### 2023-11-06 manuscript_BiSC.tex
Initial Overleaf Import

\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}
\usepackage{caption}
% Set page size and margins
% Replace `letterpaper' with`a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{authblk} % this is for multiple authors 
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{cleveref}

% Configure for "Fig.":
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Fig.}{Figs.}

% reference configuration
\usepackage[authoryear,round]{natbib}
\bibliographystyle{apalike} % for an APA-like style
\usepackage{amsmath}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Tittle of the paper --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Authors and affiliations --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% --- Authors --- %%%
\author[1, 2]{Alan Domínguez}
\author[1, 3, 4]{Payam Dadvand}
\author[1]{Marta Cirach}
\author[1]{Bruno Raimbault}
\author[1]{Toni Galmes}
\author[1]{Karl Samuelsson}
\author[1, 2, 3]{Mark Nieuwenhuijsen}
\author[1, 2, 3]{Jordi Sunyer}
\author[1, 2, 3]{Xavier Basagaña}
\author[1, 3]{Ioar Rivas}

%%% --- Affiliations --- %%%
\affil[1]{Barcelona Institute for Global Health (ISGlobal), Barcelona, Spain.}
\affil[2]{Universitat Popmpeu Fabra (UPF), Barcelona, Spain.}
\affil[3]{CIBER Epidemiología y Salud Pública (CIBERESP), Madrid, Spain.}
\affil[4]{London School of Hygiene and Tropical Medicine (LSHTM), London, UK.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Abstract --- %%%
%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Air pollution is the leading environmental risk factor for human health worldwide. A detailed and comprehensive exposure assessment of outdoor air pollution considering both spatial and temporal variability is fundamental to conduct epidemiological studies.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Introduction --- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Air pollution stands as the primary environmental risk factor for human health and the fourth most deadly health risk worldwide, playing a significant role in the global burden of disease \cite{cohen2017, he2020state}. Exposure to ambient air pollution has been linked with a extensive array of adverse health outcomes \cite{boogaard2022, guxens2022hei, haddad2023}. In this context urbanization have been playing a major role increasing air pollution exposure as cities grow and more population moves to urban areas \cite{nieuwenhuijsen2016}. This trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas \cite{boogaard2022}.

Over the past few decades, several modelling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution \cite{hoek2017, di2019no2, di2019pm25, stafoggia2019, stafoggia2020}. Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches for estimating air pollution exposure concentrations in epidemiological studies \cite{gulliver2015, dehoogh2014}. LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that works as proxies for traffic related air pollution \cite{briggs1997, hoek2008}. In contrast, DMs rely on mathematical simulations to estimate air pollution levels. These predictions are based on data from emissions sources, chemical transformations models, pollutant dispersion models, topography, roughness, and meteorological data \cite{hoek2017}. Although these methods have been extensively used they come with inherent limitations related to lack of temporal and spatial variation. In order to overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}.

While significant advancement have been made in spatiotemporal modelling of air pollution over large-scale domains, such models (multi-country and national domain models) often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}. This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potential lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.

Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patternts that difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}. In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees based models \cite{liu2022}. These family of models, are often used due to their ability to model non-linear and potential interaction among predictive variables \cite{liu2022treebased}. Among the different ML algorithm, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.

Furthermore, the COVID-19 pandemic has introduced unprecedented challenges for air pollution modelling and exposure assessment. These challenges encompass changes in air pollution patterns, constraints on data collection due to health and safety restrictions (lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}.

In this study, within the framework of the Barcelona Life Study Cohort (BiSC), we developed and compared LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses. LUR and DM, were constructed using conventional state-of the art techniques, HM was innovetely designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model. The overall aim is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates, intended to use in the prenatal exposure assessment within the BiSC cohort. Additionally, we aim to evaluate the impact of the COVID-19 pandemic on these outdoor air pollution exposure assessments.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Material and methods--- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Study domain and Monitoring Campaigns --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study domain and monitoring campaigns}

Barcelona (\textbf{\cref{fig1}}), is one of the most populated and dense cities in Spain, with 1,656,725 inhabitants ($\approx16,255.4$ inhabitants/km\textsuperscript{2}). It is located on the Mediterranean coast of the Iberian peninsula, protected by the Collserola mountain range and delimited by two delta rives (Besòs and LLobregat). The city has a Mediterranean climate, the wind patterns are dominated by a breeze that blows from the sea during the daytime and from the land during nigh-time. Barcelona is one of the cities in Europe with the highest vehicle traffic density (6,000 vehicles/km\textsuperscript{2}) \cite{casallas2018}.

% Put the figure text closer to the Figure 1
\captionsetup[figure]{skip=-4pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/fig1_op2.png}
\caption{Study domain and spatial distribution (BiSC participants, BiSCAPE campaign, monitoring stations, and traffic points)}
\label{fig1}
\end{figure}

Air pollution measurements were conducted during two separated monitoring campaigns. The first campaign was carried out in accordance with the ESCAPE guidelines. This included four distinct measurements periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022. During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet). This monitoring campaign was named "BiSCAPE" campaign. The second monitoring campaign was conducted at homes of the BiSC participants, one week during the first trimester and one week during the third trimester of pregnancy, during the period 2018-2021, and only NO\textsubscript{2} levels were measured. This monitoring campaign was named "BiSC-home" campaign. These monitoring campaigns where used in the development of the LUR and hybrid models.

%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- LUR model --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LUR models}

LUR models for NO\textsubscript{2}  were developed using only data from the locations that had measured home-outdoor (BiSC-home campaign) levels for both sampling campaigns (i.e., we did not use data from participants who had data available for only one campaign, either first or third trimester). Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites during the pandemic period and its resulting restrictions. For PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC models were developed using BiSCAPE data from the 34 sites.

For the development of the LUR models, spatial variables that better represent the pollutant concentration distribution across the study area were created. Following the ESCAPE guidelines \cite{eeftens2012, beelen2013},  predictor variables were selected from three broad categories: traffic, land use, and urban configuration. A wide variety of buffer sizes were created, i.e, 25m, 50m, 100m, 300m, 500m, and 1000m to capture the area-specific predictor variable at every single location.

We applied a supervised stepwise process in which we sequentially add a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model. This procedure was followed to built all the specific pollutant models. Previously to model development, we specified the direction expected of the effect that each predictor have in the model. For all predictor variables, univariate regression models with annual average air pollution concentration (NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC) were conducted, models with the highest adjusted explained variance (R\textsuperscript{2}) were selected. We then repeated the process using the previous model and all the remaining predictor variables. Details of the LUR modelling and variables used could be found in the \textbf{Supplementary material X}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Dispersion model --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dispersion models}

An emission inventory was compiled for the period 2018-2021, which included data from diverse emission sources such as vehicular traffic, household activities, industrial plants, airports, harbours, factories, and the commercial sector. The Atmospheric Dispersion Modelling System software (ADMS-Urban) from Cambridge Environmental Research Consultants (CERC) was employed to calculate the concentration of pollutants like NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC originating from these sources. ADMS-Urban is a specialized Gaussian model designed for analyzing pollutant emissions in urban and metropolitan areas, particularly at road level \cite{mchugh1997adms}.

ADMS-Urban employs various algorithms to assess the chemical transportation and dispersion of pollutants, factoring in ground-level turbulence and turbulence induced by surrounding buildings \cite{stocker2012adms}. The model also features a meteorological pre-processing system that calculates fluxes and turbulence, enabling it to model airflow over complex terrains.

For enhanced accuracy, the dispersion model integrates data from a three-dimensional building and street width model from the Institut Cartogràfic i Geològic de Catalunya \cite{stoter2016state, 3Dcity}. This data aids in the calculation of the street canyon effect, a phenomenon characterized by tall buildings situated adjacent to heavily-trafficked roads, which hinders atmospheric dispersion, leading to higher and relatively consistent pollutant concentrations compared to areas unaffected by the canyon effect.

The model also utilized meteorological data including wind speed, direction, air temperature, cloud cover, solar radiation, relative humidity, precipitation rate, and Planetary Boundary Layer Height (PBLH). This meteorological data was collected from the Zona Universitària station in Barcelona, managed by the Servei Meteorològic de Catalunya \cite{xema2013}.

In order to incorporate background pollution levels at a regional scale, data from Copernicus Atmosphere Monitoring Service Information (CAMS) was used. Background air pollution data has been extracted from the CAMS reanalysis ensemble model for 2018, and from CAMS ensemble forecast model for 2019-2020 \cite{cams2020, franceinstitut}.

In addition, bias-adjustment techniques have been used to fit the model data to the observations. Through hourly bias spatial interpolation and generation of hourly maps. After that, data were processed  by cluster analysis techniques in order to obtain a mean cluster map. Each bias map by pollutant were associated to a cluster map class and an hourly-varying profile.

Further details regarding the ADMS-Urban model and the variables used can be found in \textbf{Supplementary materials X}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Hybrid model --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hybrid models}

For the hybrid models, we sought to improve their predictive power by integrating comprehensive data from multiple information sources and leveraging ML. For NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC were largely based on daily concentrations from dispersion models, summarized as weekly averages, and spatial predictor variables from the LUR models.

For a more comprehensive approach, we integrated additional data from the Catalan Air Pollution Monitoring and Forecasting Network (XVPCA) for Barcelona and surrounding areas \cite{xarxa2012}. This data was interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}. The data came from 16 monitoring stations, including: Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St. Adria, St. Coloma, Vall Hebron, Zona Universitária \cite{xarxa2012}.

While our dispersion model initially utilized meteorological and traffic data, the hybrid model broadened its scope by integrating daily traffic counts from 180 locations throughout the city \cite{trafficbcn}. This broader data helps us better understand the spatial and time-specific variations in urban air pollution \cite{pinto2020}. To further refine our insights into the influence of weather patterns on pollution, we sourced additional meteorological variables, such as wind speed and air temperature, from the Raval monitoring station managed by Servei Meteorològic de Catalunya \cite{xema2013}.

We employed the RF algorithm to capture non-linear relationships and potential interactions between predictor variables and the response variable, utilizing \textit{the ranger package} in R implemented in the \textit{caret} framework \cite{wright2019}, a set of functions that attempt to streamline the process for creating predictive models \cite{caret2008}. Briefly, RF is a machine learning (ML) algorithm that construct an ensemble of multiple decisions trees \cite{breiman2001}. In this approach, each tree is formulated using a bootstrap sample of the data. During this process, every node within a tree represents the optimal split from a randomly selected subset of predictors. Predictions are then derived by averaging the outputs from all trees.

In summary, by combining data from multiple sources, our hybrid models aim to provide more accurate and reliable predictions. These predictions are essential for studying and mitigating urban air pollution. The distribution of the air pollution monitoring stations, meteorological stations and traffic count measures are shown in the (\textbf{\cref{fig1}}, \textbf{Panel D}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Model Evaluation and Comparison--- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Performance assessment}

We evaluated model performance using various statistical approaches and on different temporal bases. The LUR models were evaluated on an annual basis, dispersion models were assessed both hourly and daily, while the hybrid models were evaluated on a weekly basis.

\subsubsection{LUR model evaluation}
Performance of the LUR models was assessed using \textit{k-fold cross-validation}. Specifically, we employed a 10-fold cross-validation technique. In this approach, the data was split into ten distinct sub-samples. For each iteration, the model was trained on k-1 folds and then validated using the leftover fold \cite{ziegel2003}.For every fold, we computed metrics such as \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)). After each iteration, both the error and adjusted \(R\textsuperscript{2}\) were recorded. This process was repeated \(k-times\), and the average of these measures was reported as the indicators of the model validity. 
\vspace{0.5cm}

These metrics are defined as follows:

% --- R2 --- %
\begin{equation}
R^2 = \left( \frac{\sum (\hat{y}_i - \bar{y})^2}{\sum (y_i - \bar{y})^2} \right)
\end{equation}

% --- root mean squared error (RMSE) --- % 
\begin{equation}
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)^2}
\end{equation}

% --- mean bias (MB) --- % 
\begin{equation}
\text{Mean Bias} = \frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)
\end{equation}

% --- correlation coefficient (r) --- %
\begin{equation}
r = \frac{\sum (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum (x_i - \bar{x})^2 \sum (y_i - \bar{y})^2}}
\end{equation}
\vspace{0.5 cm}

We carried out both internal and external validation for the NO\textsubscript{2} LUR model. Internal validation was conducted by applying a K-fold cross-validation analysis. The 10-fold cross-validation procedure and its significance have been detailed above. For the external validation of the NO\textsubscript{2} LUR model, we applied data from BiSC-home participants' homes for those participants who had data available only for one measurement campaign, and therefore their data were not used for developing NO\textsubscript{2} LUR model. For LUR models BC, PM\textsubscript{2.5}, and its constituents (Cu, Fe, Sb, and Zn), internal validation was conducted using the \textit{leave-one-out-cross-validation (LOOCV)} process. For these models, we were not able to conduct external validation due to the lack of external data.

\subsection{Hybrid model evaluation}
We assessed the performance of each hybrid model by contrasting predictions with observations. Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and the \textit{out-of-bag (oob) score}. While both methods vary in the sample size designated for training and testing, their core difference lies in the data splitting strategy. In the 10-fold CV, the model trains on nine out of the ten data partitions and then validates against the remaining one. In contrast, the oob approach trains the model on two-thirds of the data and uses the residual third for external validation. We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}

\subsection{Dispersion model evaluation}
Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations (XVPCA), including: Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.\vspace{0.5cm}

The metrics used for the dispersion models we:

% --- Percentage of observed-modelled pairs (n) --- %
\begin{equation}
n = \frac{N_{\text{pairs used}}}{N_{\text{total pairs}}} \times 100
\end{equation}

% --- Factor of Two (FAC2) --- % 
\begin{equation}
\text{FAC2} = \frac{1}{n} \sum_{i=1}^{n} U,
\end{equation}
%Where \( U \) is 1 if \( 0.5 \leq \frac{M_i}{O_i} \leq 2 \) and 0 otherwise.

% --- Mean Bias (MB) --- % 
\begin{equation}
\text{MB} = \frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)
\end{equation}

% --- Mean Gross Error (MGE) --- % 
\begin{equation}
\text{MGE} = \frac{1}{n} \sum_{i=1}^{n} |O_i - M_i|
\end{equation}

% --- Normalized Mean Bias (NMB) --- % 
\begin{equation}
\text{NMB} = \frac{\sum_{i=1}^{n} (M_i - O_i)}{\sum_{i=1}^{n} O_i}
\end{equation}

% --- Root Mean Square Error (RMSE) --- % 
\begin{equation}
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)^2}
\end{equation}

% --- Correlation Coefficient (r) --- % 
\begin{equation}
r = \frac{\sum_{i=1}^{n} (O_i - \bar{O})(M_i - \bar{M})}{\sqrt{\sum_{i=1}^{n} (O_i - \bar{O})^2 \sum_{i=1}^{n} (M_i - \bar{M})^2}}
\end{equation}
\vspace{0.5cm}

%%% --- Statistical Methods --- %%% 
All statistical analyses were performed in R version 4.3.1 \cite{Rstudio}. GIS predictor variables were computed using X software.

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Results --- %%%
%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

\textbf{\cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5} and BC concentrations from the LUR, DM, and HM. The data shown are long-term estimated concentrations of pollutants, after averaging weekly estimates for all the study period (2019 - 2021).

%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/fig_HM_test_all_models.png}
\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
\label{fig2}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
We successfully developed spatiotemporal models using three different approaches (LUR, dispersion and hybrid models). All models had moderate to good performance evaluated by the defined metrics in the \textbf{methods section}.

As we expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed. We believe that by combining multiple data sources, and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.

\subsection{Comparison between LUR, DM, and Hybrid model}

\subsection{Comparison with Previous studies}

\subsection{Strengths and Limitations}
One of the main strengths of this study is the extensive exposure assessment

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Conclusions --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

Air pollution exposure concentration estimates from LUR, dispersion and hybrid models were moderately to strongly correlated. We observed

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- References --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

\bibliography{references}

\end{document}

### 2023-11-06 supplementary.tex
Initial Overleaf Import

--- 2023-11-06 manuscript_BiSC.tex
@@
 \bibliographystyle{apalike} % for an APA-like style
-\usepackage{amsmath}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}
+\title{Suppementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}
 
@@
 \author[1, 2]{Alan Domínguez}
-\author[1, 3, 4]{Payam Dadvand}
 \author[1]{Marta Cirach}
+\author[1,3]{Ioar Rivas}
 \author[1]{Bruno Raimbault}
 \author[1]{Toni Galmes}
-\author[1]{Karl Samuelsson}
-\author[1, 2, 3]{Mark Nieuwenhuijsen}
-\author[1, 2, 3]{Jordi Sunyer}
-\author[1, 2, 3]{Xavier Basagaña}
-\author[1, 3]{Ioar Rivas}
+\author[1,2,3]{Jordi Sunyer}
+\author[1, 3, 4]{Payam Dadvand}
+\author[1,2,3]{Xavier Basagaña}
 
@@
 \affil[3]{CIBER Epidemiología y Salud Pública (CIBERESP), Madrid, Spain.}
-\affil[4]{London School of Hygiene and Tropical Medicine (LSHTM), London, UK.}
+\affil[4]{London School of Hygiene and Tropical Medicine ( LSHTM), London, UK.}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Abstract --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-
-Air pollution is the leading environmental risk factor for human health worldwide.
-A detailed and comprehensive exposure assessment of outdoor air pollution considering both spatial and temporal variability is fundamental to conduct epidemiological studies.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Introduction --- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
 
-Air pollution stands as the primary environmental risk factor for human health and the fourth most deadly health risk worldwide, playing a significant role in the global burden of disease \cite{cohen2017, he2020state}.
-Exposure to ambient air pollution has been linked with a extensive array of adverse health outcomes \cite{boogaard2022, guxens2022hei, haddad2023}.
-In this context urbanization have been playing a major role increasing air pollution exposure as cities grow and more population moves to urban areas \cite{nieuwenhuijsen2016}.
-This trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas \cite{boogaard2022}.
+Air pollution stands as the primary environmental risk factor for human health \cite{} and plays a significant role in the global burden of disease \cite{}.
+Exposure to ambient air pollution has been linked with a extensive array of adverse health outcomes\cite{}.
+Urbanization has played a pivotal in this context, increasing air pollution exposure as cities grow and more population moves to urban areas\cite{}.
+This trend presents challenges for epidemiological studies, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the short- and long-term health impacts of air pollution\cite{}.
+Recent years have witnessed advancements in spatiotemporal modelling of air pollution across large-scale domains.
+While European-wide models
+aimed to assess long-term air pollution exposure effects, their ability to predict at micro-scales has been limited\cite{}.
+Therefore, a detailed and comprehensive assessment of air pollution exposure at smaller areas is essential for understanding its impact on health, especially during critical developmental phases\cite{}.
 
-Over the past few decades, several modelling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution \cite{hoek2017, di2019no2, di2019pm25, stafoggia2019, stafoggia2020}.
-Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches for estimating air pollution exposure concentrations in epidemiological studies \cite{gulliver2015, dehoogh2014}.
-LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that works as proxies for traffic related air pollution \cite{briggs1997, hoek2008}.
+Over the past few decades, several modelling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution\cite{}.
+Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches for estimating air pollution exposure concentrations in cohort studies.
+LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that works as proxies for traffic related air pollution\cite{}.
 In contrast, DMs rely on mathematical simulations to estimate air pollution levels.
-These predictions are based on data from emissions sources, chemical transformations models, pollutant dispersion models, topography, roughness, and meteorological data \cite{hoek2017}.
-Although these methods have been extensively used they come with inherent limitations related to lack of temporal and spatial variation.
-In order to overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}.
+These predictions are based on data from emissions sources, chemical transformations, pollutant dispersion, topography, roughness, and meteorological data\cite{}.
+In order to overcome the shortcomings of the aforementioned approaches, hybrid models (HM) have emerged recently, blending both spatial and temporal advantages, leading to refined and enhanced estimations\cite{}.
+These models combined outputs from dispersion, land use variables, and  data from monitoring station\cite{}.
 
-While significant advancement have been made in spatiotemporal modelling of air pollution over large-scale domains, such models (multi-country and national domain models) often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
-This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potential lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.
+Further complexity arise in urban areas when modelling air pollution, due to complex spatiotemporal patterns that difficult accurate estimation of ground-level air pollutant concentrations\cite{}.
+As a consequence, machine learning (ML) algorithms have been used, particularly decision trees based methods used due their ability to model non-linear and potential interaction among predictive variables \cite{}.
+Among the different ML algorithm, random forest has emerged as one of the most robust techniques for predicting air pollution concentrations\cite{}.
 
-Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patternts that difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}.
-In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees based models \cite{liu2022}.
-These family of models, are often used due to their ability to model non-linear and potential interaction among predictive variables \cite{liu2022treebased}.
-Among the different ML algorithm, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.
-
-Furthermore, the COVID-19 pandemic has introduced unprecedented challenges for air pollution modelling and exposure assessment.
-These challenges encompass changes in air pollution patterns, constraints on data collection due to health and safety restrictions (lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}.
-
-In this study, within the framework of the Barcelona Life Study Cohort (BiSC), we developed and compared LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses.
-LUR and DM, were constructed using conventional state-of the art techniques, HM was innovetely designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
-The overall aim is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates, intended to use in the prenatal exposure assessment within the BiSC cohort.
-Additionally, we aim to evaluate the impact of the COVID-19 pandemic on these outdoor air pollution exposure assessments.
+In this study, we developed spatiotemporal models using different methodologies (LUR, DM, and HM) to estimate outdoor air pollution exposure for specific air pollutants, all under the framework of the Barcelona Life Study Cohort (BiSC).
+While LUR and DM, will be built using conventional techniques, the HM will harness the Random Forest (RF) algorithm, integrating outputs from the DM and explanatory variables from the LUR model.
+The overall objective is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates for use in prenatal exposure assessment for the BiSC cohort study.
 
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/fig1_op2.png}
-\caption{Study domain and spatial distribution (BiSC participants, BiSCAPE campaign, monitoring stations, and traffic points)}
+\includegraphics[width=1\textwidth]{figures/bcn_map.png}
+\caption{ Study domain and spatial distribution of the BiSC participants (n = 1,083)}
 \label{fig1}
@@
 The second monitoring campaign was conducted at homes of the BiSC participants, one week during the first trimester and one week during the third trimester of pregnancy, during the period 2018-2021, and only NO\textsubscript{2} levels were measured.
-This monitoring campaign was named "BiSC-home" campaign.
-These monitoring campaigns where used in the development of the LUR and hybrid models.
+This monitoring campaign was named "BiSC-Home" campaign.
+These monitoring campaigns where used in the development of the LUR and HM models.
 
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-LUR models for NO\textsubscript{2}  were developed using only data from the locations that had measured home-outdoor (BiSC-home campaign) levels for both sampling campaigns (i.e., we did not use data from participants who had data available for only one campaign, either first or third trimester).
+LUR models for NO\textsubscript{2}  were developed using only data from the locations that had measured home-outdoor (BiSC-Home campaign) levels for both sampling campaigns (i.e., we did not use data from participants who had data available for only one campaign, either first or third trimester).
 Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites during the pandemic period and its resulting restrictions.
 For PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC models were developed using BiSCAPE data from the 34 sites.
+In addition, for the development of the LUR models, spatial variables that better represent the pollutant concentration distribution across the study area were created.
+Following the ESCAPE guidelines, we selected predictor variables from three broad categories: traffic, land use, and urban configuration.
+We selected a wide variety of buffer sizes, i.e, 25m, 50m, 100m, 300m, 500m, and 1000m to capture the area-specific predictor variable at every single location.
 
-For the development of the LUR models, spatial variables that better represent the pollutant concentration distribution across the study area were created.
-Following the ESCAPE guidelines \cite{eeftens2012, beelen2013},  predictor variables were selected from three broad categories: traffic, land use, and urban configuration.
-A wide variety of buffer sizes were created, i.e, 25m, 50m, 100m, 300m, 500m, and 1000m to capture the area-specific predictor variable at every single location.
-
-We applied a supervised stepwise process in which we sequentially add a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model.
+Following the ESCAPE guidelines \cite{eeftens2012, beelen2013}, we applied a supervised stepwise process in which we sequentially add a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model.
 This procedure was followed to built all the specific pollutant models.
 Previously to model development, we specified the direction expected of the effect that each predictor have in the model.
-For all predictor variables, univariate regression models with annual average air pollution concentration (NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC) were conducted, models with the highest adjusted explained variance (R\textsuperscript{2}) were selected.
-We then repeated the process using the previous model and all the remaining predictor variables.
-Details of the LUR modelling and variables used could be found in the \textbf{Supplementary material X}.
 
@@
 
-An emission inventory was compiled for the period 2018-2021, which included data from diverse emission sources such as vehicular traffic, household activities, industrial plants, airports, harbours, factories, and the commercial sector.
-The Atmospheric Dispersion Modelling System software (ADMS-Urban) from Cambridge Environmental Research Consultants (CERC) was employed to calculate the concentration of pollutants like NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC originating from these sources.
-ADMS-Urban is a specialized Gaussian model designed for analyzing pollutant emissions in urban and metropolitan areas, particularly at road level \cite{mchugh1997adms}.
-
-ADMS-Urban employs various algorithms to assess the chemical transportation and dispersion of pollutants, factoring in ground-level turbulence and turbulence induced by surrounding buildings \cite{stocker2012adms}.
-The model also features a meteorological pre-processing system that calculates fluxes and turbulence, enabling it to model airflow over complex terrains.
-
-For enhanced accuracy, the dispersion model integrates data from a three-dimensional building and street width model from the Institut Cartogràfic i Geològic de Catalunya \cite{stoter2016state, 3Dcity}.
-This data aids in the calculation of the street canyon effect, a phenomenon characterized by tall buildings situated adjacent to heavily-trafficked roads, which hinders atmospheric dispersion, leading to higher and relatively consistent pollutant concentrations compared to areas unaffected by the canyon effect.
-
-The model also utilized meteorological data including wind speed, direction, air temperature, cloud cover, solar radiation, relative humidity, precipitation rate, and Planetary Boundary Layer Height (PBLH).
-This meteorological data was collected from the Zona Universitària station in Barcelona, managed by the Servei Meteorològic de Catalunya \cite{xema2013}.
-
-In order to incorporate background pollution levels at a regional scale, data from Copernicus Atmosphere Monitoring Service Information (CAMS) was used.
-Background air pollution data has been extracted from the CAMS reanalysis ensemble model for 2018, and from CAMS ensemble forecast model for 2019-2020 \cite{cams2020, franceinstitut}.
-
-In addition, bias-adjustment techniques have been used to fit the model data to the observations.
-Through hourly bias spatial interpolation and generation of hourly maps.
-After that, data were processed  by cluster analysis techniques in order to obtain a mean cluster map.
-Each bias map by pollutant were associated to a cluster map class and an hourly-varying profile.
-
-Further details regarding the ADMS-Urban model and the variables used can be found in \textbf{Supplementary materials X}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-For the hybrid models, we sought to improve their predictive power by integrating comprehensive data from multiple information sources and leveraging ML.
-For NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC were largely based on daily concentrations from dispersion models, summarized as weekly averages, and spatial predictor variables from the LUR models.
-
-For a more comprehensive approach, we integrated additional data from the Catalan Air Pollution Monitoring and Forecasting Network (XVPCA) for Barcelona and surrounding areas \cite{xarxa2012}.
-This data was interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
-The data came from 16 monitoring stations, including: Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St.
-Adria, St.
-Coloma, Vall Hebron, Zona Universitária \cite{xarxa2012}.
-
-While our dispersion model initially utilized meteorological and traffic data, the hybrid model broadened its scope by integrating daily traffic counts from 180 locations throughout the city \cite{trafficbcn}.
-This broader data helps us better understand the spatial and time-specific variations in urban air pollution \cite{pinto2020}.
-To further refine our insights into the influence of weather patterns on pollution, we sourced additional meteorological variables, such as wind speed and air temperature, from the Raval monitoring station managed by Servei Meteorològic de Catalunya \cite{xema2013}.
-
-We employed the RF algorithm to capture non-linear relationships and potential interactions between predictor variables and the response variable, utilizing \textit{the ranger package} in R implemented in the \textit{caret} framework \cite{wright2019}, a set of functions that attempt to streamline the process for creating predictive models \cite{caret2008}.
-Briefly, RF is a machine learning (ML) algorithm that construct an ensemble of multiple decisions trees \cite{breiman2001}.
-In this approach, each tree is formulated using a bootstrap sample of the data.
-During this process, every node within a tree represents the optimal split from a randomly selected subset of predictors.
-Predictions are then derived by averaging the outputs from all trees.
-
-In summary, by combining data from multiple sources, our hybrid models aim to provide more accurate and reliable predictions.
-These predictions are essential for studying and mitigating urban air pollution.
-The distribution of the air pollution monitoring stations, meteorological stations and traffic count measures are shown in the (\textbf{\cref{fig1}}, \textbf{Panel D}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Model Evaluation and Comparison--- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Performance assessment}
-
-We evaluated model performance using various statistical approaches and on different temporal bases.
-The LUR models were evaluated on an annual basis, dispersion models were assessed both hourly and daily, while the hybrid models were evaluated on a weekly basis.
-
-\subsubsection{LUR model evaluation}
-Performance of the LUR models was assessed using \textit{k-fold cross-validation}.
-Specifically, we employed a 10-fold cross-validation technique.
-In this approach, the data was split into ten distinct sub-samples.
-For each iteration, the model was trained on k-1 folds and then validated using the leftover fold \cite{ziegel2003}.For every fold, we computed metrics such as \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)).
-After each iteration, both the error and adjusted \(R\textsuperscript{2}\) were recorded.
-This process was repeated \(k-times\), and the average of these measures was reported as the indicators of the model validity.
-\vspace{0.5cm}
-
-These metrics are defined as follows:
-
-% --- R2 --- %
-\begin{equation}
-R^2 = \left( \frac{\sum (\hat{y}_i - \bar{y})^2}{\sum (y_i - \bar{y})^2} \right)
-\end{equation}
-
-% --- root mean squared error (RMSE) --- % 
-\begin{equation}
-\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)^2}
-\end{equation}
-
-% --- mean bias (MB) --- % 
-\begin{equation}
-\text{Mean Bias} = \frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)
-\end{equation}
-
-% --- correlation coefficient (r) --- %
-\begin{equation}
-r = \frac{\sum (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum (x_i - \bar{x})^2 \sum (y_i - \bar{y})^2}}
-\end{equation}
-\vspace{0.5 cm}
-
-We carried out both internal and external validation for the NO\textsubscript{2} LUR model.
-Internal validation was conducted by applying a K-fold cross-validation analysis.
-The 10-fold cross-validation procedure and its significance have been detailed above.
-For the external validation of the NO\textsubscript{2} LUR model, we applied data from BiSC-home participants' homes for those participants who had data available only for one measurement campaign, and therefore their data were not used for developing NO\textsubscript{2} LUR model.
-For LUR models BC, PM\textsubscript{2.5}, and its constituents (Cu, Fe, Sb, and Zn), internal validation was conducted using the \textit{leave-one-out-cross-validation (LOOCV)} process.
-For these models, we were not able to conduct external validation due to the lack of external data.
-
-\subsection{Hybrid model evaluation}
-We assessed the performance of each hybrid model by contrasting predictions with observations.
-Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and the \textit{out-of-bag (oob) score}.
-While both methods vary in the sample size designated for training and testing, their core difference lies in the data splitting strategy.
-In the 10-fold CV, the model trains on nine out of the ten data partitions and then validates against the remaining one.
-In contrast, the oob approach trains the model on two-thirds of the data and uses the residual third for external validation.
-We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}
-
-\subsection{Dispersion model evaluation}
-Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations (XVPCA), including: Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.\vspace{0.5cm}
-
-The metrics used for the dispersion models we:
-
-% --- Percentage of observed-modelled pairs (n) --- %
-\begin{equation}
-n = \frac{N_{\text{pairs used}}}{N_{\text{total pairs}}} \times 100
-\end{equation}
-
-% --- Factor of Two (FAC2) --- % 
-\begin{equation}
-\text{FAC2} = \frac{1}{n} \sum_{i=1}^{n} U,
-\end{equation}
-%Where \( U \) is 1 if \( 0.5 \leq \frac{M_i}{O_i} \leq 2 \) and 0 otherwise.
-
-% --- Mean Bias (MB) --- % 
-\begin{equation}
-\text{MB} = \frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)
-\end{equation}
-
-% --- Mean Gross Error (MGE) --- % 
-\begin{equation}
-\text{MGE} = \frac{1}{n} \sum_{i=1}^{n} |O_i - M_i|
-\end{equation}
-
-% --- Normalized Mean Bias (NMB) --- % 
-\begin{equation}
-\text{NMB} = \frac{\sum_{i=1}^{n} (M_i - O_i)}{\sum_{i=1}^{n} O_i}
-\end{equation}
-
-% --- Root Mean Square Error (RMSE) --- % 
-\begin{equation}
-\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)^2}
-\end{equation}
-
-% --- Correlation Coefficient (r) --- % 
-\begin{equation}
-r = \frac{\sum_{i=1}^{n} (O_i - \bar{O})(M_i - \bar{M})}{\sqrt{\sum_{i=1}^{n} (O_i - \bar{O})^2 \sum_{i=1}^{n} (M_i - \bar{M})^2}}
-\end{equation}
-\vspace{0.5cm}
-
-%%% --- Statistical Methods --- %%% 
-All statistical analyses were performed in R version 4.3.1 \cite{Rstudio}. GIS predictor variables were computed using X software.
-
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Results --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-\textbf{\cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5} and BC concentrations from the LUR, DM, and HM.
-The data shown are long-term estimated concentrations of pollutants, after averaging weekly estimates for all the study period (2019 - 2021).
-
-%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
-% Put the figure text closer to the Figure 3
-\captionsetup[figure]{skip=6pt}
-% We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/fig_HM_test_all_models.png}
-\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
-\label{fig2}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Discussion --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-We successfully developed spatiotemporal models using three different approaches (LUR, dispersion and hybrid models).
-All models had moderate to good performance evaluated by the defined metrics in the \textbf{methods section}.
-
-As we expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed.
-We believe that by combining multiple data sources, and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.
-
-\subsection{Comparison between LUR, DM, and Hybrid model}
-
-\subsection{Comparison with Previous studies}
-
-\subsection{Strengths and Limitations}
-One of the main strengths of this study is the extensive exposure assessment
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Conclusions --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-Air pollution exposure concentration estimates from LUR, dispersion and hybrid models were moderately to strongly correlated. We observed
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evaluation}
+We performed different assessment to evaluate the performance and accuracy of the obtained models. For LUR models
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
 

### 2023-11-07 manuscript_BiSC.tex

@@
 In this study, within the framework of the Barcelona Life Study Cohort (BiSC), we developed and compared LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses.
-LUR and DM, were constructed using conventional state-of the art techniques, HM was innovetely designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
+LUR and DM, were constructed using conventional state-of the art techniques, while HM was innovetely designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
 The overall aim is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates, intended to use in the prenatal exposure assessment within the BiSC cohort.
-Additionally, we aim to evaluate the impact of the COVID-19 pandemic on these outdoor air pollution exposure assessments.
 

### 2023-11-10 manuscript_BiSC.tex
Updates from Overleaf

@@
 
-Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patternts that difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}.
+Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}.
 In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees based models \cite{liu2022}.
-These family of models, are often used due to their ability to model non-linear and potential interaction among predictive variables \cite{liu2022treebased}.
+This family of models is often used due to its ability to model potential, nonlinear interaction between predictor variables \cite{liu2022treebased}.
 Among the different ML algorithm, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.
@@
 
-Air pollution measurements were conducted during two separated monitoring campaigns.
-The first campaign was carried out in accordance with the ESCAPE guidelines.
-This included four distinct measurements periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
-During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).
-This monitoring campaign was named "BiSCAPE" campaign.
-The second monitoring campaign was conducted at homes of the BiSC participants, one week during the first trimester and one week during the third trimester of pregnancy, during the period 2018-2021, and only NO\textsubscript{2} levels were measured.
-This monitoring campaign was named "BiSC-home" campaign.
-These monitoring campaigns where used in the development of the LUR and hybrid models.
+Air pollution measurements were conducted during two separate monitoring campaigns.
+The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021.
+The second campaign was carried out in accordance to the ESCAPE guidelines, named "BiSCAPE", included four distinct measurements periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
+During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns where used in the development of the LUR and hybrid models.
 

### 2023-11-10 manuscript_BiSC.tex
bisc logo

@@
 We then repeated the process using the previous model and all the remaining predictor variables.
-Details of the LUR modelling and variables used could be found in the \textbf{Supplementary material X}.
+Details of the LUR modelling and variables used can be found in the \textbf{Supplementary material X}.
 
@@
 
+Details from the metrics used for model validation can be found in the \textbf{Supplementary material X}.
+
 These metrics are defined as follows:
-
 % --- R2 --- %
@@
 
-\subsection{Hybrid model evaluation}
-We assessed the performance of each hybrid model by contrasting predictions with observations.
-Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and the \textit{out-of-bag (oob) score}.
-While both methods vary in the sample size designated for training and testing, their core difference lies in the data splitting strategy.
-In the 10-fold CV, the model trains on nine out of the ten data partitions and then validates against the remaining one.
-In contrast, the oob approach trains the model on two-thirds of the data and uses the residual third for external validation.
-We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}
-
 \subsection{Dispersion model evaluation}
-Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations (XVPCA), including: Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.\vspace{0.5cm}
+Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations (XVPCA), including: Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
+Details from the metrics used for dispersion model validation can be found in the \textbf{Supplementary material X}.\vspace{0.5cm}
 
@@
 
-%%% --- Statistical Methods --- %%% 
+\subsection{Hybrid model evaluation}
+We assessed the performance of each hybrid model by contrasting predictions with observations.
+Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and the \textit{out-of-bag (oob) score}.
+While both methods vary in the sample size designated for training and testing, their core difference lies in the data splitting strategy.
+In the 10-fold CV, the model trains on nine out of the ten data partitions and then validates against the remaining one.
+In contrast, the oob approach trains the model on two-thirds of the data and uses the residual third for external validation.
+We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}
+
 All statistical analyses were performed in R version 4.3.1 \cite{Rstudio}. GIS predictor variables were computed using X software.

### 2023-11-11 manuscript_BiSC.tex
Updates from Overleaf

@@
 Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches for estimating air pollution exposure concentrations in epidemiological studies \cite{gulliver2015, dehoogh2014}.
-LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that works as proxies for traffic related air pollution \cite{briggs1997, hoek2008}.
+LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that work as proxies for traffic-related air pollution \cite{briggs1997, hoek2008}.
 In contrast, DMs rely on mathematical simulations to estimate air pollution levels.
-These predictions are based on data from emissions sources, chemical transformations models, pollutant dispersion models, topography, roughness, and meteorological data \cite{hoek2017}.
+These predictions are based on data from emissions sources, chemical transformation models, pollutant dispersion models, topography, roughness, and meteorological data \cite{hoek2017}.
 Although these methods have been extensively used they come with inherent limitations related to lack of temporal and spatial variation.
-In order to overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}.
+To overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}.
 
-While significant advancement have been made in spatiotemporal modelling of air pollution over large-scale domains, such models (multi-country and national domain models) often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
+While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains, such models (multi-country and national domain models) often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
 This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potential lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.
@@
 Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}.
-In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees based models \cite{liu2022}.
+In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees-based models \cite{liu2022}.
 This family of models is often used due to its ability to model potential, nonlinear interaction between predictor variables \cite{liu2022treebased}.
-Among the different ML algorithm, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.
+Among the different ML algorithms, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.
 
@@
 In this study, within the framework of the Barcelona Life Study Cohort (BiSC), we developed and compared LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses.
-LUR and DM, were constructed using conventional state-of the art techniques, while HM was innovetely designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
-The overall aim is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates, intended to use in the prenatal exposure assessment within the BiSC cohort.
+LUR and DM, were constructed using conventional state-of-the-art techniques, while HM was innovatively designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
+The overall aim is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates, intended to be used in the prenatal exposure assessment within the BiSC cohort.
 
@@
 Barcelona (\textbf{\cref{fig1}}), is one of the most populated and dense cities in Spain, with 1,656,725 inhabitants ($\approx16,255.4$ inhabitants/km\textsuperscript{2}).
-It is located on the Mediterranean coast of the Iberian peninsula, protected by the Collserola mountain range and delimited by two delta rives (Besòs and LLobregat).
-The city has a Mediterranean climate, the wind patterns are dominated by a breeze that blows from the sea during the daytime and from the land during nigh-time.
+It is located on the Mediterranean coast of the Iberian peninsula, protected by the Collserola mountain range and delimited by two delta rivers (Besòs and LLobregat).
+The city has a Mediterranean climate, the wind patterns are dominated by a breeze that blows from the sea during the daytime and from the land during night-time.
 Barcelona is one of the cities in Europe with the highest vehicle traffic density (6,000 vehicles/km\textsuperscript{2}) \cite{casallas2018}.

### 2023-11-12 01.Code/01.-descriptive_AP.R
update code

@@ -1,0 +1,121 @@
+# Install packages
+pacman::p_load(tidyverse)
+
+##############################
+### --- Load the data --- ###
+############################
+
+#################################################
+### --- Air pollutants measurements data --- #####################################################################
+################################################
+no2_measures <- read.csv("01.Data/BiSC_comparison_data/measures/no2_final_model_data.csv")
+pm25_measures <- read.csv("01.Data/BiSC_comparison_data/measures/pm25_final_model_data.csv")
+bc_measures <- read.csv("01.Data/BiSC_comparison_data/measures/bc_final_model_data.csv")
+pm25_constituents_measures <- read.csv("01.Data/BiSC_comparison_data/measures/pm25_constituents_model_data.csv")
+
+
+# Have a quick look to the data 
+dplyr::glimpse(no2_measures)
+dplyr::glimpse(pm25_measures)
+dplyr::glimpse(bc_measures)
+dplyr::glimpse(pm25_constituents_measures)
+
+### --- selection of specific variables --- ###
+no2_measures <- no2_measures %>%
+  dplyr::select(gid, id, year, week, NO2.DirectMeasurements, lon, lat) 
+
+pm25_measures <- pm25_measures %>% 
+  dplyr::select(gid, site_id, sid, c_biscape, lon, lat)
+
+bc_measures <- bc_measures %>% 
+  dplyr::select(gid, site_id, sid, c_biscape, lon, lat)
+
+pm25_constituents_measures <- pm25_constituents_measures %>% 
+  dplyr::select(gid, site_id, sid, 
+                biscape_Fe2O3,biscape_Cu, biscape_Zn, lon, lat)
+
+### --- rename some variables --- ###
+no2_measures <- no2_measures %>% 
+  dplyr::rename(no2_measures = NO2.DirectMeasurements)
+
+pm25_measures <- pm25_measures %>% 
+  dplyr::rename(pm25_measures = c_biscape) 
+
+bc_measures <- bc_measures %>% 
+  dplyr::rename(bc_measures = c_biscape)
+
+pm25_constituents_measures <- pm25_constituents_measures %>% 
+  dplyr::rename(fe2o3_measures = biscape_Fe2O3, 
+                cu_measures = biscape_Cu, 
+                zn_measures = biscape_Zn)
+
+#######################################
+### --- Model predictions data --- #################################################################################################
+######################################
+
[…]

### 2023-11-12 02.Code/02.-descriptive2.R
commit

@@ -1,0 +1,1 @@
+# 02.descriptive_AP2

### 2023-11-12 manuscript_BiSC.tex

@@
 The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021.
-The second campaign was carried out in accordance to the ESCAPE guidelines, named "BiSCAPE", included four distinct measurements periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
-During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns where used in the development of the LUR and hybrid models.
+The second campaign was carried out in accordance with the ESCAPE guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
+During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and hybrid models.
 
@@
 LUR models for NO\textsubscript{2}  were developed using only data from the locations that had measured home-outdoor (BiSC-home campaign) levels for both sampling campaigns (i.e., we did not use data from participants who had data available for only one campaign, either first or third trimester).
-Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites during the pandemic period and its resulting restrictions.
+Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites were during the pandemic period and its resulting restrictions.
 For PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC models were developed using BiSCAPE data from the 34 sites.
@@
 
-We applied a supervised stepwise process in which we sequentially add a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model.
-This procedure was followed to built all the specific pollutant models.
-Previously to model development, we specified the direction expected of the effect that each predictor have in the model.
+We applied a supervised stepwise process in which we sequentially added a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model.
+This procedure was followed to build all the specific pollutant models.
+Previously to model development, we specified the direction expected of the effect that each predictor has in the model.
 For all predictor variables, univariate regression models with annual average air pollution concentration (NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC) were conducted, models with the highest adjusted explained variance (R\textsuperscript{2}) were selected.
@@
 
-An emission inventory was compiled for the period 2018-2021, which included data from diverse emission sources such as vehicular traffic, household activities, industrial plants, airports, harbours, factories, and the commercial sector.
+An emission inventory was compiled for the period 2018-2021, which included data from diverse emission sources such as vehicular traffic, household activities, industrial plants, airports, harbors, factories, and the commercial sector.
 The Atmospheric Dispersion Modelling System software (ADMS-Urban) from Cambridge Environmental Research Consultants (CERC) was employed to calculate the concentration of pollutants like NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC originating from these sources.
@@
 For enhanced accuracy, the dispersion model integrates data from a three-dimensional building and street width model from the Institut Cartogràfic i Geològic de Catalunya \cite{stoter2016state, 3Dcity}.
-This data aids in the calculation of the street canyon effect, a phenomenon characterized by tall buildings situated adjacent to heavily-trafficked roads, which hinders atmospheric dispersion, leading to higher and relatively consistent pollutant concentrations compared to areas unaffected by the canyon effect.
+This data aids in the calculation of the street canyon effect, a phenomenon characterized by tall buildings situated adjacent to heavily trafficked roads, which hinders atmospheric dispersion, leading to higher and relatively consistent pollutant concentrations compared to areas unaffected by the canyon effect.
 
@@
 
-In order to incorporate background pollution levels at a regional scale, data from Copernicus Atmosphere Monitoring Service Information (CAMS) was used.
+To incorporate background pollution levels at a regional scale, data from Copernicus Atmosphere Monitoring Service Information (CAMS) was used.
 Background air pollution data has been extracted from the CAMS reanalysis ensemble model for 2018, and from CAMS ensemble forecast model for 2019-2020 \cite{cams2020, franceinstitut}.
@@
 Through hourly bias spatial interpolation and generation of hourly maps.
-After that, data were processed  by cluster analysis techniques in order to obtain a mean cluster map.
-Each bias map by pollutant were associated to a cluster map class and an hourly-varying profile.
+After that, data were processed  by cluster analysis techniques to obtain a mean cluster map.
+Each bias map by pollutant was associated to a cluster map class and an hourly-varying profile.
 
@@
 This data was interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
-The data came from 16 monitoring stations, including: Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St.
+The data came from 16 monitoring stations, including Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St.
 Adria, St.
@@
 We employed the RF algorithm to capture non-linear relationships and potential interactions between predictor variables and the response variable, utilizing \textit{the ranger package} in R implemented in the \textit{caret} framework \cite{wright2019}, a set of functions that attempt to streamline the process for creating predictive models \cite{caret2008}.
-Briefly, RF is a machine learning (ML) algorithm that construct an ensemble of multiple decisions trees \cite{breiman2001}.
+Briefly, RF is a machine learning (ML) algorithm that constructs an ensemble of multiple decision trees \cite{breiman2001}.
 In this approach, each tree is formulated using a bootstrap sample of the data.
@@
 These predictions are essential for studying and mitigating urban air pollution.
-The distribution of the air pollution monitoring stations, meteorological stations and traffic count measures are shown in the (\textbf{\cref{fig1}}, \textbf{Panel D}).
+The distribution of the air pollution monitoring stations, meteorological stations, and traffic count measures are shown in the (\textbf{\cref{fig1}}, \textbf{Panel D}).
 
@@
 We evaluated model performance using various statistical approaches and on different temporal bases.
-The LUR models were evaluated on an annual basis, dispersion models were assessed both hourly and daily, while the hybrid models were evaluated on a weekly basis.
+The LUR models were evaluated on an annual basis, dispersion models were assessed both hourly and daily, and the hybrid models were evaluated on a weekly basis.
 
@@
 In this approach, the data was split into ten distinct sub-samples.
-For each iteration, the model was trained on k-1 folds and then validated using the leftover fold \cite{ziegel2003}.For every fold, we computed metrics such as \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)).
+For each iteration, the model was trained on k-1 folds and then validated using the leftover fold \cite{ziegel2003}.
+For every fold, we computed metrics such as \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)).
 After each iteration, both the error and adjusted \(R\textsuperscript{2}\) were recorded.
-This process was repeated \(k-times\), and the average of these measures was reported as the indicators of the model validity.
+This process was repeated \(k-times\), and the average of these measures was reported as the indicator of the model validity.
 \vspace{0.5cm}
@@
 \subsection{Dispersion model evaluation}
-Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations (XVPCA), including: Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
+Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
 Details from the metrics used for dispersion model validation can be found in the \textbf{Supplementary material X}.\vspace{0.5cm}
@@
 We assessed the performance of each hybrid model by contrasting predictions with observations.
-Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and the \textit{out-of-bag (oob) score}.
-While both methods vary in the sample size designated for training and testing, their core difference lies in the data splitting strategy.
+Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and the \textit{out-of-bag (OBB) score}.
+While both methods vary in the sample size designated for training and testing, their core difference lies in the data-splitting strategy.
 In the 10-fold CV, the model trains on nine out of the ten data partitions and then validates against the remaining one.
-In contrast, the oob approach trains the model on two-thirds of the data and uses the residual third for external validation.
+In contrast, the OOB approach trains the model on two-thirds of the data and uses the residual third for external validation.
 We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}

### 2023-11-12 02.Code/01.-descriptive_AP.R
bc

@@ -5,9 +5,9 @@ pacman::p_load(tidyverse)
 ### --- Load the data --- ###
 ############################
 
-#################################################
+#############################################
 ### --- Air pollutants measurements data --- #####################################################################
-################################################
+############################################
 no2_measures <- read.csv("01.Data/BiSC_comparison_data/measures/no2_final_model_data.csv")
 pm25_measures <- read.csv("01.Data/BiSC_comparison_data/measures/pm25_final_model_data.csv")
 bc_measures <- read.csv("01.Data/BiSC_comparison_data/measures/bc_final_model_data.csv")
@@ -72,8 +72,11 @@ no2_hybrid <- no2_hybrid %>%
               dplyr::rename(no2_hm = no2.total_hybridmodel)
 
 pm25_hybrid <- pm25_hybrid %>% 
-               dplyr::rename()
+               dplyr::rename(pm25_hm = pm25.total_hybridmodel)
                 
+bc_hybrid <- bc_hybrid %>%  
+             dplyr::rename(bc_hm = bc.total_hybridmodel)
+
 
 
 

### 2023-11-12 01.Code/01.-descriptive_AP.R
Delete 01.Code directory

@@ -1,121 +1,0 @@
-# Install packages
-pacman::p_load(tidyverse)
-
-##############################
-### --- Load the data --- ###
-############################
-
-#################################################
-### --- Air pollutants measurements data --- #####################################################################
-################################################
-no2_measures <- read.csv("01.Data/BiSC_comparison_data/measures/no2_final_model_data.csv")
-pm25_measures <- read.csv("01.Data/BiSC_comparison_data/measures/pm25_final_model_data.csv")
-bc_measures <- read.csv("01.Data/BiSC_comparison_data/measures/bc_final_model_data.csv")
-pm25_constituents_measures <- read.csv("01.Data/BiSC_comparison_data/measures/pm25_constituents_model_data.csv")
-
-
-# Have a quick look to the data 
-dplyr::glimpse(no2_measures)
-dplyr::glimpse(pm25_measures)
-dplyr::glimpse(bc_measures)
-dplyr::glimpse(pm25_constituents_measures)
-
-### --- selection of specific variables --- ###
-no2_measures <- no2_measures %>%
-  dplyr::select(gid, id, year, week, NO2.DirectMeasurements, lon, lat) 
-
-pm25_measures <- pm25_measures %>% 
-  dplyr::select(gid, site_id, sid, c_biscape, lon, lat)
-
-bc_measures <- bc_measures %>% 
-  dplyr::select(gid, site_id, sid, c_biscape, lon, lat)
-
-pm25_constituents_measures <- pm25_constituents_measures %>% 
-  dplyr::select(gid, site_id, sid, 
-                biscape_Fe2O3,biscape_Cu, biscape_Zn, lon, lat)
-
-### --- rename some variables --- ###
-no2_measures <- no2_measures %>% 
-  dplyr::rename(no2_measures = NO2.DirectMeasurements)
-
-pm25_measures <- pm25_measures %>% 
-  dplyr::rename(pm25_measures = c_biscape) 
-
-bc_measures <- bc_measures %>% 
-  dplyr::rename(bc_measures = c_biscape)
-
-pm25_constituents_measures <- pm25_constituents_measures %>% 
-  dplyr::rename(fe2o3_measures = biscape_Fe2O3, 
-                cu_measures = biscape_Cu, 
-                zn_measures = biscape_Zn)
-
-#######################################
-### --- Model predictions data --- #################################################################################################
-######################################
-
[…]

### 2023-11-13 manuscript_BiSC.tex

@@
 
+% --- Performance --- %
+The performance of the models varied substantially by pollutant.
+In table x.
+performance metrics for the LUR and HM are shown.
+Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE in all models.
+The performance of the PM25
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{|l|l|l|l|l|l|}
+\hline
+\textbf{Model} & \textbf{Pollutant} & \textbf{Dataset campaign} & \textbf{N} & \textbf{R\textsuperscript{2} (10-CV)} & \textbf{RMSE (10-CV)} \\ \hline
+HM & NO\textsubscript{2} & BiSC-home & 1232 & 0.64 & 7.5 \\ \hline
+HM & PM\textsubscript{2.5} & BiSCAPE & 161 & 0.66 & 3.45 \\ \hline
+HM & BC & BiSCAPE & 74 & 0.86 & ... \\ \hline
+HM & Fe & BiSCAPE & ... & ... & ... \\ \hline
+HM & Cu & BiSCAPE & ... & ... & ... \\ \hline
+HM & Zn & BiSCAPE & ... & ... & ... \\ \hline
+LUR & NO\textsubscript{2} & BiSC-home & ... & ... & ... \\ \hline
+LUR & PM\textsubscript{2.5} & BiSCAPE & ... & ... & ... \\ \hline
+LUR & BC & BiSCAPE & ... & ... & ... \\ \hline
+LUR & Fe & BiSCAPE & ... & ... & ... \\ \hline
+LUR & Cu & BiSCAPE & ... & ... & ... \\ \hline
+LUR & Zn & BiSCAPE & ... & ... & ... \\ \hline
+\end{tabular}
+\caption{Your caption here}
+\label{table:your_label_here}
+\end{table}
+
+% --- Exposure concentrations --- %
+Outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model are shown in FIGURE X. The exposure range was larger for the LUR model
+
+% --- Correlation between measures and model estimates --- %
+Spearman correlation of air pollution exposure during the entire pregnancy between measurements, LUR, DM and HM at the residential addresses ranged from r = x to x (figure X).
+
 \textbf{\cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5} and BC concentrations from the LUR, DM, and HM.

### 2023-11-13 02.Code/01.-descriptive_AP.R
changes

@@ -78,8 +78,6 @@ bc_hybrid <- bc_hybrid %>%
              dplyr::rename(bc_hm = bc.total_hybridmodel)
 
 
-
-
 ##################################
 ### --- Dispersion models --- ###
 #################################

### 2023-11-13 02.Code/01.-descriptive_AP.R
change

@@ -67,7 +67,6 @@ dplyr::glimpse(bc_hybrid)
 dplyr::glimpse(pm25_constituents_hybrid)
 
 # --- rename variables --- # 
-
 no2_hybrid <- no2_hybrid %>% 
               dplyr::rename(no2_hm = no2.total_hybridmodel)
 
@@ -77,13 +76,23 @@ pm25_hybrid <- pm25_hybrid %>%
 bc_hybrid <- bc_hybrid %>%  
              dplyr::rename(bc_hm = bc.total_hybridmodel)
 
+pm25_constituents_hybrid <- pm25_constituents_hybrid %>% 
+                            dplyr::rename(fe_hm = fe.total_hybridmodel,
+                                          cu_hm = cu.total_hybridmodel,
+                                          zn_hm = zn.total_hybridmodel)
+
+# --- long-format --- # 
+
+
+
+
+
 
 ##################################
 ### --- Dispersion models --- ###
 #################################
 
-
-
+# Here list of files and process all the data 
 
 
 
@@ -100,23 +109,3 @@ bc_hybrid <- bc_hybrid %>%
 
 
 
-#######################################
-### --- Model predictions data --- ###
-######################################
-
-
-
-
-
-
-
-
-
-
-#####################################################################################
-
-# have a quick look to the data 
-no2_measures %>% dplyr::glimpse()
-pm25_measures %>% dplyr::glimpse()
-bc_measures %>% dplyr::glimpse()
-pm25_constituents_measures %>% dplyr::glimpse()

### 2023-11-13 manuscript_BiSC.tex

@@
 \usepackage{cleveref}
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
 \centering
-\begin{tabular}{|l|l|l|l|l|l|}
-\hline
-\textbf{Model} & \textbf{Pollutant} & \textbf{Dataset campaign} & \textbf{N} & \textbf{R\textsuperscript{2} (10-CV)} & \textbf{RMSE (10-CV)} \\ \hline
-HM & NO\textsubscript{2} & BiSC-home & 1232 & 0.64 & 7.5 \\ \hline
-HM & PM\textsubscript{2.5} & BiSCAPE & 161 & 0.66 & 3.45 \\ \hline
-HM & BC & BiSCAPE & 74 & 0.86 & ... \\ \hline
-HM & Fe & BiSCAPE & ... & ... & ... \\ \hline
-HM & Cu & BiSCAPE & ... & ... & ... \\ \hline
-HM & Zn & BiSCAPE & ... & ... & ... \\ \hline
-LUR & NO\textsubscript{2} & BiSC-home & ... & ... & ... \\ \hline
-LUR & PM\textsubscript{2.5} & BiSCAPE & ... & ... & ... \\ \hline
-LUR & BC & BiSCAPE & ... & ... & ... \\ \hline
-LUR & Fe & BiSCAPE & ... & ... & ... \\ \hline
-LUR & Cu & BiSCAPE & ... & ... & ... \\ \hline
-LUR & Zn & BiSCAPE & ... & ... & ... \\ \hline
+\begin{tabular}{llllll}
+\toprule
+\textbf{Model} & \textbf{Pollutant} & \textbf{Dataset campaign} & \textbf{N} & \textbf{R2 (10-CV)} & \textbf{RMSE (10-CV)} \\
+\midrule
+\multirow{6}{*}{LUR} & NO2 & BiSC-home & ... & ... & ... \\
+& PM2.5 & BiSCAPE & ... & ... & ... \\
+& BC & BiSCAPE & ... & ... & ... \\
+& Fe & BiSCAPE & ... & ... & ... \\
+& Cu & BiSCAPE & ... & ... & ... \\
+& Zn & BiSCAPE & ... & ... & ... \\
+\midrule
+\multirow{6}{*}{HM} & NO2 & BiSC-home & 1232 & 0.64 & 7.5 \\
+& PM2.5 & BiSCAPE & ... & ... & ... \\
+& BC & BiSCAPE & ... & ... & ... \\
+& Fe & BiSCAPE & ... & ... & ... \\
+& Cu & BiSCAPE & ... & ... & ... \\
+& Zn & BiSCAPE & ... & ... & ... \\
+\bottomrule
 \end{tabular}
-\caption{Your caption here}
-\label{table:your_label_here}
+\caption{Descriptive caption for the table.}
+\label{tab:my_label}
 \end{table}

### 2023-11-13 02.Code/01.-descriptive_AP.R
checking dispersion estimates

@@ -1,5 +1,6 @@
 # Install packages
-pacman::p_load(tidyverse)
+pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate,  mice, ggmice, plyr, data.table)
+
 
 ##############################
 ### --- Load the data --- ###
@@ -85,14 +86,18 @@ pm25_constituents_hybrid <- pm25_constituents_hybrid %>%
 
 
 
-
-
-
 ##################################
 ### --- Dispersion models --- ###
 #################################
+weekly_home_dm <- read.csv("01.Data/BiSC_dispersion_models/imputed/weekly_estimate_DM_home_imputed.csv")
 
-# Here list of files and process all the data 
+# quick check of the data
+dplyr::glimpse(weekly_home_dm)
+skimr::skim(weekly_home_dm)
+
+# dates in proper format
+weekly_home_dm$date_start <- as.Date(weekly_home_dm$date_start)
+weekly_home_dm$date_end <- as.Date(weekly_home_dm$date_end)
 
 
 

### 2023-11-13 02.Code/01.-descriptive_AP.R
last update

@@ -14,7 +14,6 @@ pm25_measures <- read.csv("01.Data/BiSC_comparison_data/measures/pm25_final_mode
 bc_measures <- read.csv("01.Data/BiSC_comparison_data/measures/bc_final_model_data.csv")
 pm25_constituents_measures <- read.csv("01.Data/BiSC_comparison_data/measures/pm25_constituents_model_data.csv")
 
-
 # Have a quick look to the data 
 dplyr::glimpse(no2_measures)
 dplyr::glimpse(pm25_measures)
@@ -23,17 +22,17 @@ dplyr::glimpse(pm25_constituents_measures)
 
 ### --- selection of specific variables --- ###
 no2_measures <- no2_measures %>%
-  dplyr::select(gid, id, year, week, NO2.DirectMeasurements, lon, lat) 
+                dplyr::select(gid, id, year, week, NO2.DirectMeasurements, lon, lat) 
 
 pm25_measures <- pm25_measures %>% 
-  dplyr::select(gid, site_id, sid, c_biscape, lon, lat)
+                 dplyr::select(gid, site_id, sid, c_biscape, lon, lat)
 
 bc_measures <- bc_measures %>% 
-  dplyr::select(gid, site_id, sid, c_biscape, lon, lat)
+               dplyr::select(gid, site_id, sid, c_biscape, lon, lat)
 
 pm25_constituents_measures <- pm25_constituents_measures %>% 
-  dplyr::select(gid, site_id, sid, 
-                biscape_Fe2O3,biscape_Cu, biscape_Zn, lon, lat)
+                              dplyr::select(gid, site_id, sid, 
+                                            biscape_Fe2O3,biscape_Cu, biscape_Zn, lon, lat)
 
 ### --- rename some variables --- ###
 no2_measures <- no2_measures %>% 
@@ -82,35 +81,63 @@ pm25_constituents_hybrid <- pm25_constituents_hybrid %>%
                                           cu_hm = cu.total_hybridmodel,
                                           zn_hm = zn.total_hybridmodel)
 
-# --- long-format --- # 
-
-
 
 ##################################
 ### --- Dispersion models --- ###
 #################################
[…]

### 2023-11-14 manuscript_BiSC_outdoor_AP.tex
changes in name of the file and other stuff

--- 2023-11-13 manuscript_BiSC.tex
@@
 
-In this study, within the framework of the Barcelona Life Study Cohort (BiSC), we developed and compared LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses.
+In this study, within the framework of the Barcelona Life Study Cohort (BiSC), we developed LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses.
 LUR and DM, were constructed using conventional state-of-the-art techniques, while HM was innovatively designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
@@
 % --- Performance --- %
+Since models were built based on different approaches there where not directly comparable due to differences on a temporal basis
+
+Since different approaches were to built outdoor air pollution models were not directly comparable due to differences on temporal
+
 The performance of the models varied substantially by pollutant.
 In table x.
-performance metrics for the LUR and HM are shown.
-Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE in all models.
+performance metrics for the HM are shown.
+Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE.
 The performance of the PM25
+
+Performance metrics for LUR and DM are shown in the supplementary materials.
 
@@
 \toprule
-\textbf{Model} & \textbf{Pollutant} & \textbf{Dataset campaign} & \textbf{N} & \textbf{R2 (10-CV)} & \textbf{RMSE (10-CV)} \\
+\textbf{Model} & \textbf{Pollutant} & \textbf{Dataset campaign} & \textbf{N} & \textbf{R\textsuperscript{2} (10-CV)} & \textbf{RMSE (10-CV)} \\
 \midrule
-\multirow{6}{*}{LUR} & NO2 & BiSC-home & ... & ... & ... \\
-& PM2.5 & BiSCAPE & ... & ... & ... \\
-& BC & BiSCAPE & ... & ... & ... \\
-& Fe & BiSCAPE & ... & ... & ... \\
-& Cu & BiSCAPE & ... & ... & ... \\
-& Zn & BiSCAPE & ... & ... & ... \\
-\midrule
-\multirow{6}{*}{HM} & NO2 & BiSC-home & 1232 & 0.64 & 7.5 \\
-& PM2.5 & BiSCAPE & ... & ... & ... \\
-& BC & BiSCAPE & ... & ... & ... \\
+\multirow{6}{*}{HM} & NO\textsubscript{2} & BiSC-home & 1232 & 0.64 & 7.48 \\
+& PM\textsubscript{2.5} & BiSCAPE & 161 & 0.66 & 3.45 \\
+& BC & BiSCAPE & 74 & 0.86 &  0.23 \\
 & Fe & BiSCAPE & ... & ... & ... \\

### 2023-11-14 supplementary_BiSC_outdoor_AP.tex
changes in name of the file and other stuff

--- 2023-11-13 manuscript_BiSC.tex
@@
 \usepackage{cleveref}
-\usepackage{booktabs}
-\usepackage{multirow}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}
+\title{Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Abstract --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
+% Index table with other stuff
 
-Air pollution is the leading environmental risk factor for human health worldwide.
-A detailed and comprehensive exposure assessment of outdoor air pollution considering both spatial and temporal variability is fundamental to conduct epidemiological studies.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Introduction --- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-
-Air pollution stands as the primary environmental risk factor for human health and the fourth most deadly health risk worldwide, playing a significant role in the global burden of disease \cite{cohen2017, he2020state}.
-Exposure to ambient air pollution has been linked with a extensive array of adverse health outcomes \cite{boogaard2022, guxens2022hei, haddad2023}.
-In this context urbanization have been playing a major role increasing air pollution exposure as cities grow and more population moves to urban areas \cite{nieuwenhuijsen2016}.
-This trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas \cite{boogaard2022}.
-
-Over the past few decades, several modelling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution \cite{hoek2017, di2019no2, di2019pm25, stafoggia2019, stafoggia2020}.
-Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches for estimating air pollution exposure concentrations in epidemiological studies \cite{gulliver2015, dehoogh2014}.
-LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that work as proxies for traffic-related air pollution \cite{briggs1997, hoek2008}.
-In contrast, DMs rely on mathematical simulations to estimate air pollution levels.
-These predictions are based on data from emissions sources, chemical transformation models, pollutant dispersion models, topography, roughness, and meteorological data \cite{hoek2017}.
-Although these methods have been extensively used they come with inherent limitations related to lack of temporal and spatial variation.
-To overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}.
-
-While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains, such models (multi-country and national domain models) often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
-This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potential lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.
-
-Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}.
-In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees-based models \cite{liu2022}.
-This family of models is often used due to its ability to model potential, nonlinear interaction between predictor variables \cite{liu2022treebased}.
-Among the different ML algorithms, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.
-
-Furthermore, the COVID-19 pandemic has introduced unprecedented challenges for air pollution modelling and exposure assessment.
-These challenges encompass changes in air pollution patterns, constraints on data collection due to health and safety restrictions (lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}.
-
-In this study, within the framework of the Barcelona Life Study Cohort (BiSC), we developed and compared LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses.
-LUR and DM, were constructed using conventional state-of-the-art techniques, while HM was innovatively designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
-The overall aim is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates, intended to be used in the prenatal exposure assessment within the BiSC cohort.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Material and methods--- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and methods}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Study domain and Monitoring Campaigns --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Study domain and monitoring campaigns}
-
-Barcelona (\textbf{\cref{fig1}}), is one of the most populated and dense cities in Spain, with 1,656,725 inhabitants ($\approx16,255.4$ inhabitants/km\textsuperscript{2}).
-It is located on the Mediterranean coast of the Iberian peninsula, protected by the Collserola mountain range and delimited by two delta rivers (Besòs and LLobregat).
-The city has a Mediterranean climate, the wind patterns are dominated by a breeze that blows from the sea during the daytime and from the land during night-time.
-Barcelona is one of the cities in Europe with the highest vehicle traffic density (6,000 vehicles/km\textsuperscript{2}) \cite{casallas2018}.
-
-% Put the figure text closer to the Figure 1
-\captionsetup[figure]{skip=-4pt}
-% We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/fig1_op2.png}
-\caption{Study domain and spatial distribution (BiSC participants, BiSCAPE campaign, monitoring stations, and traffic points)}
-\label{fig1}
-\end{figure}
-
-Air pollution measurements were conducted during two separate monitoring campaigns.
-The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021.
-The second campaign was carried out in accordance with the ESCAPE guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
-During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and hybrid models.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- LUR model --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{LUR models}
-
-LUR models for NO\textsubscript{2}  were developed using only data from the locations that had measured home-outdoor (BiSC-home campaign) levels for both sampling campaigns (i.e., we did not use data from participants who had data available for only one campaign, either first or third trimester).
-Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites were during the pandemic period and its resulting restrictions.
-For PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC models were developed using BiSCAPE data from the 34 sites.
-
-For the development of the LUR models, spatial variables that better represent the pollutant concentration distribution across the study area were created.
-Following the ESCAPE guidelines \cite{eeftens2012, beelen2013},  predictor variables were selected from three broad categories: traffic, land use, and urban configuration.
-A wide variety of buffer sizes were created, i.e, 25m, 50m, 100m, 300m, 500m, and 1000m to capture the area-specific predictor variable at every single location.
-
-We applied a supervised stepwise process in which we sequentially added a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model.
-This procedure was followed to build all the specific pollutant models.
-Previously to model development, we specified the direction expected of the effect that each predictor has in the model.
-For all predictor variables, univariate regression models with annual average air pollution concentration (NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC) were conducted, models with the highest adjusted explained variance (R\textsuperscript{2}) were selected.
-We then repeated the process using the previous model and all the remaining predictor variables.
-Details of the LUR modelling and variables used can be found in the \textbf{Supplementary material X}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Dispersion model --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Dispersion models}
-
-An emission inventory was compiled for the period 2018-2021, which included data from diverse emission sources such as vehicular traffic, household activities, industrial plants, airports, harbors, factories, and the commercial sector.
-The Atmospheric Dispersion Modelling System software (ADMS-Urban) from Cambridge Environmental Research Consultants (CERC) was employed to calculate the concentration of pollutants like NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC originating from these sources.
-ADMS-Urban is a specialized Gaussian model designed for analyzing pollutant emissions in urban and metropolitan areas, particularly at road level \cite{mchugh1997adms}.
-
-ADMS-Urban employs various algorithms to assess the chemical transportation and dispersion of pollutants, factoring in ground-level turbulence and turbulence induced by surrounding buildings \cite{stocker2012adms}.
-The model also features a meteorological pre-processing system that calculates fluxes and turbulence, enabling it to model airflow over complex terrains.
-
-For enhanced accuracy, the dispersion model integrates data from a three-dimensional building and street width model from the Institut Cartogràfic i Geològic de Catalunya \cite{stoter2016state, 3Dcity}.
-This data aids in the calculation of the street canyon effect, a phenomenon characterized by tall buildings situated adjacent to heavily trafficked roads, which hinders atmospheric dispersion, leading to higher and relatively consistent pollutant concentrations compared to areas unaffected by the canyon effect.
-
-The model also utilized meteorological data including wind speed, direction, air temperature, cloud cover, solar radiation, relative humidity, precipitation rate, and Planetary Boundary Layer Height (PBLH).
-This meteorological data was collected from the Zona Universitària station in Barcelona, managed by the Servei Meteorològic de Catalunya \cite{xema2013}.
-
-To incorporate background pollution levels at a regional scale, data from Copernicus Atmosphere Monitoring Service Information (CAMS) was used.
-Background air pollution data has been extracted from the CAMS reanalysis ensemble model for 2018, and from CAMS ensemble forecast model for 2019-2020 \cite{cams2020, franceinstitut}.
-
-In addition, bias-adjustment techniques have been used to fit the model data to the observations.
-Through hourly bias spatial interpolation and generation of hourly maps.
-After that, data were processed  by cluster analysis techniques to obtain a mean cluster map.
-Each bias map by pollutant was associated to a cluster map class and an hourly-varying profile.
-
-Further details regarding the ADMS-Urban model and the variables used can be found in \textbf{Supplementary materials X}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Hybrid model --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hybrid models}
-
-For the hybrid models, we sought to improve their predictive power by integrating comprehensive data from multiple information sources and leveraging ML.
-For NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC were largely based on daily concentrations from dispersion models, summarized as weekly averages, and spatial predictor variables from the LUR models.
-
-For a more comprehensive approach, we integrated additional data from the Catalan Air Pollution Monitoring and Forecasting Network (XVPCA) for Barcelona and surrounding areas \cite{xarxa2012}.
-This data was interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
-The data came from 16 monitoring stations, including Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St.
-Adria, St.
-Coloma, Vall Hebron, Zona Universitária \cite{xarxa2012}.
-
-While our dispersion model initially utilized meteorological and traffic data, the hybrid model broadened its scope by integrating daily traffic counts from 180 locations throughout the city \cite{trafficbcn}.
-This broader data helps us better understand the spatial and time-specific variations in urban air pollution \cite{pinto2020}.
-To further refine our insights into the influence of weather patterns on pollution, we sourced additional meteorological variables, such as wind speed and air temperature, from the Raval monitoring station managed by Servei Meteorològic de Catalunya \cite{xema2013}.
-
-We employed the RF algorithm to capture non-linear relationships and potential interactions between predictor variables and the response variable, utilizing \textit{the ranger package} in R implemented in the \textit{caret} framework \cite{wright2019}, a set of functions that attempt to streamline the process for creating predictive models \cite{caret2008}.
-Briefly, RF is a machine learning (ML) algorithm that constructs an ensemble of multiple decision trees \cite{breiman2001}.
-In this approach, each tree is formulated using a bootstrap sample of the data.
-During this process, every node within a tree represents the optimal split from a randomly selected subset of predictors.
-Predictions are then derived by averaging the outputs from all trees.
-
-In summary, by combining data from multiple sources, our hybrid models aim to provide more accurate and reliable predictions.
-These predictions are essential for studying and mitigating urban air pollution.
-The distribution of the air pollution monitoring stations, meteorological stations, and traffic count measures are shown in the (\textbf{\cref{fig1}}, \textbf{Panel D}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Model Evaluation and Comparison--- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Performance assessment}
-
-We evaluated model performance using various statistical approaches and on different temporal bases.
-The LUR models were evaluated on an annual basis, dispersion models were assessed both hourly and daily, and the hybrid models were evaluated on a weekly basis.
-
-\subsubsection{LUR model evaluation}
-Performance of the LUR models was assessed using \textit{k-fold cross-validation}.
-Specifically, we employed a 10-fold cross-validation technique.
-In this approach, the data was split into ten distinct sub-samples.
-For each iteration, the model was trained on k-1 folds and then validated using the leftover fold \cite{ziegel2003}.
-For every fold, we computed metrics such as \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)).
-After each iteration, both the error and adjusted \(R\textsuperscript{2}\) were recorded.
-This process was repeated \(k-times\), and the average of these measures was reported as the indicator of the model validity.
-\vspace{0.5cm}
-
-Details from the metrics used for model validation can be found in the \textbf{Supplementary material X}.
-
-These metrics are defined as follows:
-% --- R2 --- %
-\begin{equation}
-R^2 = \left( \frac{\sum (\hat{y}_i - \bar{y})^2}{\sum (y_i - \bar{y})^2} \right)
-\end{equation}
-
-% --- root mean squared error (RMSE) --- % 
-\begin{equation}
-\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)^2}
-\end{equation}
-
-% --- mean bias (MB) --- % 
-\begin{equation}
-\text{Mean Bias} = \frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)
-\end{equation}
-
-% --- correlation coefficient (r) --- %
-\begin{equation}
-r = \frac{\sum (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum (x_i - \bar{x})^2 \sum (y_i - \bar{y})^2}}
-\end{equation}
-\vspace{0.5 cm}
-
-We carried out both internal and external validation for the NO\textsubscript{2} LUR model.
-Internal validation was conducted by applying a K-fold cross-validation analysis.
-The 10-fold cross-validation procedure and its significance have been detailed above.
-For the external validation of the NO\textsubscript{2} LUR model, we applied data from BiSC-home participants' homes for those participants who had data available only for one measurement campaign, and therefore their data were not used for developing NO\textsubscript{2} LUR model.
-For LUR models BC, PM\textsubscript{2.5}, and its constituents (Cu, Fe, Sb, and Zn), internal validation was conducted using the \textit{leave-one-out-cross-validation (LOOCV)} process.
-For these models, we were not able to conduct external validation due to the lack of external data.
-
-\subsection{Dispersion model evaluation}
-Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
-Details from the metrics used for dispersion model validation can be found in the \textbf{Supplementary material X}.\vspace{0.5cm}
-
-The metrics used for the dispersion models we:
-
-% --- Percentage of observed-modelled pairs (n) --- %
-\begin{equation}
-n = \frac{N_{\text{pairs used}}}{N_{\text{total pairs}}} \times 100
-\end{equation}
-
-% --- Factor of Two (FAC2) --- % 
-\begin{equation}
-\text{FAC2} = \frac{1}{n} \sum_{i=1}^{n} U,
-\end{equation}
-%Where \( U \) is 1 if \( 0.5 \leq \frac{M_i}{O_i} \leq 2 \) and 0 otherwise.
-
-% --- Mean Bias (MB) --- % 
-\begin{equation}
-\text{MB} = \frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)
-\end{equation}
-
-% --- Mean Gross Error (MGE) --- % 
-\begin{equation}
-\text{MGE} = \frac{1}{n} \sum_{i=1}^{n} |O_i - M_i|
-\end{equation}
-
-% --- Normalized Mean Bias (NMB) --- % 
-\begin{equation}
-\text{NMB} = \frac{\sum_{i=1}^{n} (M_i - O_i)}{\sum_{i=1}^{n} O_i}
-\end{equation}
-
-% --- Root Mean Square Error (RMSE) --- % 
-\begin{equation}
-\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)^2}
-\end{equation}
-
-% --- Correlation Coefficient (r) --- % 
-\begin{equation}
-r = \frac{\sum_{i=1}^{n} (O_i - \bar{O})(M_i - \bar{M})}{\sqrt{\sum_{i=1}^{n} (O_i - \bar{O})^2 \sum_{i=1}^{n} (M_i - \bar{M})^2}}
-\end{equation}
-\vspace{0.5cm}
-
-\subsection{Hybrid model evaluation}
-We assessed the performance of each hybrid model by contrasting predictions with observations.
-Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and the \textit{out-of-bag (OBB) score}.
-While both methods vary in the sample size designated for training and testing, their core difference lies in the data-splitting strategy.
-In the 10-fold CV, the model trains on nine out of the ten data partitions and then validates against the remaining one.
-In contrast, the OOB approach trains the model on two-thirds of the data and uses the residual third for external validation.
-We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}
-
-All statistical analyses were performed in R version 4.3.1 \cite{Rstudio}. GIS predictor variables were computed using X software.
-
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Results --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-% --- Performance --- %
-The performance of the models varied substantially by pollutant.
-In table x.
-performance metrics for the LUR and HM are shown.
-Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE in all models.
-The performance of the PM25
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{llllll}
-\toprule
-\textbf{Model} & \textbf{Pollutant} & \textbf{Dataset campaign} & \textbf{N} & \textbf{R2 (10-CV)} & \textbf{RMSE (10-CV)} \\
-\midrule
-\multirow{6}{*}{LUR} & NO2 & BiSC-home & ... & ... & ... \\
-& PM2.5 & BiSCAPE & ... & ... & ... \\
-& BC & BiSCAPE & ... & ... & ... \\
-& Fe & BiSCAPE & ... & ... & ... \\
-& Cu & BiSCAPE & ... & ... & ... \\
-& Zn & BiSCAPE & ... & ... & ... \\
-\midrule
-\multirow{6}{*}{HM} & NO2 & BiSC-home & 1232 & 0.64 & 7.5 \\
-& PM2.5 & BiSCAPE & ... & ... & ... \\
-& BC & BiSCAPE & ... & ... & ... \\
-& Fe & BiSCAPE & ... & ... & ... \\
-& Cu & BiSCAPE & ... & ... & ... \\
-& Zn & BiSCAPE & ... & ... & ... \\
-\bottomrule
-\end{tabular}
-\caption{Descriptive caption for the table.}
-\label{tab:my_label}
-\end{table}
-
-% --- Exposure concentrations --- %
-Outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model are shown in FIGURE X. The exposure range was larger for the LUR model
-
-% --- Correlation between measures and model estimates --- %
-Spearman correlation of air pollution exposure during the entire pregnancy between measurements, LUR, DM and HM at the residential addresses ranged from r = x to x (figure X).
-
-\textbf{\cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5} and BC concentrations from the LUR, DM, and HM.
-The data shown are long-term estimated concentrations of pollutants, after averaging weekly estimates for all the study period (2019 - 2021).
-
-%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
-% Put the figure text closer to the Figure 3
-\captionsetup[figure]{skip=6pt}
-% We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/fig_HM_test_all_models.png}
-\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
-\label{fig2}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Discussion --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-We successfully developed spatiotemporal models using three different approaches (LUR, dispersion and hybrid models).
-All models had moderate to good performance evaluated by the defined metrics in the \textbf{methods section}.
-
-As we expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed.
-We believe that by combining multiple data sources, and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.
-
-\subsection{Comparison between LUR, DM, and Hybrid model}
-
-\subsection{Comparison with Previous studies}
-
-\subsection{Strengths and Limitations}
-One of the main strengths of this study is the extensive exposure assessment
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Conclusions --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-Air pollution exposure concentration estimates from LUR, dispersion and hybrid models were moderately to strongly correlated. We observed
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2023-11-14 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 This process was repeated \(k-times\), and the average of these measures was reported as the indicator of the model validity.
-\vspace{0.5cm}
-
-Details from the metrics used for model validation can be found in the \textbf{Supplementary material X}.
-
-These metrics are defined as follows:
-% --- R2 --- %
-\begin{equation}
-R^2 = \left( \frac{\sum (\hat{y}_i - \bar{y})^2}{\sum (y_i - \bar{y})^2} \right)
-\end{equation}
-
-% --- root mean squared error (RMSE) --- % 
-\begin{equation}
-\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)^2}
-\end{equation}
-
-% --- mean bias (MB) --- % 
-\begin{equation}
-\text{Mean Bias} = \frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)
-\end{equation}
-
-% --- correlation coefficient (r) --- %
-\begin{equation}
-r = \frac{\sum (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum (x_i - \bar{x})^2 \sum (y_i - \bar{y})^2}}
-\end{equation}
-\vspace{0.5 cm}
+\vspace{0.5cm} Details from the metrics used for model validation can be found in the \textbf{Supplementary material X}.
 
@@
 Details from the metrics used for dispersion model validation can be found in the \textbf{Supplementary material X}.\vspace{0.5cm}
-
-The metrics used for the dispersion models we:
-
-% --- Percentage of observed-modelled pairs (n) --- %
-\begin{equation}
-n = \frac{N_{\text{pairs used}}}{N_{\text{total pairs}}} \times 100
-\end{equation}
-
-% --- Factor of Two (FAC2) --- % 
-\begin{equation}
-\text{FAC2} = \frac{1}{n} \sum_{i=1}^{n} U,
-\end{equation}
-%Where \( U \) is 1 if \( 0.5 \leq \frac{M_i}{O_i} \leq 2 \) and 0 otherwise.
-
-% --- Mean Bias (MB) --- % 
-\begin{equation}
-\text{MB} = \frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)
-\end{equation}
-
-% --- Mean Gross Error (MGE) --- % 
-\begin{equation}
-\text{MGE} = \frac{1}{n} \sum_{i=1}^{n} |O_i - M_i|
-\end{equation}
-
-% --- Normalized Mean Bias (NMB) --- % 
-\begin{equation}
-\text{NMB} = \frac{\sum_{i=1}^{n} (M_i - O_i)}{\sum_{i=1}^{n} O_i}
-\end{equation}
-
-% --- Root Mean Square Error (RMSE) --- % 
-\begin{equation}
-\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)^2}
-\end{equation}
-
-% --- Correlation Coefficient (r) --- % 
-\begin{equation}
-r = \frac{\sum_{i=1}^{n} (O_i - \bar{O})(M_i - \bar{M})}{\sqrt{\sum_{i=1}^{n} (O_i - \bar{O})^2 \sum_{i=1}^{n} (M_i - \bar{M})^2}}
-\end{equation}
-\vspace{0.5cm}
 

### 2023-11-14 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}
+\title{\textbf{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}
 
@@
 % --- Exposure concentrations --- %
-Outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model are shown in FIGURE X. The exposure range was larger for the LUR model
+Outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in FIGURE X. The exposure range was larger for the LUR model
 
@@
 We successfully developed spatiotemporal models using three different approaches (LUR, dispersion and hybrid models).
-All models had moderate to good performance evaluated by the defined metrics in the \textbf{methods section}.
+All models had moderate to good performance evaluated by the defined metrics in the \textbf{methods section} and displayed in more detail in the \textbf{Supplementary materials}
 

### 2023-11-15 02.Code/01.-descriptive_AP.R
renaming variables

@@ -137,7 +137,18 @@ no2_lur <- no2_lur %>%
            dplyr::rename(no2_lur = no2_week_ratio)
 
 pm25_lur <- pm25_lur %>% 
-            dplyr::select(subject_id, weeks, date_start, date_end)
+            dplyr::select(subject_id, weeks, date_start, date_end) %>% 
+            dplyr::rename(pm25_lur  = pm25_week_lur_ratio)
 
 bc_lur <- bc_lur %>% 
           dplyr::select(subject_id, weeks, date_start, date_end)
+
+fe_lur <- fe_lur %>% 
+          dplyr::select(subject_id, weeks, date_start, date_end)
+
+
+
+
+
+
+

### 2023-11-15 02.Code/01.-descriptive_AP.R
renaming variables

@@ -137,11 +137,12 @@ no2_lur <- no2_lur %>%
            dplyr::rename(no2_lur = no2_week_ratio)
 
 pm25_lur <- pm25_lur %>% 
-            dplyr::select(subject_id, weeks, date_start, date_end) %>% 
-            dplyr::rename(pm25_lur  = pm25_week_lur_ratio)
+            dplyr::select(subject_id, weeks, date_start, date_end, pm25_week_lur_ratio) %>% 
+            dplyr::rename(pm25_lur= pm25_week_lur_ratio)
 
 bc_lur <- bc_lur %>% 
-          dplyr::select(subject_id, weeks, date_start, date_end)
+          dplyr::select(subject_id, weeks, date_start, date_end, bc_week_ratio) %>% 
+          dplyr::rename(bc_lur = bc_week_ratio)
 
 fe_lur <- fe_lur %>% 
           dplyr::select(subject_id, weeks, date_start, date_end)

### 2023-11-15 supplementary_BiSC_outdoor_AP.tex

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}
+\title{\textbf{Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}
 
@@
 %%% --- Authors --- %%%
-\author[1, 2]{Alan Domínguez}
+\author[1, 2, 3]{Alan Domínguez}
 \author[1, 3, 4]{Payam Dadvand}
@@
 
-% Index table with other stuff
+% Index table with to put order to supplementary materials
 
-\section{Introduction}
+\section{Predictor variables}
+\section{}
+
+\section{Model evaluation metrics}
+
+The following metrics were used to evaluate the LUR model for all the air pollution assessed.
+
+% --- R2 --- %
+\begin{equation}
+R^2 = \left( \frac{\sum (\hat{y}_i - \bar{y})^2}{\sum (y_i - \bar{y})^2} \right)
+\end{equation}
+
+% --- root mean squared error (RMSE) --- % 
+\begin{equation}
+\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)^2}
+\end{equation}
+
+% --- mean bias (MB) --- % 
+\begin{equation}
+\text{Mean Bias} = \frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)
+\end{equation}
+
+% --- correlation coefficient (r) --- %
+\begin{equation}
+r = \frac{\sum (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum (x_i - \bar{x})^2 \sum (y_i - \bar{y})^2}}
+\end{equation}
+\vspace{0.5 cm}
+
+For the DM models we evaluate their performance using a different set of metrics as following:
+
+% --- Percentage of observed-modelled pairs (n) --- %
+\begin{equation}
+n = \frac{N_{\text{pairs used}}}{N_{\text{total pairs}}} \times 100
+\end{equation}
+
+% --- Factor of Two (FAC2) --- % 
+\begin{equation}
+\text{FAC2} = \frac{1}{n} \sum_{i=1}^{n} U,
+\end{equation}
+%Where \( U \) is 1 if \( 0.5 \leq \frac{M_i}{O_i} \leq 2 \) and 0 otherwise.
+
+% --- Mean Bias (MB) --- % 
+\begin{equation}
+\text{MB} = \frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)
+\end{equation}
+
+% --- Mean Gross Error (MGE) --- % 
+\begin{equation}
+\text{MGE} = \frac{1}{n} \sum_{i=1}^{n} |O_i - M_i|
+\end{equation}
+
+% --- Normalized Mean Bias (NMB) --- % 
+\begin{equation}
+\text{NMB} = \frac{\sum_{i=1}^{n} (M_i - O_i)}{\sum_{i=1}^{n} O_i}
+\end{equation}
+
+% --- Root Mean Square Error (RMSE) --- % 
+\begin{equation}
+\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)^2}
+\end{equation}
+
+% --- Correlation Coefficient (r) --- % 
+\begin{equation}
+r = \frac{\sum_{i=1}^{n} (O_i - \bar{O})(M_i - \bar{M})}{\sqrt{\sum_{i=1}^{n} (O_i - \bar{O})^2 \sum_{i=1}^{n} (M_i - \bar{M})^2}}
+\end{equation}
+\vspace{0.5cm}
 

### 2023-11-16 02.Code/01.-descriptive_AP.R
renaming lur estimates

@@ -1,10 +1,15 @@
 # Install packages
 pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate,  mice, ggmice, plyr, data.table)
 
-
 ##############################
 ### --- Load the data --- ###
 ############################
+
+###############################
+### --- Home addresses --- ###
+#############################
+bisc_home_coords <- read.csv() # wee need assign to dm and lur data the lon/lat to then plot the predictions (entire pregnancy)
+
 
 #############################################
 ### --- Air pollutants measurements data --- #####################################################################
@@ -115,7 +120,6 @@ dplyr::glimpse(bc_dm)
 ###########################
 ### --- LUR models --- ###
 #########################
-
 no2_lur <- readRDS("01.Data/BiSC_comparison_data/predictions/LUR/no2_weekly_preg_29092023.rds")
 pm25_lur <- readRDS("01.Data/BiSC_comparison_data/predictions/LUR/pm25_weekly_preg_29092023.rds")
 bc_lur <- readRDS("01.Data/BiSC_comparison_data/predictions/LUR/bc_weekly_preg_29092023.rds")
@@ -131,7 +135,7 @@ dplyr::glimpse(fe_lur)
 dplyr::glimpse(cu_lur)
 dplyr::glimpse(zn_lur)
 
-# re-arrenge the data 
+# re-arrange the data 
 no2_lur <- no2_lur %>% 
            dplyr::select(subject_id, weeks, date_start, date_end, no2_week_ratio) %>% 
            dplyr::rename(no2_lur = no2_week_ratio)
@@ -145,8 +149,16 @@ bc_lur <- bc_lur %>%
           dplyr::rename(bc_lur = bc_week_ratio)
 
 fe_lur <- fe_lur %>% 
-          dplyr::select(subject_id, weeks, date_start, date_end)
+          dplyr::select(subject_id, weeks, date_start, date_end, fe_week_ratio) %>% 
+          dplyr::rename(fe_lur = fe_week_ratio)
 
+cu_lur <- cu_lur %>% 
+          dplyr::select(subject_id, weeks, date_start, date_end, cu_week_ratio) %>% 
+          dplyr::rename(cu_lur = cu_week_ratio)
+
+zn_lur <- zn_lur %>% 
+          dplyr::select(subject_id, weeks, date_start, date_end, zn_week_ratio) %>% 
+          dplyr::rename(zn_lue = zn_week_ratio)
 
 
 

### 2023-11-17 supplementary_BiSC_outdoor_AP.tex

@@
 %%% --- Authors --- %%%
-\author[1, 2, 3]{Alan Domínguez}
-\author[1, 3, 4]{Payam Dadvand}
-\author[1]{Marta Cirach}
-\author[1]{Bruno Raimbault}
-\author[1]{Toni Galmes}
-\author[1]{Karl Samuelsson}
-\author[1, 2, 3]{Mark Nieuwenhuijsen}
-\author[1, 2, 3]{Jordi Sunyer}
-\author[1, 2, 3]{Xavier Basagaña}
-\author[1, 3]{Ioar Rivas}
-
-%%% --- Affiliations --- %%%
-\affil[1]{Barcelona Institute for Global Health (ISGlobal), Barcelona, Spain.}
-\affil[2]{Universitat Popmpeu Fabra (UPF), Barcelona, Spain.}
-\affil[3]{CIBER Epidemiología y Salud Pública (CIBERESP), Madrid, Spain.}
-\affil[4]{London School of Hygiene and Tropical Medicine ( LSHTM), London, UK.}
-
+\author{Alan Domínguez, Payam Dadvand, Marta Cirach, Bruno Raimbault, Toni Galmes, Karl Samuelsson, Mark Nieuwenhuijsen, Jordi Sunyer, Xavier Basagaña, Ioar Rivas}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Predictor variables}
-\section{}
+\section{Land Use Regression modelling details}
+
+\subsection{Predictor variables}
+\subsection{Performance metrics}
+
+\section{Dispersion modelling details}
+
+\subsection{Predictor variables}
+\subsection{Performance metrics}
 

### 2023-11-20 02.Code/01.-descriptive_AP.R
re-arrenging data

@@ -135,30 +135,136 @@ dplyr::glimpse(fe_lur)
 dplyr::glimpse(cu_lur)
 dplyr::glimpse(zn_lur)
 
-# re-arrange the data 
+##################################
+### --- Re-ordering weeks --- ###
+#################################
+
+# --- NO2 --- #
+no2_lur$weeks <-  dplyr::recode_factor(no2_lur$weeks, 
+                                       w1 = 'w01', w2 = 'w02', w3 = 'w03', w4 = 'w04', w5 = 'w05',
+                                       w6 = 'w06', w7 = 'w07', w8 = 'w08', w9 = 'w09', w10 = 'w10',
+                                       w11 = 'w11', w12 = 'w12', w13 = 'w13', w14 = 'w14', w15 = 'w15',
+                                       w16 = 'w16', w17 = 'w17', w18 = 'w18', w19 = 'w19', w20 = 'w20',
+                                       w21 = 'w21', w22 = 'w22', w23 = 'w23', w24 = 'w24', w25 = 'w25',
+                                       w26 = 'w26', w27 = 'w27', w28 = 'w28', w29 = 'w29', w30 = 'w30',
+                                       w31 = 'w31', w32 = 'w32', w33 = 'w33', w34 = 'w34', w35 = 'w35',
+                                       w36 = 'w36', w37 = 'w37', w38 = 'w38', w39 = 'w39', w40 = 'w40',
+                                       w41 = 'w41', w42 = 'w42', w43 = 'w43', w44 = 'w44')
+
+# --- PM25 --- #  
+pm25_lur$weeks <- dplyr::recode_factor(pm25_lur$weeks, 
+                                       w1 = 'w01', w2 = 'w02', w3 = 'w03', w4 = 'w04', w5 = 'w05',
+                                       w6 = 'w06', w7 = 'w07', w8 = 'w08', w9 = 'w09', w10 = 'w10',
+                                       w11 = 'w11', w12 = 'w12', w13 = 'w13', w14 = 'w14', w15 = 'w15',
+                                       w16 = 'w16', w17 = 'w17', w18 = 'w18', w19 = 'w19', w20 = 'w20',
+                                       w21 = 'w21', w22 = 'w22', w23 = 'w23', w24 = 'w24', w25 = 'w25',
[…]

### 2023-11-20 02.Code/01.-descriptive_AP.R
plotting maps

@@ -1,5 +1,6 @@
 # Install packages
-pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate,  mice, ggmice, plyr, data.table)
+pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate, 
+               mice, ggmice, plyr, data.table, mapview, ggmap, omsdata)
 
 ##############################
 ### --- Load the data --- ###
@@ -8,8 +9,24 @@ pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate,  mice, ggmice,
 ###############################
 ### --- Home addresses --- ###
 #############################
-bisc_home_coords <- read.csv() # wee need assign to dm and lur data the lon/lat to then plot the predictions (entire pregnancy)
+gid_data <- read.csv("01.Data/hybrid_data_final_040823.csv") # wee need assign to dm and lur data the lon/lat to then plot the predictions (entire pregnancy)
+dplyr::glimpse(gid_data)
 
+gid_data <- gid_data %>% dplyr::select(gid, subject_id, weeks, lon, lat)
+
+gid_data$weeks <- dplyr::recode_factor(gid_data$weeks, 
+                                       w1 = 'w01', w2 = 'w02', w3 = 'w03', w4 = 'w04', w5 = 'w05',
+                                       w6 = 'w06', w7 = 'w07', w8 = 'w08', w9 = 'w09', w10 = 'w10',
+                                       w11 = 'w11', w12 = 'w12', w13 = 'w13', w14 = 'w14', w15 = 'w15',
+                                       w16 = 'w16', w17 = 'w17', w18 = 'w18', w19 = 'w19', w20 = 'w20',
+                                       w21 = 'w21', w22 = 'w22', w23 = 'w23', w24 = 'w24', w25 = 'w25',
+                                       w26 = 'w26', w27 = 'w27', w28 = 'w28', w29 = 'w29', w30 = 'w30',
+                                       w31 = 'w31', w32 = 'w32', w33 = 'w33', w34 = 'w34', w35 = 'w35',
+                                       w36 = 'w36', w37 = 'w37', w38 = 'w38', w39 = 'w39', w40 = 'w40',
+                                       w41 = 'w41', w42 = 'w42', w43 = 'w43', w44 = 'w44')
+
+gid_data <- gid_data %>% dplyr::arrange(subject_id, weeks)
+dplyr::glimpse(gid_data)
 
 #############################################
[…]

### 2023-11-21 02.Code/01.-descriptive_AP.R
plotting figures

@@ -1,10 +1,16 @@
 # Install packages
 pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate, 
-               mice, ggmice, plyr, data.table, mapview, ggmap, omsdata)
+               mice, ggmice, plyr, data.table, mapview, ggmap, omsdata, sf, sp, patchwork)
 
 ##############################
 ### --- Load the data --- ###
 ############################
+
+#######################################################
+### --- Load shapefile  Barcelona metropolitan --- ###
+#####################################################
+bcn_shp <- sf::read_sf("01.Data/BCN_GIS_data/bcn/bcn.shp")
+amb_shp <- sf::st_read("01.Data/amb.gpkg")
 
 ###############################
 ### --- Home addresses --- ###
@@ -297,30 +303,148 @@ rio::export(lur_estimates, "01.Data/lur_estimates.csv")
 ##################################
 
 ### -- LUR estimates --- ### 
-lur_estimates_plot <- lur_estimates %>% 
-                      dplyr::group_by(gid) # we have 1.235 home address
+dplyr::glimpse(lur_estimates)
 
+lur_estimates_plot <- lur_estimates %>% 
+                      dplyr::ungroup()
+
+# summarise all the pollutants 
+lur_estimates_plot <- lur_estimates_plot %>% dplyr::group_by(gid) %>%  dplyr::summarise_all(funs(mean))
 dplyr::glimpse(lur_estimates_plot)
+lur_estimates_plot <- lur_estimates_plot %>% dplyr::select(gid, subject_id, no2_lur:lat) %>% tidyr::drop_na()
 
 ### --- Load shapefile  Barcelona metropolitan --- ###b
 bcn_shp <- sf::read_sf("01.Data/BCN_GIS_data/bcn/bcn.shp")
 amb_shp <- sf::st_read("01.Data/amb.gpkg")
 plot(amb_shp[1])
 
+#####################################
+# --- Entire pregnancy LUR --- # 
+###################################
 
-dplyr::glimpse(amb_shp)
-
-
-##################################################
-# --- Here we plot entire pregnancy LUR NO2 --- # 
-##################################################
-
-ggplot() + 
-geom_sf(data = amb_shp, fill = "grey80")+ 
[…]

### 2023-11-21 02.Code/01.-descriptive_AP.R
creating figure

@@ -322,25 +322,31 @@ plot(amb_shp[1])
 # --- Entire pregnancy LUR --- # 
 ###################################
 
-library(sf)
 lur_data <- st_as_sf(lur_estimates_plot, coords = c("lon", "lat"), crs = 32632)  # replace with your actual CRS
-
 
 amb_plot <- ggplot() + 
   geom_sf(data = amb_shp, fill = "grey", color = "white")+ 
   theme(plot.title = element_text(color = "black", size = 14, face = "bold")) +
   facet_grid(. ~ "NO2 LUR model") + theme_bw()
 
+filter_amb_shp <- amb_shp %>% 
+                  dplyr::filter(NOMMUNI %in% c("Badalona", "Barcelona", "Esplugues de Llobregat",
+                                             "l'Hospitalet de Llobregat", "Sant Joan Despí", "Sant Just Desvern",
+                                             "Sant Feliu de Llobregat", "Sant Adrià de Besòs", 
+                                             "Santa Coloma de Gramenet", "Ripollet", "Cerdanyola del Vallès",
+                                             "Sant Cugat del Vallès", "Sant Boi de Llobregat","Cornellà de Llobregat"))
+
+
+amb_plot_filtered <- ggplot() + 
+  geom_sf(data = filter_amb_shp, fill = "grey", color = "white")+ 
+  theme(plot.title = element_text(color = "black", size = 14, face = "bold")) +
+  facet_grid(. ~ "NO2 LUR model") + theme_bw() +
+  xlim(c(st_bbox(filter_amb_shp)[1], st_bbox(filter_amb_shp)[3])) +
+  ylim(c(st_bbox(filter_amb_shp)[2], st_bbox(filter_amb_shp)[4])) 
+
 # Check CRS of shapefile
 print(sf::st_crs(amb_shp))
-
 print(sf::st_crs(lur_estimates_plot))
-
-
-
-library(sf)
-library(dplyr)
-library(ggplot2)
 
 # Assuming lur_estimates_plot is already loaded and as shown in your output
 
@@ -367,7 +373,7 @@ lur_estimates_wgs84 <- st_transform(lur_estimates_sf, 4326)
 
 # Plotting with ggplot2
 
-amb_shp %>% dplyr::select()
+dplyr::glimpse(amb_shp)
 
 
 ggplot() +
@@ -384,56 +390,67 @@ print(st_bbox(amb_shp))
 
 # LUR estimates NO2 
 no2_plot_lur <- ggplot() +
[…]

### 2023-11-21 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 \affil[1]{Barcelona Institute for Global Health (ISGlobal), Barcelona, Spain.}
-\affil[2]{Universitat Popmpeu Fabra (UPF), Barcelona, Spain.}
+\affil[2]{Universitat Pompeu Fabra (UPF), Barcelona, Spain.}
 \affil[3]{CIBER Epidemiología y Salud Pública (CIBERESP), Madrid, Spain.}
@@
 \centering
+\caption{Performance metric for hybrid models.}
 \begin{tabular}{llllll}
@@
 \midrule
-\multirow{6}{*}{HM} & NO\textsubscript{2} & BiSC-home & 1232 & 0.64 & 7.48 \\
+\multirow{6}{*}{HM} & NO\textsubscript{2} & BiSC-hchatome & 1232 & 0.64 & 7.48 \\
 & PM\textsubscript{2.5} & BiSCAPE & 161 & 0.66 & 3.45 \\
 & BC & BiSCAPE & 74 & 0.86 &  0.23 \\
-& Fe & BiSCAPE & ... & ... & ... \\
-& Cu & BiSCAPE & ... & ... & ... \\
-& Zn & BiSCAPE & ... & ... & ... \\
+& Fe & BiSCAPE & ... & 0.54 & ... \\
+& Cu & BiSCAPE & ... & 0.70 & ... \\
+& Zn & BiSCAPE & ... & 0.42 & ... \\
 \bottomrule
 \end{tabular}
-\caption{Descriptive caption for the table.}
 \label{tab:my_label}
@@
 \caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
+\label{fig2}
+\end{figure}
+
+%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
+% Put the figure text closer to the Figure 3
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=1.0\textwidth]{figures/lur_map_plot.png}
+\caption{Pregnancy exposure for the entire pregnancy period \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
 \label{fig2}

### 2023-11-21 02.Code/01.-descriptive_AP.R
data cleaning dispersion estimates

@@ -115,26 +115,47 @@ pm25_constituents_hybrid <- pm25_constituents_hybrid %>%
 #################################
 dm_home_weekly <- read.csv("01.Data/BiSC_dispersion_models/imputed/weekly_estimate_DM_home_imputed.csv")
 
-# quick check of the data
-dplyr::glimpse(dm_home_weekly)
-skimr::skim(dm_home_weekly)
+no2_dm <- read.csv("01.Data/BiSC_comparison_data/predictions/DM/no2_dm.csv")
+pm25_dm <- read.csv("01.Data/BiSC_comparison_data/predictions/DM/pm25_dm.csv") 
+bc_dm <- read.csv("01.Data/BiSC_comparison_data/predictions/DM/bc_dm.csv") 
 
-# dates in proper format
-dm_home_weekly$date_start <- as.Date(dm_home_weekly$date_start)
-dm_home_weekly$date_end <- as.Date(dm_home_weekly$date_end)
+# quick check of the data 
+dplyr::glimpse(no2_dm)
+dplyr::glimpse(pm25_dm)
+dplyr::glimpse(bc_dm)
 
-# --- generate separete files --- #
-no2_dm <- dm_home_weekly %>% 
-          dplyr::select(subject_id, weeks, date_start, date_end, no2.total) %>% 
-          dplyr::rename(no2_dm = no2.total)
+#######################################
+### --- dates in proper format --- ###
+#####################################
 
-pm25_dm <- dm_home_weekly %>% 
-           dplyr::select(subject_id, weeks, date_start, date_end, pm2.5.total) %>% 
+### --- NO2 --- ###
+no2_dm$date_start <- as.Date(no2_dm$date_start)
+no2_dm$date_end <- as.Date(no2_dm$date_end)
+
+### --- PM25 --- ###
+pm25_dm$date_start <- as.Date(pm25_dm$date_start)
+pm25_dm$date_end <- as.Date(pm25_dm$date_end)
+
+### --- BC --- ###
+bc_dm$date_start <- as.Date(bc_dm$date_start)
+bc_dm$date_end <- as.Date(bc_dm$date_end)
+
+# --- select some variables per each dataset --- # 
+no2_dm <- no2_dm %>% 
+          dplyr::select(gid, subject_id, weeks, date_start, date_end,
+                        NO2.total_DispersionModels, lon, lat) %>% 
+          dplyr::rename(no2_dm = NO2.total_DispersionModels)
+
+pm25_dm <- pm25_dm %>%
[…]

### 2023-11-21 02.Code/01.-descriptive_AP.R
dispersion estimates

@@ -161,6 +161,7 @@ dplyr::glimpse(no2_dm)
 dplyr::glimpse(pm25_dm)
 dplyr::glimpse(bc_dm)
 
+
 ###########################
 ### --- LUR models --- ###
 #########################

### 2023-11-21 02.Code/01.-descriptive_AP.R
dm plot

@@ -140,6 +140,9 @@ pm25_dm$date_end <- as.Date(pm25_dm$date_end)
 bc_dm$date_start <- as.Date(bc_dm$date_start)
 bc_dm$date_end <- as.Date(bc_dm$date_end)
 
+# arrange the no2 data 
+no2_dm <- no2_dm %>% dplyr::arrange(subject_id, weeks)
+
 # --- select some variables per each dataset --- # 
 no2_dm <- no2_dm %>% 
           dplyr::select(gid, subject_id, weeks, date_start, date_end,
@@ -147,19 +150,27 @@ no2_dm <- no2_dm %>%
           dplyr::rename(no2_dm = NO2.total_DispersionModels)
 
 pm25_dm <- pm25_dm %>%
-           dplyr::select(gid, subject_id, weeks, date_start, date_end, 
-                         pm2.5.total, lon, lat) %>% 
+           dplyr::select(subject_id, weeks, pm2.5.total) %>% 
            dplyr::rename(pm25_dm = pm2.5.total)
 
 bc_dm <- bc_dm %>% 
-         dplyr::select(gid, subject_id, weeks, date_start, date_end,
-                       BC.total, lon, lat) %>% 
+         dplyr::select(subject_id, weeks, BC.total) %>% 
          dplyr::rename(bc_dm = BC.total)
 
 
 dplyr::glimpse(no2_dm)
 dplyr::glimpse(pm25_dm)
 dplyr::glimpse(bc_dm)
+
+### --- Join all DM estimates --- ### 
+dm_estimates <- no2_dm %>% 
+                dplyr::inner_join(pm25_dm, by = c("subject_id", "weeks")) %>% 
+                dplyr::inner_join(bc_dm, by = c("subject_id", "weeks"))
+
+dplyr::glimpse(dm_estimates)
+
+### --- Export data --- ###
+rio::export(dm_estimates, "01.Data/dm_estimates.csv")
 
 
 ###########################
@@ -289,9 +300,7 @@ zn_lur <- zn_lur %>%
 # We use lur data to arrange(sort the data)
 no2_lur <- no2_lur %>% dplyr::arrange(subject_id, weeks)
 
-#######################################
 ### --- Join all LUR estimates --- ### 
-#####################################
 lur_estimates <- no2_lur %>% 
                  dplyr::inner_join(pm25_lur, by = c("subject_id", "weeks")) %>% 
[…]

### 2023-11-22 02.Code/01.-descriptive_AP.R
LUR, DM, HM figure finished

@@ -110,6 +110,37 @@ pm25_constituents_hybrid <- pm25_constituents_hybrid %>%
                                           zn_hm = zn.total_hybridmodel)
 
 
+# --- select variables --- #
+no2_hybrid <- no2_hybrid %>% 
+              dplyr::select(gid, subject_id, weeks, date_start, date_end, no2_hm)
+
+
+pm25_hybrid <- pm25_hybrid %>% 
+               dplyr::select(subject_id, weeks, pm25_hm)
+
+bc_hybrid <- bc_hybrid %>% 
+             dplyr::select(subject_id, weeks, bc_hm)
+
+lon_lat <- no2_dm %>% 
+  dplyr::select(subject_id, weeks, lon, lat)
+
+# --- join estimates --- # 
+hm_estimates <- no2_hybrid %>%  
+                dplyr::inner_join(pm25_hybrid, by = c("subject_id", "weeks")) %>% 
+                dplyr::inner_join(bc_hybrid, by = c("subject_id", "weeks")) %>% 
+                dplyr::inner_join(lon_lat, by = c("subject_id", "weeks"))
+                
+dplyr::glimpse(hm_estimates)            
+
+
+
+
+
+
+
+
+
+
 ##################################
 ### --- Dispersion models --- ###
 #################################
@@ -531,7 +562,7 @@ library(ggplot2)
 # Lets check the extent of the shapefile
 print(st_bbox(amb_shp))
 
-# LUR estimates NO2 
+# DM estimates NO2 
 no2_plot_dm <- ggplot() +
   geom_sf(data = filter_amb_shp, fill = "lightgrey", color = "white") +
   geom_sf(data = dm_estimates_sf, aes(color = no2_dm), size = 1.0, alpha = 0.7) +
@@ -541,7 +572,7 @@ no2_plot_dm <- ggplot() +
        color = expression(paste("NO"[2], " (µg/m"^3*")"))) +
   xlim(c(st_bbox(filter_amb_shp)[1], st_bbox(filter_amb_shp)[3])) +
   ylim(c(st_bbox(filter_amb_shp)[2], st_bbox(filter_amb_shp)[4])) +
-  #facet_grid(. ~ "NO2 LUR model") +
+  #facet_grid(. ~ "NO2 DM model") +
   theme_bw() + 
   theme(plot.title = element_text(face = "bold"),
         legend.position = "bottom", 
@@ -558,7 +589,7 @@ no2_plot_dm <- ggplot() +
[…]

### 2023-11-22 02.Code/01.-descriptive_AP.R
notched boxplot

@@ -121,14 +121,18 @@ pm25_hybrid <- pm25_hybrid %>%
 bc_hybrid <- bc_hybrid %>% 
              dplyr::select(subject_id, weeks, bc_hm)
 
+pm25_constituents_hybrid <- pm25_constituents_hybrid %>% 
+                            dplyr::select(subject_id, weeks, fe_hm, cu_hm, zn_hm)
+
 lon_lat <- no2_dm %>% 
   dplyr::select(subject_id, weeks, lon, lat)
 
 # --- join estimates --- # 
 hm_estimates <- no2_hybrid %>%  
                 dplyr::inner_join(pm25_hybrid, by = c("subject_id", "weeks")) %>% 
-                dplyr::inner_join(bc_hybrid, by = c("subject_id", "weeks")) %>% 
-                dplyr::inner_join(lon_lat, by = c("subject_id", "weeks"))
+                dplyr::inner_join(bc_hybrid, by = c("subject_id", "weeks")) %>%
+                dplyr::inner_join(pm25_constituents_hybrid, by = c("subject_id", "weeks")) %>% 
+                dplyr::inner_join(lon_lat, by = c("subject_id", "weeks")) 
                 
 dplyr::glimpse(hm_estimates)            
 
@@ -782,6 +786,17 @@ hm_map_plot
 ##################################
 ggsave(plot = hm_map_plot, "03.Outputs/figures/hm_map_plot.png", 
        dpi = 600, width = 10, height = 5, units = "in")
+
+
+
+###################################################################################################
+
+########################
+### --- Boxplot --- ###
+######################
+dplyr::glimpse(lur_estimates)
+dplyr::glimpse(dm_estimates)
+dplyr::glimpse(hm_estimates)
 
 
 

### 2023-11-22 02.Code/01.-descriptive_AP.R
details

@@ -1,6 +1,7 @@
 # Install packages
 pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate, 
-               mice, ggmice, plyr, data.table, mapview, ggmap, omsdata, sf, sp, patchwork)
+               mice, ggmice, plyr, data.table, mapview, ggmap, omsdata, sf, sp, patchwork,
+               ggpattern)
 
 ##############################
 ### --- Load the data --- ###
@@ -787,8 +788,6 @@ hm_map_plot
 ggsave(plot = hm_map_plot, "03.Outputs/figures/hm_map_plot.png", 
        dpi = 600, width = 10, height = 5, units = "in")
 
-
-
 ###################################################################################################
 
 ########################
@@ -799,7 +798,125 @@ dplyr::glimpse(dm_estimates)
 dplyr::glimpse(hm_estimates)
 
 
+# create a label and renaming variables 
+lur_estimates <- lur_estimates %>% 
+                  dplyr::mutate(model = "LUR") %>% 
+                  dplyr::ungroup() %>% 
+                  dplyr::rename(no2 = no2_lur,
+                                pm25 = pm25_lur,
+                                bc = bc_lur,
+                                fe = fe_lur, 
+                                cu = cu_lur,
+                                zn = zn_lur)
 
 
+dm_estimates <- dm_estimates %>% 
+                dplyr::mutate(model = "DM") %>% 
+                dplyr::rename(no2 = no2_dm, 
+                              pm25 = pm25_dm, 
+                              bc = bc_dm)
+
+
+hm_estimates <- hm_estimates %>% 
+                dplyr::mutate(model = "HM") %>% 
+                dplyr::rename(no2 = no2_hm, 
+                              pm25 = pm25_hm, 
+                              bc = bc_hm,
+                              fe = fe_hm, 
+                              cu = cu_hm,
+                              zn = zn_hm)
+
+
+lur_estimates <- lur_estimates %>% dplyr::select(gid, subject_id, weeks, date_start, date_end, 
+                                                 no2, pm25, bc, model)
+
+
+dm_estimates <- dm_estimates %>% dplyr::select(gid, subject_id, weeks, date_start, date_end, 
+                                                 no2, pm25, bc, model)
+
+
+hm_estimates <- hm_estimates %>% dplyr::select(gid, subject_id, weeks, date_start, date_end, 
[…]

### 2023-11-23 02.Code/01.-descriptive_AP.R
correlation plot estimates

@@ -7,6 +7,17 @@ pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate,
 ### --- Load the data --- ###
 ############################
 
+meteo_variables <- readxl::read_xlsx("03.Outputs/predictions/hybrid_data_final_130923.xlsx")
+dplyr::glimpse(meteo_variables)
+
+bisc_meteo_weekly <- meteo_variables %>% 
+                     dplyr::select(gid, subject_id, weeks, date_start, date_end, 
+                                   lon, lat, avg_temperature, avg_relative_humidity, avg_precipitation) 
+
+
+rio::export(bisc_meteo_weekly, "03.Outputs/predictions/bisc_meteo_weekly.csv")
+
+
 #######################################################
 ### --- Load shapefile  Barcelona metropolitan --- ###
 #####################################################
@@ -126,7 +137,7 @@ pm25_constituents_hybrid <- pm25_constituents_hybrid %>%
                             dplyr::select(subject_id, weeks, fe_hm, cu_hm, zn_hm)
 
 lon_lat <- no2_dm %>% 
-  dplyr::select(subject_id, weeks, lon, lat)
+           dplyr::select(subject_id, weeks, lon, lat)
 
 # --- join estimates --- # 
 hm_estimates <- no2_hybrid %>%  
@@ -135,16 +146,10 @@ hm_estimates <- no2_hybrid %>%
                 dplyr::inner_join(pm25_constituents_hybrid, by = c("subject_id", "weeks")) %>% 
                 dplyr::inner_join(lon_lat, by = c("subject_id", "weeks")) 
                 
-dplyr::glimpse(hm_estimates)            
+dplyr::glimpse(hm_estimates)  # 42,894   
 
-
-
-
-
-
-
-
-
+### --- Export the data --- ###
+rio::export(hm_estimates, "01.Data/hm_estimates.csv")
 
 ##################################
 ### --- Dispersion models --- ###
@@ -220,6 +225,7 @@ cu_lur <- readRDS("01.Data/BiSC_comparison_data/predictions/LUR/cu_weekly_preg_2
 zn_lur <- readRDS("01.Data/BiSC_comparison_data/predictions/LUR/zn_weekly_preg_29092023.rds")
 
 # have a quick check to the data 
+view(no2_lur)
 dplyr::glimpse(no2_lur)
 dplyr::glimpse(pm25_lur)
 dplyr::glimpse(bc_lur)
[…]

### 2023-11-23 02.Code/00.-weekly_average_dispersion.R
changes

@@ -1,0 +1,10 @@
+# 00.- weekly_average_dispersion
+pacman::p_load(tidyverse, skimr, caret, tictoc, lubridate,  mice, ggmice, plyr, data.table)
+
+
+################################################
+### --- Define paths and load some data --- ###
+###############################################
+load("01.Data/FUR_updated/2023_07_28_finalversion_FUR_delivery.RData") # File FUR updated
+path_workplace_dm_reconstructed <- "01.Data/BiSC_dispersion_hourly/workplace_reconstructed_correct/" # this is new route in Toni's folder 
+path_weekly_workplace <- "01.Data/BiSC_dispersion_hourly/work_weekly_avg/"

### 2023-11-24 02.Code/01.-descriptive_AP.R
correlation plots

@@ -1,7 +1,7 @@
 # Install packages
 pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate, 
                mice, ggmice, plyr, data.table, mapview, ggmap, omsdata, sf, sp, patchwork,
-               ggpattern)
+               ggpattern, GGally)
 
 ##############################
 ### --- Load the data --- ###
@@ -556,7 +556,9 @@ ggsave(plot = lur_map_plot, "03.Outputs/figures/lur_map_plot.png",
 dm_estimates_plot <- dm_estimates %>% 
                      dplyr::ungroup()
 
-dm_estimates_plot <- dm_estimates_plot %>% dplyr::group_by(gid) %>%  dplyr::summarise_all(funs(mean))
+dm_estimates_plot <- dm_estimates_plot %>% dplyr::group_by(gid) %>%  
+  dplyr::summarise_all(funs(mean))
+
 dplyr::glimpse(dm_estimates_plot)
 
 dm_estimates_plot <- dm_estimates_plot %>% 
@@ -971,5 +973,63 @@ estimates_correlation_data <- lur_estimates_filtered %>%
                               dplyr::inner_join(hm_estimates_filtered, by = c("subject_id", "weeks"))
 
 dplyr::glimpse(estimates_correlation_data)
+
+
+#######################################
+### --- export correlation data --- ###
+#######################################
+rio::export(estimates_correlation_data, "01.Data/estimates_correlation_data.csv")
+
+####################################################################################
+
+
+####################################################
+### --- Correlation plot long term exposure --- ###
+##################################################
+pacman::p_load(GGally)
+library("GGally")
+
+estimates_correlation_data <- read.csv("01.Data/estimates_correlation_data.csv")
+dplyr::glimpse(estimates_correlation_data)
+estimates_correlation_data %>% dplyr::group_by(subject_id)
+
+###################################
+### --- Data for the plots --- ###
+#################################
+
+# short - term exposure data 
+estimates_correlation_data <- estimates_correlation_data %>% 
+                              dplyr::select(subject_id, weeks, 
+                                            no2_lur:bc_lur, no2_dm:bc_hm)
+
+dplyr::glimpse(estimates_correlation_data)
+
+names(estimates_correlation_data) <- c("ID", "weeks", 
+                                       "NO2 LUR", "PM25 LUR", "BC LUR",
[…]

### 2023-11-24 02.Code/01.-descriptive_AP.R
saving correlation plot

@@ -1012,7 +1012,7 @@ names(estimates_correlation_data) <- c("ID", "weeks",
 # long - term exposure data 
 preg_cor_data <- estimates_correlation_data  %>% 
                  dplyr::group_by(ID) %>% 
-                 dplyr::summarise(dplyr::across(3:9, mean))
+                 dplyr::summarise(dplyr::across(2:10, mean))
 
 dplyr::glimpse(preg_cor_data)
 
@@ -1021,15 +1021,33 @@ dplyr::glimpse(preg_cor_data)
 ### --- PLotting correlation matrix --- ###
 ###########################################
 
-### ---  Short - term exposure correlation matrix --- ### 
-ggcorr(estimates_correlation_data[, 3:11], 
-       method = c("pairwise", "spearman"), label = TRUE, label_size = 4 , label_round = 2)
-
-
 ### ---  Long - term exposure correlation matrix --- ### 
-ggcorr(predexpos[predexpos,3:7], method = c("pairwise", "spearman"), 
-       # nbreaks = 7, palette = "PRGn",
-       label = TRUE, label_size = 4, label_round = 2)
+preg_corr <- ggcorr(preg_cor_data[, 2:10], 
+                    method = c("pairwise", "spearman"), 
+                    label = TRUE, label_size = 4 , label_round = 2) +
+  ggplot2::labs(title = "(A) Long-term exposure (Entire pregnancy)") + 
+  theme(legend.position = "none", 
+        plot.title = element_text(face = "bold", size = 20))
 
+preg_corr
+
+
+### ---  Short - term exposure correlation matrix --- ### 
+weekly_corr <-  ggcorr(estimates_correlation_data[, 3:11], 
+              method = c("pairwise", "spearman"),
+              label = TRUE, label_size = 4 , label_round = 2) +
+              ggplot2::labs(title = "(B) Short-term exposure (Weekly exposure)") +
+              theme(legend.position = "none", 
+                    plot.title = element_text(face = "bold", size = 20))
+
+weekly_corr
+
+
+# plotting things together
+correlation_fig <- preg_corr | weekly_corr
+correlation_fig
+
+### --- save figure -- ###
+ggsave(correlation_fig )
 
 

### 2023-11-26 02.Code/01.-descriptive_AP.R
HEXBIN plot

@@ -367,7 +367,7 @@ rio::export(lur_estimates, "01.Data/lur_estimates.csv")
 ##############################################################################################
 
 # Note 1: to plot the estimates we going to group by gid
-# Note 2: to summarize the long-term exposure we group by 
+# Note 2: to summarize the long-term exposure we group by subject_id
 
 ####################################
 ### --- Plotting estimates --- ###
@@ -1151,12 +1151,154 @@ short_corr_pollutant
 ggsave(plot = short_corr_pollutant, "03.Outputs/figures/short_corr_pollutant.png",
        dpi = 600, width = 10, height = 5, units = "in")
 
-
-
-
+######################################################
+### --- long-and-short-term correlation plots --- ### 
+####################################################
 corrplot_windows <- long_corr_pollutant / short_corr_pollutant
 ggsave(plot = corrplot_windows, "03.Outputs/figures/corrplot_windows.png",
        dpi = 600, width = 15, height = 7, units = "in")
+
+
+####################################
+### --- Hexbin scatterplots --- ###
+##################################
+library("tidyverse")
+library("colorspace")
+library("lubridate")
+library("gridExtra")
+
+
+# Note: This could be added in the supplementary materials 
+
+########################################
+### --- NO2 Hexbin scatterplots --- ###
+######################################
+
+NO2_HM_LUR <- estimates_correlation_data %>%
+              ggplot() +
+              geom_hex(aes(y =`NO2 HM`, x = `NO2 LUR`)) +
+              #scale_fill_continuous_sequential(palette = "spectral") +
+                scale_fill_distiller(palette = "Spectral") + 
+              coord_equal() +
+              theme_bw() + 
+              labs(fill="Count") +
+              ylab(expression(NO[2] ~ "HM predictions" ~ (mu * g/m^3))) + 
+              xlab(expression(NO[2] ~ "LUR predictions" ~ (mu * g/m^3))) +
[…]

### 2023-11-26 02.Code/01.-descriptive_AP.R
fix the map scales for NO2 models (LUR, DM, HM)

@@ -373,6 +373,9 @@ rio::export(lur_estimates, "01.Data/lur_estimates.csv")
 ### --- Plotting estimates --- ###
 ##################################
 
+lur_estimates <- read.csv("01.Data/lur_estimates.csv")
+
+
 ### -- LUR estimates --- ### 
 dplyr::glimpse(lur_estimates)
 
@@ -400,6 +403,8 @@ amb_plot <- ggplot() +
   theme(plot.title = element_text(color = "black", size = 14, face = "bold")) +
   facet_grid(. ~ "NO2 LUR model") + theme_bw()
 
+amb_plot 
+
 filter_amb_shp <- amb_shp %>% 
                   dplyr::filter(NOMMUNI %in% c("Badalona", "Barcelona", "Esplugues de Llobregat",
                                              "l'Hospitalet de Llobregat", "Sant Joan Despí", "Sant Just Desvern",
@@ -414,6 +419,8 @@ amb_plot_filtered <- ggplot() +
   facet_grid(. ~ "NO2 LUR model") + theme_bw() +
   xlim(c(st_bbox(filter_amb_shp)[1], st_bbox(filter_amb_shp)[3])) +
   ylim(c(st_bbox(filter_amb_shp)[2], st_bbox(filter_amb_shp)[4])) 
+
+amb_plot_filtered
 
 # Check CRS of shapefile
 print(sf::st_crs(amb_shp))
@@ -463,7 +470,8 @@ no2_plot_lur <- ggplot() +
   geom_sf(data = filter_amb_shp, fill = "lightgrey", color = "white") +
   geom_sf(data = lur_estimates_sf, aes(color = no2_lur), size = 1.0, alpha = 0.7) +
   theme_minimal() +
-  scale_color_viridis_c(option = "viridis") + # Using viridis green palette
+  scale_color_viridis_c(option = "viridis", direction = -1) + # Using viridis green palette
+  #scale_color_distiller(palette = "Spectral", direction = -1) + # Using Spectral palette
   labs(title = "(A) LUR model",  
        color = expression(paste("NO"[2], " (µg/m"^3*")"))) +
   xlim(c(st_bbox(filter_amb_shp)[1], st_bbox(filter_amb_shp)[3])) +
@@ -490,7 +498,8 @@ pm25_plot_lur <- ggplot() +
   geom_sf(data = filter_amb_shp, fill = "lightgrey", color = "white") +
[…]

### 2023-11-26 02.Code/01.-descriptive_AP.R
Legends fixed for all the plots

@@ -1239,6 +1239,8 @@ NO2_HEX <- NO2_HM_LUR | NO2_HM_DM
 NO2_HEX
 
 ### --- Export HEXBIN plot --- ### 
+ggsave(plot = NO2_HEX, "03.Outputs/figures/NO2_HEX.png",
+       dpi = 600, width = 10, height = 5, units = "in")
 
 #########################################
 ### --- PM25 Hexbin scatterplots --- ###
@@ -1278,7 +1280,8 @@ PM25_HEX <- PM25_HM_LUR | PM25_HM_DM
 PM25_HEX
 
 ### --- Export HEXBIN plot --- ###
-
+ggsave(plot = PM25_HEX, "03.Outputs/figures/PM25_HEX.png",
+       dpi = 600, width = 10, height = 5, units = "in")
 
 #########################################
 ### --- BC Hexbin scatterplots --- ###
@@ -1313,9 +1316,18 @@ BC_HM_DM <- estimates_correlation_data %>%
 
 BC_HM_DM
 
-# PM25 predictions comparison 
+# BC predictions comparison 
 BC_HEX <- BC_HM_LUR | BC_HM_DM
 BC_HEX
+
+
+### --- Export HEXBIN plot --- ###
+ggsave(plot = BC_HEX, "03.Outputs/figures/BC_HEX.png",
+       dpi = 600, width = 10, height = 5, units = "in")
+
+
+
+
 
 
 # Putting all the figures together 
@@ -1323,8 +1335,12 @@ NO2_HEX / PM25_HEX / BC_HEX
 
 
 ##################################################
-### --- Rescalling MAP legend comparison --- ####
+### --- Rescalling MAP legend comparison --- ###############################################
 ################################################
+
+###########################
+### --- NO2 models --- ###
+#########################
 
 # Calculate overall min and max for NO2 from the summaries you have
 min_no2 <- min(lur_estimates_sf$no2_lur, dm_estimates_sf$no2_dm, hm_estimates_sf$no2_hm, na.rm = TRUE)
@@ -1333,7 +1349,6 @@ max_no2 <- max(lur_estimates_sf$no2_lur, dm_estimates_sf$no2_dm, hm_estimates_sf
 # Define the breaks you want to show in the legend
 legend_breaks <- c(min_no2, 50, 100, max_no2)
 legend_labels <- c(as.integer(min_no2), "50", "100", as.integer(max_no2))
-
 
 ##############################
[…]

### 2023-11-28 supplementary_BiSC_outdoor_AP.tex
changes in the manuscript

@@
 \usepackage{graphicx}
-\usepackage{authblk} % this is for multiple authors 
+\usepackage{authblk} % this is for multiple autho  rs 
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \usepackage{cleveref}
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
 \Crefname{figure}{Fig.}{Figs.}
+
+% reference configuration
+\usepackage[authoryear,round]{natbib}
+\bibliographystyle{apalike} % for an APA-like style
+\usepackage{amsmath}
 
@@
 
-% Index table with to put order to supplementary materials
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Land Use Regression modelling --- %%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Land Use Regression modelling details}
 
-\subsection{Predictor variables}
+\subsection{Predictor variables and sources of information}
+
 \subsection{Performance metrics}
-
-\section{Dispersion modelling details}
-
-\subsection{Predictor variables}
-\subsection{Performance metrics}
-
-\section{Model evaluation metrics}
-
 The following metrics were used to evaluate the LUR model for all the air pollution assessed.
@@
 
-For the DM models we evaluate their performance using a different set of metrics as following:
+\subsection{Model performance}
 
-% --- Percentage of observed-modelled pairs (n) --- %
-\begin{equation}
-n = \frac{N_{\text{pairs used}}}{N_{\text{total pairs}}} \times 100
-\end{equation}
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Dispersion modelling --- %%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% --- Factor of Two (FAC2) --- % 
-\begin{equation}
-\text{FAC2} = \frac{1}{n} \sum_{i=1}^{n} U,
-\end{equation}
-%Where \( U \) is 1 if \( 0.5 \leq \frac{M_i}{O_i} \leq 2 \) and 0 otherwise.
+\section{Dispersion modelling details}
 
-% --- Mean Bias (MB) --- % 
-\begin{equation}
-\text{MB} = \frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)
-\end{equation}
+\subsection{Predictor variables and sources of information}
 
-% --- Mean Gross Error (MGE) --- % 
-\begin{equation}
-\text{MGE} = \frac{1}{n} \sum_{i=1}^{n} |O_i - M_i|
-\end{equation}
+\subsection{Performance metrics}
 
-% --- Normalized Mean Bias (NMB) --- % 
-\begin{equation}
-\text{NMB} = \frac{\sum_{i=1}^{n} (M_i - O_i)}{\sum_{i=1}^{n} O_i}
-\end{equation}
+\subsection{Model performance}
 
-% --- Root Mean Square Error (RMSE) --- % 
-\begin{equation}
-\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)^2}
-\end{equation}
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Hybrid modelling --- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% --- Correlation Coefficient (r) --- % 
-\begin{equation}
-r = \frac{\sum_{i=1}^{n} (O_i - \bar{O})(M_i - \bar{M})}{\sqrt{\sum_{i=1}^{n} (O_i - \bar{O})^2 \sum_{i=1}^{n} (M_i - \bar{M})^2}}
-\end{equation}
-\vspace{0.5cm}
+\section{Hybrid modelling details}
+
+\subsection{Predictor variables and sources of information}
+
+\subsection{Performance metrics}
+
+\subsection{Model performance}
+
+%%% --- Table with performance metrics for Hybrid models --- %%%
+\begin{table}[ht]
+\centering
+\caption{Performance metric for hybrid models.}
+\begin{tabular}{llllll}
+\toprule
+\textbf{Dataset campaign} & \textbf{Pollutant} & \textbf{N} & \textbf{R\textsuperscript{2} (10-CV)} & \textbf{RMSE (10-CV)} \\
+\midrule
+BiSC-home & NO\textsubscript{2} & 1232 & 0.64 & 7.48 \\
+\multirow{5}{*}{BiSCAPE} & PM\textsubscript{2.5} & 161 & 0.66 & 3.45 \\
+& BC & 74 & 0.86 &  0.23 \\
+& Fe & ... & 0.54 & ... \\
+& Cu & ... & 0.70 & ... \\
+& Zn & ... & 0.42 & ... \\
+\bottomrule
+\end{tabular}
+\label{tablesX}
+\end{table}
+
+\subsection{Variable importance}
+
+%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
+% Put the figure text closer to the Figure 3
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=1.0\textwidth]{figures/fig_HM_allmodels_importance.png}
+\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
+\label{fig2}
+\end{figure}
+
+\newpage
+
+\subsection{Comparison between observed measurements and predictions}
+
+%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
+% Put the figure text closer to the Figure 3
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=1.0\textwidth]{figures/fig_HM_test_all_models.png}
+\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
+\label{fig2}
+\end{figure}
 

### 2023-11-28 manuscript_BiSC_outdoor_AP.tex

@@
 \usepackage{graphicx}
-\usepackage{authblk} % this is for multiple authors 
+\usepackage{authblk} % this is for multiple autho  rs 
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 
-Air pollution is the leading environmental risk factor for human health worldwide.
+Air pollution is the leading environmental risk factor for human health worldwide linked to wide range of adverse health outcomes.
 A detailed and comprehensive exposure assessment of outdoor air pollution considering both spatial and temporal variability is fundamental to conduct epidemiological studies.
@@
 In this study, within the framework of the Barcelona Life Study Cohort (BiSC), we developed LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses.
-LUR and DM, were constructed using conventional state-of-the-art techniques, while HM was innovatively designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
+LUR and DM, were constructed using conventional techniques, while HM was designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
 The overall aim is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates, intended to be used in the prenatal exposure assessment within the BiSC cohort.
@@
 In contrast, the OOB approach trains the model on two-thirds of the data and uses the residual third for external validation.
-We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}
+We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}.
 
@@
 
-% --- Performance --- %
+% --- Distribution of the monitoring campaigns --- % 
+\subsection{Distribution of monitoring campaigns}
+
+Distribution of the concentration for the BiSCAPE and BiSC-Home campaign for all the pollutants were shown in \hyperref[table1]{Table 1}.
+For all the air pollutants assessed, concentration varied substantially between pollutants.
+
+%%% Table 1. Outdoor air pollution concentration from BiSC-home and BiSCAPE campaigns (2018-2021)
+
+%%% --- Table 1. Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns --- %%%
+\begin{table}[h!]
+\centering
+\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns}
+\begin{tabular}{lllllllll}
+\toprule
+Campaign & Pollutant & Year & Units & mean & sd & p25 & p50 & p75 \\ 
+\midrule
+BiSC-home & NO\textsubscript{2} & 2018-2021 & \(\mu\)g/m\textsuperscript{3} & 38.8 & 12.4 & 30.0 & 37.5 & 45.8 \\ 
+\multirow{5}{*}{BiSCAPE} 
+& PM\textsubscript{25} & 2021-2022 & \(\mu\)g/m\textsuperscript{3} & 13.4 & 6.14 & 9.06 & 11.3 & 16.2 \\ 
+& BC & 2021-2022 & \(\mu\)g/m\textsuperscript{3} & 1.27 & 0.50 & 0.88 & 1.22 & 1.52 \\ 
+& Fe & 2021-2022 & \(\mu\)g/m\textsuperscript{3} & 0.25 & 0.13 & 0.15 & 0.21 & 0.32 \\ 
+& Cu & 2021-2022 & ng/m\textsuperscript{3} & 7.10 & 3.82  & 4.37 & 6.36 & 8.98 \\ 
+& Mn & 2021-2022 & ng/m\textsuperscript{3} & 40.7 & 36.5 & 18.9 & 28.2 & 46.7 \\ 
+\bottomrule
+\end{tabular}
+\end{table}
+
+The spatial distribution of the campaigns were showed in \cref{fig1}.
+
 Since models were built based on different approaches there where not directly comparable due to differences on a temporal basis
 
-Since different approaches were to built outdoor air pollution models were not directly comparable due to differences on temporal
+Since different approaches were used to built outdoor air pollution models these were not directly comparable due to differences on temporal
 
@@
 
-Performance metrics for LUR and DM are shown in the supplementary materials.
+Performance metrics for LUR, DM and HM are shown in the supplementary materials files.
 
-\begin{table}[ht]
-\centering
-\caption{Performance metric for hybrid models.}
-\begin{tabular}{llllll}
-\toprule
-\textbf{Model} & \textbf{Pollutant} & \textbf{Dataset campaign} & \textbf{N} & \textbf{R\textsuperscript{2} (10-CV)} & \textbf{RMSE (10-CV)} \\
-\midrule
-\multirow{6}{*}{HM} & NO\textsubscript{2} & BiSC-hchatome & 1232 & 0.64 & 7.48 \\
-& PM\textsubscript{2.5} & BiSCAPE & 161 & 0.66 & 3.45 \\
-& BC & BiSCAPE & 74 & 0.86 &  0.23 \\
-& Fe & BiSCAPE & ... & 0.54 & ... \\
-& Cu & BiSCAPE & ... & 0.70 & ... \\
-& Zn & BiSCAPE & ... & 0.42 & ... \\
-\bottomrule
-\end{tabular}
-\label{tab:my_label}
-\end{table}
+\
+% --- Exposure concentrations --- %
+Predictions from outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in FIGURE X.
+The exposure range was larger for the LUR model for all the pollutants assessed.
+Dispersion model predicted lower exposure concentration values than LUR and Hybrid.
+In addition to that HM show less variability in the exposure concentration.
 
-% --- Exposure concentrations --- %
-Outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in FIGURE X. The exposure range was larger for the LUR model
-
-% --- Correlation between measures and model estimates --- %
-Spearman correlation of air pollution exposure during the entire pregnancy between measurements, LUR, DM and HM at the residential addresses ranged from r = x to x (figure X).
+%%% --- Figure x. Boxplot models comparison --- %%%
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=1.0\textwidth]{figures/boxplot_models.png}
+\caption{Distribution exposure concentration for the entire pregnancy period \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
+\end{figure}
 
@@
 
-%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
+\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Spatial distribution of outdoor air pollution estimates  --- %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%% --- Figure. Average exposure during the entire pregnancy period --- %%%
 % Put the figure text closer to the Figure 3
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/fig_HM_test_all_models.png}
-\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
-\label{fig2}
+\includegraphics[width=1.0\textwidth]{figures/NO2_models_map.png}
 \end{figure}
 
-%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
 % Put the figure text closer to the Figure 3
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/lur_map_plot.png}
-\caption{Pregnancy exposure for the entire pregnancy period \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
-\label{fig2}
+\includegraphics[width=1.0\textwidth]{figures/PM25_models_map.png}
 \end{figure}
 
+% Put the figure text closer to the Figure 3
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=1.0\textwidth]{figures/BC_models_map.png}
+\caption{Pregnancy exposure for the entire pregnancy period \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% --- Correlation between measures and model estimates --- %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 (figure 3.A).
+
+Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 (figure 3.B).
+
+% Here add the correlation plot (long and short term)
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=1.0\textwidth]{figures/correlation_fig.png}
+\caption{Correlation between traffic-related air pollutants for LUR, DM, HM in different windows of exposure \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
+\end{figure}
+
+Correlation between the predicted air pollutants for the different models assessed was moderate to strongly correlated for both long-and-short term exposure.
+
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Discussion}
-We successfully developed spatiotemporal models using three different approaches (LUR, dispersion and hybrid models).
-All models had moderate to good performance evaluated by the defined metrics in the \textbf{methods section} and displayed in more detail in the \textbf{Supplementary materials}
+We successfully developed spatiotemporal models using three different approaches (LUR, DM and HM).
+All models show moderate to good performance evaluated by the defined metrics in the \textbf{methods section} and displayed in more detail in the \textbf{Supplementary materials}.
 

### 2023-11-28 02.Code/01.-descriptive_AP.R
working on table 1

@@ -902,7 +902,6 @@ models_estimates$model <- factor(models_estimates$model,
 
 
 library(ggplot2)
-pacman
 library(ggpattern)
 
 ##################################
@@ -918,7 +917,8 @@ no2_boxplot <- ggplot2::ggplot(data = models_estimates,
                                'DM' = '#3b528b', 
                                'HM' = '#21918c')) +
   ylab(bquote(NO[2] ~ (mu*g/m^3))) +
-  theme_bw() + theme(legend.position = 'none')
+  theme_bw() +
+  theme(legend.position = 'none')
 
 
 ### --- PM25 model boxplot --- ### 
@@ -932,7 +932,8 @@ pm25_boxplot <- ggplot2::ggplot(data = models_estimates,
                                'DM' = '#3b528b', 
                                'HM' = '#21918c')) +
   ylab(bquote(PM[25] ~ (mu*g/m^3))) +
-  theme_bw() + theme(legend.position = 'none')
+  theme_bw() + 
+  theme(legend.position = 'none')
 
 ### --- BC model boxplot --- ### 
 bc_model_boxplot <- ggplot2::ggplot(data = models_estimates, 
@@ -945,14 +946,29 @@ bc_model_boxplot <- ggplot2::ggplot(data = models_estimates,
                                'DM' = '#3b528b', 
                                'HM' = '#21918c')) +
   ylab(bquote(BC ~ (mu*g/m^3))) + ylim(c(0, 9)) +
-  theme_bw() + theme(legend.position = 'none')
+  theme_bw() + 
+  theme(legend.position = 'none')
 
+
+### --- Putting al the graphs together --- ###
 boxplot_models <- no2_boxplot | pm25_boxplot | bc_model_boxplot
-boxplot_models
+boxplot_models + theme(aspect.ratio = 1)
 
 ### --- saving the figure --- ### 
 ggsave(plot = boxplot_models , "03.Outputs/figures/boxplot_models.png",
-       dpi = 600, width = 15, height = 5, units = "in")
+       dpi = 600, width = 10, height = 3, units = "in")
+
+
+
+
+
+
+
+
+
+
+
+
 
 
 ##################################
@@ -1326,10 +1342,6 @@ ggsave(plot = BC_HEX, "03.Outputs/figures/BC_HEX.png",
        dpi = 600, width = 10, height = 5, units = "in")
 
 
-
-
-
-
 # Putting all the figures together 
[…]

### 2023-11-29 02.Code/01.-descriptive_AP.R
changes on the Boxplot (adding pm25 elements)

@@ -813,8 +813,6 @@ ggsave(plot = hm_map_plot, "03.Outputs/figures/hm_map_plot_v2.png",
        dpi = 600, width = 10, height = 5, units = "in")
 
 
-
-
 ### check all the plots ###
 lur_map_plot / dm_map_plot/ hm_map_plot
 
@@ -822,9 +820,6 @@ lur_map_plot / dm_map_plot/ hm_map_plot
 lur_estimates_sf %>% dplyr::select(no2_lur, pm25_lur, bc_lur) %>% summary()
 dm_estimates_sf %>% dplyr::select(no2_dm, pm25_dm, bc_dm) %>% summary()
 hm_estimates_sf %>% dplyr::select(no2_hm, pm25_hm, bc_hm) %>% summary()
-
-
-
 
 
 ###################################################################################################
@@ -872,8 +867,9 @@ hm_estimates <- hm_estimates %>%
                               zn = zn_hm)
 
 
+# Note: Now I add fe, cu, zn to the datset for LUR and HM (we can't do the rbind with dm)
 lur_estimates <- lur_estimates %>% dplyr::select(gid, subject_id, weeks, date_start, date_end, 
-                                                 no2, pm25, bc, model)
+                                                 no2, pm25, bc, fe, cu, zn, model)
 
 
 dm_estimates <- dm_estimates %>% dplyr::select(gid, subject_id, weeks, date_start, date_end, 
@@ -881,7 +877,7 @@ dm_estimates <- dm_estimates %>% dplyr::select(gid, subject_id, weeks, date_star
 
 
 hm_estimates <- hm_estimates %>% dplyr::select(gid, subject_id, weeks, date_start, date_end, 
-                                               no2, pm25, bc, model)
+                                               no2, pm25, bc, fe, cu, zn, model)
 
 
 dplyr::glimpse(lur_estimates)
@@ -890,10 +886,23 @@ dplyr::glimpse(hm_estimates)
 
 # putting all the estimates together
 models_estimates <- rbind(lur_estimates, dm_estimates, hm_estimates)
+model_estimates_elements <- rbind(lur_estimates, hm_estimates)
+
+dplyr::glimpse(model_estimates_elements)
 
 # levels of model variable
 models_estimates$model <- as.factor(models_estimates$model)
-levels(models_estimates$model)
+levels(model_estimates$model)
+
[…]

### 2023-11-29 02.Code/01.-descriptive_AP.R
adding fe, cu, zn in boxplots

@@ -869,7 +869,9 @@ hm_estimates <- hm_estimates %>%
 
 # Note: Now I add fe, cu, zn to the datset for LUR and HM (we can't do the rbind with dm)
 lur_estimates <- lur_estimates %>% dplyr::select(gid, subject_id, weeks, date_start, date_end, 
-                                                 no2, pm25, bc, fe, cu, zn, model)
+                                                 no2, pm25, bc,
+                                                 #fe, cu, zn,
+                                                 model)
 
 
 dm_estimates <- dm_estimates %>% dplyr::select(gid, subject_id, weeks, date_start, date_end, 
@@ -877,7 +879,9 @@ dm_estimates <- dm_estimates %>% dplyr::select(gid, subject_id, weeks, date_star
 
 
 hm_estimates <- hm_estimates %>% dplyr::select(gid, subject_id, weeks, date_start, date_end, 
-                                               no2, pm25, bc, fe, cu, zn, model)
+                                               no2, pm25, bc, 
+                                               #fe, cu, zn, 
+                                               model)
 
 
 dplyr::glimpse(lur_estimates)
@@ -959,40 +963,81 @@ bc_model_boxplot <- ggplot2::ggplot(data = models_estimates,
   theme(legend.position = 'none')
 
 ### --- Fe model  boxplot --- ### 
-fe_model_boxplot <- ggplot2::ggplot(data = models_estimates, 
-                                    mapping = aes(x = model, y = bc, color = model, fill = model)) + 
+fe_model_boxplot <- ggplot2::ggplot(data = model_estimates_elements, 
+                                    mapping = aes(x = model, y = fe, color = model, fill = model)) + 
   geom_boxplot(alpha = 0.7, color = "black") + 
   #scale_color_manual(values = c('LUR' = '#440154', 
   #                               'DM' = '#3b528b', 
   #                              'HM' = '#21918c')) +
   scale_fill_manual(values = c('LUR' = '#440154', 
-                               'DM' = '#3b528b', 
                                'HM' = '#21918c')) +
-  ylab(bquote(BC ~ (mu*g/m^3))) + ylim(c(0, 9)) +
+  ylab(bquote(Fe ~ (mu*g/m^3))) + 
+  #ylim(c(0, 9)) +
[…]

### 2023-11-30 02.Code/01.-descriptive_AP.R
small changes

@@ -413,6 +413,16 @@ filter_amb_shp <- amb_shp %>%
                                              "Sant Cugat del Vallès", "Sant Boi de Llobregat","Cornellà de Llobregat"))
 
 
+amb_bcn <- amb_shp %>%  
+           dplyr::filter(NOMMUNI %in% c("Barcelona", "Badalona", 
+                                        "Esplugues de Llobregat",
+                                        "l'Hospitalet de Llobregat", "Sant Adrià de Besòs", 
+                                        "Santa Coloma de Gramenet", "Sant Joan Despí", "Sant Just Desvern",
+                                        "Cornellà de Llobregat"))
+
+
+
+
 amb_plot_filtered <- ggplot() + 
   geom_sf(data = filter_amb_shp, fill = "grey", color = "white")+ 
   theme(plot.title = element_text(color = "black", size = 14, face = "bold")) +
@@ -421,6 +431,19 @@ amb_plot_filtered <- ggplot() +
   ylim(c(st_bbox(filter_amb_shp)[2], st_bbox(filter_amb_shp)[4])) 
 
 amb_plot_filtered
+
+
+amb_bcn <- ggplot() + 
+  geom_sf(data = amb_bcn, fill = "grey", color = "white")+ 
+  theme(plot.title = element_text(color = "black", size = 14, face = "bold")) +
+  facet_grid(. ~ "NO2 LUR model") + theme_bw() +
+  xlim(c(st_bbox(amb_bcn)[1], st_bbox(amb_bcn)[3])) +
+  ylim(c(st_bbox(amb_bcn)[2], st_bbox(amb_bcn)[4])) 
+
+amb_bcn
+
+
+
 
 # Check CRS of shapefile
 print(sf::st_crs(amb_shp))
@@ -444,17 +467,33 @@ ggplot() +
 #####################################################################################
 
 
-# First, convert your points to an sf object
-lur_estimates_sf <- st_as_sf(lur_estimates_plot, coords = c("lon", "lat"), crs = 25831)
 
+lur_estimates <- read.csv("01.Data/lur_estimates.csv")
+
+
+# summarise all the pollutants 
+lur_estimates_plot <- lur_estimates %>% 
+  dplyr::ungroup()
+
+lur_estimates_plot <- lur_estimates_plot %>% dplyr::group_by(gid) %>%  
[…]

### 2023-11-30 manuscript_BiSC_outdoor_AP.tex

@@
 The second campaign was carried out in accordance with the ESCAPE guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
-During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and hybrid models.
+During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and HM.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % --- Distribution of the monitoring campaigns --- % 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Distribution of monitoring campaigns}
 
-Distribution of the concentration for the BiSCAPE and BiSC-Home campaign for all the pollutants were shown in \hyperref[table1]{Table 1}.
-For all the air pollutants assessed, concentration varied substantially between pollutants.
+Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants are shown in \hyperref[table1]{Table 1}.
+NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 \(\mu \text{g/m}^3\) and 97.3 \(\mu \text{g/m}^3\) with a mean (sd) of  38.8 (12.4) \( \mu \text{g/m}^3\).
+PM\textsubscript{2.5} observed levels ranged between 6.14 \(\mu \text{g/m}^3\) and 35.8 \( \mu \text{g/m}^3\) with a mean (sd) of 13.4 (6.14) \(\mu \text{g/m}^3\), while BC levels showed less variation during the BiSCAPE campaign ranged between 0.40 \(\mu \text{g/m}^3\) and 2.95 \(  \mu \text{g/m}^3\) with a mean of 1.27 (0.50) \( \mu \text{g/m}^3\).
+For PM\textsubscript{2.5} constituents concentration values observed during the BiSCAPE campaign varied substantially between elements with a mean of 0.25 for  FE,  7.10 for Cu, and a mean of 40.7 for Mn.
 
-%%% Table 1. Outdoor air pollution concentration from BiSC-home and BiSCAPE campaigns (2018-2021)
+\\
+The spatial distribution of the outdoor air pollution measurements campaigns were showed in \cref{fig1}.  Panel Fig 1.A show a high density coverage
 
-%%% --- Table 1. Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns --- %%%
+%%% ---------------------------------------------------------------- %%%
+%%% Table 1. Outdoor air pollution concentration BiSC-home, BiSCAPE  %%%
+%%% ---------------------------------------------------------------  %%%
 \begin{table}[h!]
 \centering
-\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns}
-\begin{tabular}{lllllllll}
+\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns (2018-2022)}
+\begin{tabular}{lllllllllll} % each one indicates a new column in the table 
 \toprule
-Campaign & Pollutant & Year & Units & mean & sd & p25 & p50 & p75 \\ 
+Campaign & Pollutant & Year  & mean & sd & min & p25 & p50 & p75 & max \\ 
 \midrule
-BiSC-home & NO\textsubscript{2} & 2018-2021 & \(\mu\)g/m\textsuperscript{3} & 38.8 & 12.4 & 30.0 & 37.5 & 45.8 \\ 
+BiSC-home, BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\ 
 \multirow{5}{*}{BiSCAPE} 
-& PM\textsubscript{25} & 2021-2022 & \(\mu\)g/m\textsuperscript{3} & 13.4 & 6.14 & 9.06 & 11.3 & 16.2 \\ 
-& BC & 2021-2022 & \(\mu\)g/m\textsuperscript{3} & 1.27 & 0.50 & 0.88 & 1.22 & 1.52 \\ 
-& Fe & 2021-2022 & \(\mu\)g/m\textsuperscript{3} & 0.25 & 0.13 & 0.15 & 0.21 & 0.32 \\ 
-& Cu & 2021-2022 & ng/m\textsuperscript{3} & 7.10 & 3.82  & 4.37 & 6.36 & 8.98 \\ 
-& Mn & 2021-2022 & ng/m\textsuperscript{3} & 40.7 & 36.5 & 18.9 & 28.2 & 46.7 \\ 
+& PM\textsubscript{25} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 13.4 & 6.14 & 5.01 & 9.06 & 11.3 & 16.2 & 35.8 \\ 
+& BC (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 1.27 & 0.50 & 0.40 & 0.88 & 1.22 & 1.52 & 2 .95 \\ 
+& Fe (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 0.25 & 0.13 & 0.07 & 0.15 & 0.21 & 0.32 & 0.77 \\ 
+& Cu (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
+& Mn (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
 \bottomrule
@@
 
-The spatial distribution of the campaigns were showed in \cref{fig1}.
-
-Since models were built based on different approaches there where not directly comparable due to differences on a temporal basis
-
-Since different approaches were used to built outdoor air pollution models these were not directly comparable due to differences on temporal
+Since different approaches were used to built outdoor air pollution models these were not directly comparable due to differences on temporal basis.
 
@@
 Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE.
-The performance of the PM25
+Performance metrics for LUR, DM and HM are shown in the supplementary materials.
 
-Performance metrics for LUR, DM and HM are shown in the supplementary materials files.
-
-\
 % --- Exposure concentrations --- %
@@
 
-%%% --- Figure x. Boxplot models comparison --- %%%
-\captionsetup[figure]{skip=6pt}
-% We add the figure with the study domain 
+%%% -------------------------------------------%%%%
+%%% --- Figure 2. Boxplot models comparison --- %%%
+%%% -------------------------------------------%%%%
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/boxplot_models.png}
-\caption{Distribution exposure concentration for the entire pregnancy period \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
+\includegraphics[width=1.0\textwidth]{figures/boxplot_all_models_estimates.png}
+\caption{Distribution of exposure to air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
 \end{figure}
@@
 
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 (figure 3.A).
-
-Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 (figure 3.B).
+Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 \textbf{(figure 4.A)}.
+Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 \textbf{(figure 4.B)}.
+Correlation between the air pollutant estimates for the different models assessed was moderate to strongly correlated for both long-and-short term exposure.
+Nitrogen dioxide
 
@@
 
-Correlation between the predicted air pollutants for the different models assessed was moderate to strongly correlated for both long-and-short term exposure.
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% --- Comparison/agreement between air pollution estimates --- %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%% --- Figure. Comparison/agreement between moodel predictions estimates --- %%%
+% Put the figure text closer to the Figure 3
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=1.0\textwidth]{figures/NO2_HEX.png}
+\end{figure}
+
+% Put the figure text closer to the Figure 3
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=1.0\textwidth]{figures/PM25_HEX.png}
+\end{figure}
+
+% Put the figure text closer to the Figure 3
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=1.0\textwidth]{figures/BC_HEX.png}
+\caption{Comparison/agreement between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
+\end{figure}
 
@@
 \section{Discussion}
-We successfully developed spatiotemporal models using three different approaches (LUR, DM and HM).
+We successfully developed spatiotemporal models using three different approaches (LUR, DM, and HM). Models show moderate to good performance
+
+spatiotemporal models using three different approaches (LUR, DM and HM).
 All models show moderate to good performance evaluated by the defined metrics in the \textbf{methods section} and displayed in more detail in the \textbf{Supplementary materials}.

### 2023-11-30 02.Code/01.-descriptive_AP.R
changes

@@ -1472,10 +1472,12 @@ NO2_HEX / PM25_HEX / BC_HEX
 # Calculate overall min and max for NO2 from the summaries you have
 min_no2 <- min(lur_estimates_sf$no2_lur, dm_estimates_sf$no2_dm, hm_estimates_sf$no2_hm, na.rm = TRUE)
 max_no2 <- max(lur_estimates_sf$no2_lur, dm_estimates_sf$no2_dm, hm_estimates_sf$no2_hm, na.rm = TRUE)
+min_no2
+max_no2
 
 # Define the breaks you want to show in the legend
-legend_breaks <- c(min_no2, 50, 100, max_no2)
-legend_labels <- c(as.integer(min_no2), "50", "100", as.integer(max_no2))
+legend_breaks <- c(min_no2, 26, 43, 60, max_no2)
+legend_labels <- c(as.integer(min_no2), "26", "43", "60", as.integer(max_no2))
 
 
 lur_estimates_sf$
@@ -1502,14 +1504,14 @@ NO2_LUR_model_plot <- ggplot() +
     option = "viridis", direction = 1, 
     limits = c(min_no2, max_no2),
     breaks = legend_breaks, # Set the breaks for the legend
-    labels = c(as.integer(min_no2), "50", "100", as.integer(max_no2))  # Format the labels as desired
+    labels = c(as.integer(min_no2), "26", "43", "60", as.integer(max_no2))  # Format the labels as desired
   ) + # Set the limits to the overall range
   labs(title = expression(paste("(A) NO"[2], " models")), 
        color = expression(paste("NO"[2], " (µg/m"^3*")"))) +
   xlim(c(st_bbox(amb_bcn)[1], st_bbox(amb_bcn)[3])) +
   ylim(c(st_bbox(amb_bcn)[2], st_bbox(amb_bcn)[4])) +
   theme_bw() + 
-  theme(plot.title = element_text(face = "bold"),
+  theme(plot.title = element_text(face = "bold", size = 16),
         legend.position = "none",
         axis.title.x = element_blank(),
         axis.title.y = element_blank(),
@@ -1523,7 +1525,7 @@ NO2_LUR_model_plot <- ggplot() +
         legend.key.size = unit(0.5, 'cm'),
         panel.border = element_blank(),
         panel.background = element_blank(),
-        strip.text = element_text(face = "bold", hjust = 0),
+        strip.text = element_text(face = "bold", hjust = 0, size = 12),
         strip.background = element_blank(),
         strip.placement = "outside"
   ) +
[…]

### 2023-11-30 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 \usepackage{authblk} % this is for multiple autho  rs 
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage[colorlinks=true, allcolors=black]{hyperref}
 \usepackage{cleveref}
@@
 \usepackage{multirow}
-
-% Configure for "Fig.":
-\crefname{figure}{Fig.}{Figs.}
-\Crefname{figure}{Fig.}{Figs.}
 
@@
 Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites were during the pandemic period and its resulting restrictions.
-For PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC models were developed using BiSCAPE data from the 34 sites.
+For PM\textsubscript{2.5}, BC  and PM\textsubscript{2.5} constituents models were developed using BiSCAPE data from the 34 sites.
 
-For the development of the LUR models, spatial variables that better represent the pollutant concentration distribution across the study area were created.
+For the development of the LUR models, spatial variables that better represent the air pollutant concentration distribution across the study area were created.
 Following the ESCAPE guidelines \cite{eeftens2012, beelen2013},  predictor variables were selected from three broad categories: traffic, land use, and urban configuration.
 A wide variety of buffer sizes were created, i.e, 25m, 50m, 100m, 300m, 500m, and 1000m to capture the area-specific predictor variable at every single location.
+Predictor variables were calculated using specific coordinates for each site.
 
-We applied a supervised stepwise process in which we sequentially added a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model.
+We applied a supervised stepwise process in which we sequentially added a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model \cite{eeftens2012, beelen2013}.
 This procedure was followed to build all the specific pollutant models.
 Previously to model development, we specified the direction expected of the effect that each predictor has in the model.
-For all predictor variables, univariate regression models with annual average air pollution concentration (NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC) were conducted, models with the highest adjusted explained variance (R\textsuperscript{2}) were selected.
+For all predictor variables, univariate regression models with annual average air pollution concentration (NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn)) were conducted, models with the highest adjusted explained variance (adjusted R\textsuperscript{2}) and the right direction of the effect were selected.
 We then repeated the process using the previous model and all the remaining predictor variables.
-Details of the LUR modelling and variables used can be found in the \textbf{Supplementary material X}.
+Details of the LUR modelling and variables used can be found in the \textbf{Supplementary materials} .
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \subsection{Distribution of monitoring campaigns}
 
-Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants are shown in \hyperref[table1]{Table 1}.
-NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 \(\mu \text{g/m}^3\) and 97.3 \(\mu \text{g/m}^3\) with a mean (sd) of  38.8 (12.4) \( \mu \text{g/m}^3\).
-PM\textsubscript{2.5} observed levels ranged between 6.14 \(\mu \text{g/m}^3\) and 35.8 \( \mu \text{g/m}^3\) with a mean (sd) of 13.4 (6.14) \(\mu \text{g/m}^3\), while BC levels showed less variation during the BiSCAPE campaign ranged between 0.40 \(\mu \text{g/m}^3\) and 2.95 \(  \mu \text{g/m}^3\) with a mean of 1.27 (0.50) \( \mu \text{g/m}^3\).
-For PM\textsubscript{2.5} constituents concentration values observed during the BiSCAPE campaign varied substantially between elements with a mean of 0.25 for  FE,  7.10 for Cu, and a mean of 40.7 for Mn.
-
+Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants is shown in \Cref{table1}.
+NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 and 97.3 \(\mu \text{g/m}^3\) with a mean of  38.8 \(\pm\) 12.4 \( \mu \text{g/m}^3\).
+PM\textsubscript{2.5} observed levels ranged between 6.14 and 35.8 \( \mu \text{g/m}^3\) with a mean of 13.4 \(\pm\) 6.14 \(\mu \text{g/m}^3\), while BC levels showed less variation during the BiSCAPE campaign, ranging from 0.40 to 2.95 \(  \mu \text{g/m}^3\) with a mean of 1.27 \(\pm\) 0.50 \( \mu \text{g/m}^3\).
+For PM\textsubscript{2.5} constituents  observed values during the BiSCAPE campaign varied substantially between elements with means of 0.25 \(\pm\) 0.13 \(  \mu \text{g/m}^3\)  for  Fe,  7.10 \(\pm\) 3.82 \( \text{ng/m}^3 \) for Cu, and 40.7 \(\pm\) 36.5 \( \text{ng/m}^3 \) for Zn.
 \\
-The spatial distribution of the outdoor air pollution measurements campaigns were showed in \cref{fig1}.  Panel Fig 1.A show a high density coverage
 
@@
 \caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns (2018-2022)}
+\label{table1} 
 \begin{tabular}{lllllllllll} % each one indicates a new column in the table 
@@
 & Cu (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
-& Mn (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
+& Zn (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
 \bottomrule
@@
 
-Since different approaches were used to built outdoor air pollution models these were not directly comparable due to differences on temporal basis.
+The spatial distribution of the outdoor air pollution sampling campaigns is shown in \Cref{fig1}.
+Figure 1A illustrates a high density coverage in the central area of Barcelona, where the BiSC-home campaign was conducted.
+Figure 1B exhibit a similar pattern of spatial distribution but with a reduced density in sampling locations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% --- Model performance: LUR, DM, HM --- % 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Model performance: LUR, DM, HM}
+
+Since different approaches were used to built outdoor air pollution models were not directly comparable due to differences on temporal basis.
 
@@
 
+\newpage
+
+\subsection{Comparison between models} % broad title then change this
+
 % --- Exposure concentrations --- %
-Predictions from outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in FIGURE X.
+Predictions from outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig2}.
 The exposure range was larger for the LUR model for all the pollutants assessed.
@@
 \includegraphics[width=1.0\textwidth]{figures/boxplot_all_models_estimates.png}
-\caption{Distribution of exposure to air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
+\caption{Distribution of exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
+\label{fig2}
 \end{figure}
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/NO2_models_map.png}
-\end{figure}
-
-% Put the figure text closer to the Figure 3
-\captionsetup[figure]{skip=6pt}
-% We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/PM25_models_map.png}
-\end{figure}
-
-% Put the figure text closer to the Figure 3
-\captionsetup[figure]{skip=6pt}
-% We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/BC_models_map.png}
-\caption{Pregnancy exposure for the entire pregnancy period \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
+\includegraphics[width=1.0\textwidth]{figures/all_maps.png}
+\caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
 \end{figure}

### 2023-12-01 02.Code/01.-descriptive_AP.R
changes in the script

@@ -1342,8 +1342,8 @@ NO2_HM_LUR <- estimates_correlation_data %>%
               coord_equal() +
               theme_bw() + 
               labs(fill="Count") +
-              xlim(c(0, 150)) +
-              ylim(c(0, 150)) +
+              #xlim(c(0, 150)) +
+              #ylim(c(0, 150)) +
               ylab(expression(NO[2] ~ "HM predictions" ~ (mu * g/m^3))) + 
               xlab(expression(NO[2] ~ "LUR predictions" ~ (mu * g/m^3))) +
               theme(aspect.ratio = 1) + 
@@ -1359,6 +1359,8 @@ NO2_HM_DM <- estimates_correlation_data %>%
   coord_equal() +
   theme_bw() + 
   labs(fill="Count") +
+  #xlim(c(0, 100)) +
+  #ylim(c(0, 100)) +
   ylab(expression(NO[2] ~ "HM predictions" ~ (mu * g/m^3))) + 
   xlab(expression(NO[2] ~ "DM predictions" ~ (mu * g/m^3))) +
   theme(aspect.ratio = 1)
@@ -1385,6 +1387,8 @@ PM25_HM_LUR <- estimates_correlation_data %>%
   coord_equal() +
   theme_bw() + 
   labs(fill="Count") +
+  #xlim(c(0, 60)) +
+  #ylim(c(0, 60)) +
   ylab(expression(PM[25] ~ "HM predictions" ~ (mu * g/m^3))) + 
   xlab(expression(PM[25] ~ "LUR predictions" ~ (mu * g/m^3))) +
   theme(aspect.ratio = 1) + 
@@ -1400,6 +1404,8 @@ PM25_HM_DM <- estimates_correlation_data %>%
   coord_equal() +
   theme_bw() + 
   labs(fill="Count") +
+  #xlim(c(0, 40)) +
+  #ylim(c(0, 40)) +
   ylab(expression(PM[25] ~ "HM predictions" ~ (mu * g/m^3))) + 
   xlab(expression(PM[25] ~ "DM predictions" ~ (mu * g/m^3))) +
   theme(aspect.ratio = 1)
@@ -1426,6 +1432,8 @@ BC_HM_LUR <- estimates_correlation_data %>%
   coord_equal() +
   theme_bw() + 
   labs(fill="Count") +
+ #xlim(c(0, 4)) +
+ #ylim(c(0, 4)) +
   ylab(expression(BC ~ "HM predictions" ~ (mu * g/m^3))) + 
[…]

### 2023-12-01 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 \usepackage{authblk} % this is for multiple autho  rs 
-\usepackage[colorlinks=true, allcolors=black]{hyperref}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \usepackage{cleveref}
@@
 This trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas \cite{boogaard2022}.
+\\
 
@@
 To overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}.
+\\
 
@@
 This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potential lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.
+\\
 
@@
 Among the different ML algorithms, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.
-
 Furthermore, the COVID-19 pandemic has introduced unprecedented challenges for air pollution modelling and exposure assessment.
 These challenges encompass changes in air pollution patterns, constraints on data collection due to health and safety restrictions (lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}.
+\\
 
@@
 Barcelona is one of the cities in Europe with the highest vehicle traffic density (6,000 vehicles/km\textsuperscript{2}) \cite{casallas2018}.
+Air pollution measurements were conducted during two separate monitoring campaigns.
+The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021.
+The second campaign was carried out in accordance with the ESCAPE guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
+During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and HM.
 
@@
 
-Air pollution measurements were conducted during two separate monitoring campaigns.
-The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021.
-The second campaign was carried out in accordance with the ESCAPE guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
-During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and HM.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For PM\textsubscript{2.5}, BC  and PM\textsubscript{2.5} constituents models were developed using BiSCAPE data from the 34 sites.
-
 For the development of the LUR models, spatial variables that better represent the air pollutant concentration distribution across the study area were created.
@@
 Predictor variables were calculated using specific coordinates for each site.
-
 We applied a supervised stepwise process in which we sequentially added a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model \cite{eeftens2012, beelen2013}.
@@
 We then repeated the process using the previous model and all the remaining predictor variables.
-Details of the LUR modelling and variables used can be found in the \textbf{Supplementary materials} .
+Details of the LUR modelling and variables used can be found in the \textbf{Table S1}.
 
@@
 This process was repeated \(k-times\), and the average of these measures was reported as the indicator of the model validity.
-\vspace{0.5cm} Details from the metrics used for model validation can be found in the \textbf{Supplementary material X}.
+Details from the metrics used for model validation can be found in the \textbf{Supplementary material X}.\\
 
@@
 \subsection{Dispersion model evaluation}
+ Briefly, once input data has been introduced to the ADMS Urban model, the calculation of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC concentration levels has been carried out.
 Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
-Details from the metrics used for dispersion model validation can be found in the \textbf{Supplementary material X}.\vspace{0.5cm}
+The closer these values are, the better is the air quality modelling.
+Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material X}.\vspace{0.5cm}
 
@@
 
-\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Put the figure text closer to the Figure 3
-\captionsetup[figure]{skip=6pt}
-% We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/NO2_HEX.png}
-\end{figure}
 
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/PM25_HEX.png}
-\end{figure}
-
-% Put the figure text closer to the Figure 3
-\captionsetup[figure]{skip=6pt}
-% We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/BC_HEX.png}
+\includegraphics[width=1.0\textwidth]{figures/HEX_all.png}
 \caption{Comparison/agreement between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}

### 2023-12-04 02.Code/01.-descriptive_AP.R
new correlation plots

@@ -936,8 +936,6 @@ dplyr::glimpse(hm_estimates)
 models_estimates <- rbind(lur_estimates, dm_estimates, hm_estimates)
 model_estimates_elements <- rbind(lur_estimates, hm_estimates)
 
-dplyr::glimpse(model_estimates_elements)
-
 # levels of model variable
 models_estimates$model <- as.factor(models_estimates$model)
 levels(model_estimates$model)
@@ -957,6 +955,10 @@ models_estimates$model <- factor(models_estimates$model,
                                 ordered = TRUE)
 
 
+
+
+dplyr::glimpse(model_estimates)
+view(model_estimates_elements)
 
 library(ggplot2)
 library(ggpattern)
@@ -1076,13 +1078,9 @@ ggsave(plot = boxplot_models , "03.Outputs/figures/boxplot_models.png",
 ggsave(plot = boxplot_models_constituents , "03.Outputs/figures/boxplot_models_elements.png",
        dpi = 600, width = 10, height = 3, units = "in")
 
-
-
 # NO2, PM25, BC, FE, CU, ZN (all models estimates )
 ggsave(plot = boxplot_all_models_estimates, "03.Outputs/figures/boxplot_all_models_estimates.png",
        dpi = 600, width = 10, height = 6, units = "in")
-
-
 
 ##################################
 ### --- Correlation plots --- ###
@@ -1340,7 +1338,8 @@ NO2_HM_LUR <- estimates_correlation_data %>%
               #scale_fill_continuous_sequential(palette = "spectral") +
               scale_fill_distiller(palette = "Spectral") + 
               coord_equal() +
-              theme_bw() + 
+              theme_bw(base_size = 10) + 
+              theme(axis.text = element_text(size = 10)) + 
               labs(fill="Count") +
               #xlim(c(0, 150)) +
               #ylim(c(0, 150)) +
@@ -1357,7 +1356,7 @@ NO2_HM_DM <- estimates_correlation_data %>%
   #scale_fill_continuous_sequential(palette = "spectral") +
   scale_fill_distiller(palette = "Spectral") + 
   coord_equal() +
-  theme_bw() + 
+  theme_bw(base_size = 10) + 
   labs(fill="Count") +
   #xlim(c(0, 100)) +
   #ylim(c(0, 100)) +
[…]

### 2023-12-04 02.Code/01.-descriptive_AP.R
HEXBIN final plot

@@ -955,8 +955,6 @@ models_estimates$model <- factor(models_estimates$model,
                                 ordered = TRUE)
 
 
-
-
 dplyr::glimpse(model_estimates)
 view(model_estimates_elements)
 
@@ -1126,11 +1124,15 @@ dplyr::glimpse(hm_estimates_filtered) %>% dplyr::group_by(subject_id)
 
 # select variables 
 lur_estimates_filtered 
+
 dm_estimates_filtered <- dm_estimates_filtered %>% 
-                         dplyr::select(subject_id, weeks, no2_dm, pm25_dm, bc_dm)
+                         dplyr::select(subject_id, weeks,
+                                       no2_dm, pm25_dm, bc_dm)
 
 hm_estimates_filtered <- hm_estimates_filtered %>% 
-                         dplyr::select(subject_id, weeks, no2_hm, pm25_hm, bc_hm)
+                         dplyr::select(subject_id, weeks,
+                                       no2_hm, pm25_hm, bc_hm,
+                                       fe_hm, cu_hm, zn_hm)
 
 
 estimates_correlation_data <- lur_estimates_filtered %>% 
@@ -1161,23 +1163,27 @@ estimates_correlation_data %>% dplyr::group_by(subject_id)
 ###################################
 ### --- Data for the plots --- ###
 #################################
+dplyr::glimpse(estimates_correlation_data)
 
 # short - term exposure data 
 estimates_correlation_data <- estimates_correlation_data %>% 
                               dplyr::select(subject_id, weeks, 
-                                            no2_lur:bc_lur, no2_dm:bc_hm)
+                                            no2_lur:zn_lur, 
+                                            no2_dm:zn_hm)
 
 dplyr::glimpse(estimates_correlation_data)
 
 names(estimates_correlation_data) <- c("ID", "weeks", 
                                        "NO2 LUR", "PM25 LUR", "BC LUR",
+                                       "FE LUR", "CU LUR", "ZN LUR", 
[…]

### 2023-12-04 02.Code/01.-descriptive_AP.R
Fe model map

@@ -481,13 +481,14 @@ lur_estimates_plot <- lur_estimates_plot %>% dplyr::group_by(gid) %>%
 dplyr::glimpse(lur_estimates_plot)
 
 lur_estimates_plot <- lur_estimates_plot %>% 
-  dplyr::select(gid, subject_id, no2_lur, pm25_lur, bc_lur, lon, lat) %>% 
+  dplyr::select(gid, subject_id, 
+                no2_lur, pm25_lur, bc_lur, fe_lur, cu_lur, zn_lur,
+                lon, lat) %>% 
   tidyr::drop_na()
 
 
 lur_estimates_sf <- st_as_sf(lur_estimates_plot, coords = c("lon", "lat"), crs = 25831)
 lur_estimates_wgs84 <- st_transform(lur_estimates_sf, 4326)
-
 
 # Plotting with ggplot2
 dplyr::glimpse(amb_shp)
@@ -744,7 +745,10 @@ hm_estimates_plot <- hm_estimates_plot %>%
 dplyr::glimpse(hm_estimates_plot)
 
 hm_estimates_plot <- hm_estimates_plot %>% 
-                     dplyr::select(gid, subject_id, no2_hm, pm25_hm, bc_hm, lon, lat) %>% 
+                     dplyr::select(gid, subject_id, 
+                                   no2_hm, pm25_hm, bc_hm, 
+                                   fe_hm, cu_hm, zn_hm, 
+                                   lon, lat) %>% 
                      tidyr::drop_na()
 
 
@@ -953,7 +957,6 @@ levels(model_estimates_elements$model)
 models_estimates$model <- factor(models_estimates$model,
                                 levels = c("LUR", "DM", "HM"),
                                 ordered = TRUE)
-
 
 dplyr::glimpse(model_estimates)
 view(model_estimates_elements)
@@ -1594,14 +1597,11 @@ legend_breaks <- c(min_no2, 26, 43, 60, max_no2)
 legend_labels <- c(as.integer(min_no2), "26", "43", "60", as.integer(max_no2))
 
 
-lur_estimates_sf$
-
 # We use this vector to filtered the data
 subjects_out <- c("10051711", "10032711", "12031211", 
                   "11003411", "10038011", "12032011",
                   "12025111", "10052911", "12020611", "12025611")
 
-subject_out <- c("")
 
 lur_estimates_sf <- lur_estimates_sf %>%
[…]

### 2023-12-05 02.Code/01.-descriptive_AP.R
new domain figure

@@ -2110,7 +2110,7 @@ FE_LUR_model_plot
 #################################
 ### --- Fe Hybrid model --- ###
 ################################
-Fe_HM_model_plot <- ggplot() +
+FE_HM_model_plot <- ggplot() +
   geom_sf(data = amb_bcn, fill = "lightgrey", color = "white") +
   geom_sf(data = hm_estimates_sf, aes(color = fe_hm), size = 1.0, alpha = 1.0) +
   theme_minimal() +
@@ -2145,8 +2145,6 @@ Fe_HM_model_plot <- ggplot() +
     panel.background = element_blank()
   ) +
   facet_grid(.~'Hybrid model')  # Add the model name as facet label
-
-
 
 ### --- FE models comparison ---- ###
 FE_models_map <-  FE_LUR_model_plot | FE_HM_model_plot | plot_spacer()
@@ -2187,7 +2185,7 @@ CU_LUR_model_plot <- ggplot() +
     breaks = legend_breaks, # Set the breaks for the legend
     labels = c(as.integer(min_cu), "4.54", "8.11", "11.68", "15.25", as.integer(max_cu))
   ) + # Set the limits to the overall range
-  labs(title = expression(paste("(D) Cu models")),   
+  labs(title = expression(paste("(E) Cu models")),   
        color = expression(paste("Cu" ," (ng/m"^3*")"))) +
   xlim(c(st_bbox(amb_bcn)[1], st_bbox(amb_bcn)[3])) +
   ylim(c(st_bbox(amb_bcn)[2], st_bbox(amb_bcn)[4])) +
@@ -2214,14 +2212,160 @@ CU_LUR_model_plot <- ggplot() +
 CU_LUR_model_plot
 
 
+#################################
+### --- CU Hybrid model --- ###
+################################
+CU_HM_model_plot <- ggplot() +
+  geom_sf(data = amb_bcn, fill = "lightgrey", color = "white") +
+  geom_sf(data = hm_estimates_sf, aes(color = cu_hm), size = 1.0, alpha = 1.0) +
+  theme_minimal() +
+  scale_color_viridis_c(
+    option = "viridis", direction = 1, 
+    limits = c(min_cu, max_cu),
+    breaks = legend_breaks, # Set the breaks for the legend
+    labels = c(as.integer(min_cu), "4.54", "8.11", "11.68", "15.25", as.integer(max_cu))
+  ) + # Set the limits to the overall range
+  labs(title = "",  
+       color = expression(paste("Cu", " (ng/m"^3*")"))) +
[…]

### 2023-12-07 02.Code/03.- monitoring_stations.R
monitoring stations script

@@ -1,0 +1,77 @@
+# 03.- monitoring stations 
+pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate, 
+               mice, ggmice, plyr, data.table, mapview, ggmap, omsdata, sf, sp, patchwork,
+               ggpattern, GGally)
+
+###################################
+### --- Barcelona Base Map --- ### 
+#################################
+load("03.Outputs/figures/barcelona_base_map.RData")
+
+####################################
+### --- Monitoring stations --- ###
+###################################
+
+### --- Coordinates --- ### 
+monitoring_coords <- readxl::read_xlsx("01.Data/BiSC_monitoring_stations/monitoring_coordenates.xlsx")
+dplyr::glimpse(monitoring_coords)
+view(monitoring_coords)
+
+names(monitoring_coords) <- c()
+
+
+
+### --- Mapping monitoring stations --- ### 
+  ggmap(barcelona_map2) +
+  geom_point(data = monitoring_coords,
+             aes(x = longitude, y = latitude), 
+             color = '#21918c',
+             alpha = 1, 
+             size = 2) +
+  theme_minimal() +  xlab('') + ylab('') +
+  facet_grid(.~'Monitoring Stations')
+
+
+# Change the names displaye in the map
+new_names <- c("Ciutadella", "Eixample", "Gracia", "Hospitalet", "Obs Fabra",
+                 "Palau Reial", "Poblenou", "Sants", "Vall Hebron", "Ies Goya",
+                 "Pl. Univ", "Zona Univ", "Jardins", "Sagnier", "St. Adria",
+                 "Sta. Coloma")
+  
+monitoring_coords$station <- factor(monitoring_coords$station, levels = unique(monitoring_coords$station), labels = new_names)
+  
+
+# here we assing the label direction by hand 
+label_direction <- c(0.005, -0.005, 0.005, -0.005, 0.005, 
+                     0.005, 0.005, -0.005, 0.005, 0.005, 
+                     0.005, -0.005, 0.005, 0.005, 0.005, 
+                     0.005)
+
+monitoring_coords$label_direction <- label_direction
+
+# Plotting the monitoring stations with the 
+ggmap(barcelona_map2) +
+  geom_point(data = monitoring_coords,
+             aes(x = longitude, y = latitude), 
+             color = 'firebrick',
+             alpha = 1, 
+             size = 2) +
+  geom_label(data = monitoring_coords,
[…]

### 2023-12-07 02.Code/03.- monitoring_stations.R
figure monitoring stations (location + time series)

@@ -17,8 +17,8 @@ monitoring_coords <- readxl::read_xlsx("01.Data/BiSC_monitoring_stations/monitor
 dplyr::glimpse(monitoring_coords)
 view(monitoring_coords)
 
-names(monitoring_coords) <- c()
-
+monitoring_coords <- monitoring_coords %>% 
+                     dplyr::filter(!station == "obs_fabra")
 
 
 ### --- Mapping monitoring stations --- ### 
@@ -33,7 +33,7 @@ names(monitoring_coords) <- c()
 
 
 # Change the names displaye in the map
-new_names <- c("Ciutadella", "Eixample", "Gracia", "Hospitalet", "Obs Fabra",
+new_names <- c("Ciutadella", "Eixample", "Gracia", "Hospitalet",
                  "Palau Reial", "Poblenou", "Sants", "Vall Hebron", "Ies Goya",
                  "Pl. Univ", "Zona Univ", "Jardins", "Sagnier", "St. Adria",
                  "Sta. Coloma")
@@ -42,7 +42,7 @@ monitoring_coords$station <- factor(monitoring_coords$station, levels = unique(m
   
 
 # here we assing the label direction by hand 
-label_direction <- c(0.005, -0.005, 0.005, -0.005, 0.005, 
+label_direction <- c(0.005, -0.005, 0.005, -0.005, 
                      0.005, 0.005, -0.005, 0.005, 0.005, 
                      0.005, -0.005, 0.005, 0.005, 0.005, 
                      0.005)
@@ -50,12 +50,38 @@ label_direction <- c(0.005, -0.005, 0.005, -0.005, 0.005,
 monitoring_coords$label_direction <- label_direction
 
 # Plotting the monitoring stations with the 
-ggmap(barcelona_map2) +
+fig_stations <- ggmap(barcelona_map2) +
+                geom_point(data = monitoring_coords,
+                           aes(x = longitude, y = latitude), 
+                           color = 'firebrick',
+                           alpha = 1, 
+                           size = 2) +
+                geom_label(data = monitoring_coords,
+                           aes(x = longitude, y = latitude, label = station), 
+                           color = "black",
+                           fill = "white",
+                           label.size = 0.2,
+                           label.padding = unit(0.2, "lines"),
+                           hjust = 0.5,
+                           vjust = 0.5,
+                           size = 2.5,
+                           alpha = 0.7,
[…]

### 2023-12-07 02.Code/03.- monitoring_stations.R
Palau Reial monitoring station figure

@@ -116,6 +116,9 @@ monitoring_station_fig
 bcn_stations <- read.csv("01.Data/BiSC_monitoring_stations/imputed/bcn_ms_bisc_new_imputed.csv")
 dplyr::glimpse(bcn_stations)
 
+bcn_stations <- read.csv("01.Data/BiSC_monitoring_stations/imputed/monitoring_stations_imputed.csv")
+dplyr::glimpse(bcn_stations)
+
 bcn_stations$date <- as.Date(bcn_stations$date)
 
 library(tidyr)
@@ -236,5 +239,67 @@ location_ts_stations <- monitoring_station_fig | (no2_ts_plot / pm25_ts_plot)
 ### --- Export figure --- ### 
 ggsave(plot = location_ts_stations , "03.Outputs/figures/location_ts_stations.png",
        dpi = 600, width = 12, height = 10, units = "in")
+
+
+###############################################
+### --- Supplementary materials paper  --- ### 
+#############################################
+
+bcn_stations <- read.csv("01.Data/BiSC_monitoring_stations/imputed/monitoring_stations_imputed.csv")
+dplyr::glimpse(bcn_stations)
+
+bcn_stations$date <- as.Date(bcn_stations$date)
+
+
+### --- NO2 monitoring station --- ###
+no2_pr_ts <- bcn_stations %>% dplyr::select(date, 
+                                            no2_palau_reial) %>% 
+             ggplot(aes(x = date, y = no2_palau_reial)) + 
+             geom_line(color = "firebrick") + 
+             labs(x = "", 
+                  y = expression(NO[2]*" (µg/m" ^ "3" * ")")) + 
+             theme_bw() + 
+             theme(legend.position = "none") +
+             facet_wrap(.~ "Palau Reial")
+
+no2_pr_ts
+
+### --- PM25 monitoring station --- ###
+pm25_pr_ts <- bcn_stations %>% dplyr::select(date, 
+                                            pm25_palau_reial) %>% 
+              ggplot(aes(x = date, y = pm25_palau_reial)) + 
+              geom_line(color = "firebrick") + 
+              labs(x = "", 
+                   y = expression(PM[2.5]*" (µg/m" ^ "3" * ")")) + 
+              theme_bw() + 
+              theme(legend.position = "none") 
+
+pm25_pr_ts
+
+### --- PM25 monitoring station --- ###
+bc_pr_ts <- bcn_stations %>% dplyr::select(date, 
[…]

### 2023-12-20 manuscript_BiSC_outdoor_AP.tex

@@
 Air pollution is the leading environmental risk factor for human health worldwide linked to wide range of adverse health outcomes.
-A detailed and comprehensive exposure assessment of outdoor air pollution considering both spatial and temporal variability is fundamental to conduct epidemiological studies.
+A detailed and comprehensive exposure assessment of outdoor air pollution considering both spatial and temporal variability in urban areas is essential to identify critical windows of susceptibility during pregnancy to traffic-related air pollution.
+Within the framework of the Barcelona Life Study Cohort (BiSC), we developed Land use regression (LUR), dispersion (DM), and hybrid (HM) models to estimate outdoor air pollution exposure at residential addresses during pregnancy for several traffic-related air pollutants; NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn).
+The overall aim is to obtain accurate estimates of traffic-related air pollution, intended to be used in the prenatal exposure assessment for a cohort of pregnant women in the city of Barcelona.
+We create 101 spatial predictor variables
 
@@
 Air pollution stands as the primary environmental risk factor for human health and the fourth most deadly health risk worldwide, playing a significant role in the global burden of disease \cite{cohen2017, he2020state}.
-Exposure to ambient air pollution has been linked with a extensive array of adverse health outcomes \cite{boogaard2022, guxens2022hei, haddad2023}.
-In this context urbanization have been playing a major role increasing air pollution exposure as cities grow and more population moves to urban areas \cite{nieuwenhuijsen2016}.
-This trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas \cite{boogaard2022}.
+Exposure to ambient air pollution has been linked with many adverse health outcomes \cite{boogaard2022, guxens2022hei, haddad2023}.
+In this context, urbanization has been playing a major role in increasing air pollution exposure as cities grow and more population moves to urban areas \cite{nieuwenhuijsen2016}.
+These trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas \cite{boogaard2022}.
 \\
 
-Over the past few decades, several modelling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution \cite{hoek2017, di2019no2, di2019pm25, stafoggia2019, stafoggia2020}.
+Over the past few decades, several modeling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution \cite{hoek2017, di2019no2, di2019pm25, stafoggia2019, stafoggia2020}.
 Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches for estimating air pollution exposure concentrations in epidemiological studies \cite{gulliver2015, dehoogh2014}.
@@
 While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains, such models (multi-country and national domain models) often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
-This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potential lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.
+This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potentially lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.
 \\
 
-Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}.
+Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that make difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}.
 In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees-based models \cite{liu2022}.
@@
 % Put the figure text closer to the Figure 1
-\captionsetup[figure]{skip=-4pt}
+\captionsetup[figure]{skip=0pt}
 % We add the figure with the study domain 
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/fig1_op2.png}
-\caption{Study domain and spatial distribution (BiSC participants, BiSCAPE campaign, monitoring stations, and traffic points)}
+\includegraphics[width=1.0\textwidth]{figures/figure_campaigns.png}
+\caption{Study domain and spatial distribution of the monitoring campaigns}
 \label{fig1}
@@
 BiSC-home, BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\
+\hline
 \multirow{5}{*}{BiSCAPE} 
-& PM\textsubscript{25} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 13.4 & 6.14 & 5.01 & 9.06 & 11.3 & 16.2 & 35.8 \\ 
+& PM\textsubscript{2.5} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 13.4 & 6.14 & 5.01 & 9.06 & 11.3 & 16.2 & 35.8 \\ 
 & BC (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 1.27 & 0.50 & 0.40 & 0.88 & 1.22 & 1.52 & 2 .95 \\ 
-& Fe (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 0.25 & 0.13 & 0.07 & 0.15 & 0.21 & 0.32 & 0.77 \\ 
-& Cu (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
-& Zn (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
+& PM\textsubscript{2.5 - Fe} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 0.25 & 0.13 & 0.07 & 0.15 & 0.21 & 0.32 & 0.77 \\ 
+& PM\textsubscript{2.5 - Cu} (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
+& PM\textsubscript{2.5 - Zn} (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
 \bottomrule
@@
 Figure 1A illustrates a high density coverage in the central area of Barcelona, where the BiSC-home campaign was conducted.
-Figure 1B exhibit a similar pattern of spatial distribution but with a reduced density in sampling locations.
+Figure 1B exhibit a similar pattern of spatial distribution of the monitoring campaign but with a reduced density in sampling locations.
 
@@
 
-Since different approaches were used to built outdoor air pollution models were not directly comparable due to differences on temporal basis.
+Since different approaches and sampling campaigns were used to built outdoor air pollution models were not directly comparable.
 
@@
 
-\textbf{\cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5} and BC concentrations from the LUR, DM, and HM.
+\textbf{\Cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5} and BC concentrations from the LUR, DM, and HM.
 The data shown are long-term estimated concentrations of pollutants, after averaging weekly estimates for all the study period (2019 - 2021).
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/HEX_all.png}
-\caption{Comparison/agreement between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
+\includegraphics[width=1.0\textwidth]{figures/final_HEXBIN.png}
+\caption{Comparison between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, Zn).}}
 \end{figure}

### 2023-12-20 supplementary_BiSC_outdoor_AP.tex

@@
 \usepackage{multirow}
+\usepackage{threeparttable} % For footnotes in tables
+\usepackage{graphicx}  % For including graphics
+\usepackage{caption}   % For customizing captions
+\usepackage{cleveref}  % For enhanced referencing
 
 % Configure for "Fig.":
-\crefname{figure}{Fig.}{Figs.}
-\Crefname{figure}{Fig.}{Figs.}
+\crefname{figure}{figure}{figures}
+\Crefname{figure}{Figure}{Figures}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Predictor variables and sources of information --- %%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Predictor variables and sources of information}
+To develop LUR models, we calculate a set of spatial variables that could hep us to explain pollutant concentration distribution across the study area.
+We calculate a total of 101 predictor variables, see \textbf{\Cref{Table S1}}.
+To calculate the predictor variables we applied a wide range of local and regional GIS datasets, see \textbf{\Cref{Table S2}}.
 
+%%%% ----------------------------------------------------------------------------------%%%%
+%%%% --- Table S1. Predictor variables categories created for LUR model development --- %%%
+%%%% ----------------------------------------------------------------------------------%%%%
+
+\renewcommand{\thetable}{S\arabic{table}}
+\begin{table}[ht]
+\centering
+\caption{Predictor variables categories created for LUR model development.}
+\begin{tabular}{lll}
+\toprule
+\textbf{Traffic} & \textbf{Land-uses} & \textbf{Urban configuration} \\
+\midrule
+Road length & Residential & Population \\
+Major road length & Industry & Buildings density \\
+Distance to road & Natural & Imperviousness \\
+Distance to major road & Urban green & Slope \\
+Traffic intensity & Road density & Buildings height \\
+Traffic intensity only major roads &  & Street bearing \\
+Traffic intensity in nearest road &  & Low Emission Zone (ZBE)\\
+Traffic intensity in nearest major road &  & Street width (BiSCAPE)\\
+                                         &  & Street canyon (BiSCAPE)\\
+                                         &  & Sampler height (BiSCAPE)\\
+\bottomrule
+\label{Table S1}
+\end{tabular}
+\end{table}
+
+\vspace{-2em}
+
+%%%% ------------------------------------------------------%%%%
+%%%% --- Table S2. Data sources for predictor variables --- %%%
+%%%% ------------------------------------------------------%%%%
+\begin{table}[ht]
+\centering
+\caption{Predictor varaibles categories created for LUR model development.}
+\begin{tabular}{lll}
+\toprule
+\textbf{Predictor variable} & \textbf{Year} & \textbf{Source} \\
+\midrule
+Road network and traffic data  & 2014 & Departament de mobilitat, Ajuntament de Barcelona \\
+Land uses & 2018 & Urban atlas, Copernicus \\
+Population & 2016 & Institut d'Estadística de Catalunya \\
+Buildings height & 2009 & Modelo Digital de Superficies Edificación  - MDSn2.5 (MDSNE)\\
+Imperviousness & 2018 & Copernicus \\
+Slope & 2012 & Institut Cartográfic de Catalunya \\
+ZBE & 2012 & OSM\\
+\bottomrule
+\label{Table S2}
+\end{tabular}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Deseasonalizing air pollution data --- %%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\subsection{Deseasonalizing pollution data}
+
+For the development of LUR models data collected at different points of time need to be deseasonalized (process explained below) before being averaged to represent the average value of the baseline period.
+
+\subsubsection{NO\textsubscript{2} deseasonalization}
+
+For the NO\textsubscript{2} data (BiSCAPE + BiSC) the \textit{difference method} was used for deseasonalisation.
+
+\begin{enumerate}
+    \item Define the baseline period, that is, the total period in which the air pollution measurements were done.
+For NO\textsubscript{2} the baseline period was from 23\textsuperscript{rd} October 2018 to 16\textsuperscript{th} February 2022.
+    \item Collect hourly air pollution data from a routine monitoring site covering the baseline period.
+We used data from Palau Reial monitoring station, as this is the reference site for the BiSCAPE campaigns and the BiSC cohort.
+NO\textsubscript{2}, data at Palau Reial was missing for only 5 \% of the period.
+Even though the missing rate is low, we imputed missing data by doing a linear regression with NO\textsubscript{2} levels at the monitoring site of Sants (r = 0.84).
+After recovering those data, the percentage of missing data at Palau Reial decreased to 0.68 \%.
+    \item Calculate the average concentration of NO\textsubscript{2} for the routine monitoring site covering the baseline measurement period: C\textsubscript{PR}(avg).
+    \item Calculate the concentrations of the pollutant at the routine monitoring site for the exact same period (t) as the sample collected at the BiSC and BiSCAPE locations in the different periods: CP\textsubscript{PR}(t).
+    \item Calculate for the routine monitoring site for each sample the difference between the concentration (C\textsubscript{PR}) and the average covering the baseline measurement period: DC\textsubscript{PR}(t) = C\textsubscript{PR}(t) - C\textsubscript{PR}(avg).
+    \item Calculate for each sample at site i (C\textsubscript{i}(t)) the adjusted concentration (C\textsubscript{i, adj}(t)) by computing: C\textsubscript{i,adj}(t) = C\textsubscript{i}/DC\textsubscript{PR}(t).
+\end{enumerate}
+
+\subsubsection{PM\textsubscript{2.5}, PM\textsubscript{2.5} components (Fe, Cu, Zn), and BC deseasonalization}
+
+For the PM\textsubscript{2.5} data (BiSCAPE + BiSC) the \textit{difference method} was used for deseasonalisation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Performance metrics --- %%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
 \subsection{Performance metrics}
-The following metrics were used to evaluate the LUR model for all the air pollution assessed.
+The following metrics were used to evaluate the LUR model for all the air pollutants assessed.
 
@@
 
-% --- mean bias (MB) --- % 
-\begin{equation}
-\text{Mean Bias} = \frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)
-\end{equation}
-
 % --- correlation coefficient (r) --- %
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Model development --- %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Model development}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Model performance --- %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Model performance}
+The performance of the final models obtained for NO\textsubscript{2}, PM\textsubscript{2.5}, BC and PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} is described in \textbf{\Cref{Table S3}}.
+
+%%% --- Table with performance metrics for Hybrid models --- %%%
+\begin{table}[ht]
+\centering
+\begin{threeparttable}
+\caption{Performance metric for LUR models.}
+\label{Table S3}
+\begin{tabular}{llllll}
+\toprule
+\textbf{Pollutant} & \textbf{Year} & \textbf{N}  &  \textbf{R\textsuperscript{2}\textsubscript{10-CV}} & \textbf{RMSE\textsubscript{10-CV}} & \textbf{Predictor variables}\tnote{1} \\
+\midrule
+\multirow{2}{*}{NO\textsubscript{2}} & \multirow{2}{*}{2018-2021} & \multirow{2}{*}{489} & \multirow{2}{*}{0.62} & \multirow{2}{*}{4.03} & trafload25 + sqralt + majorroadlength50 + \\ 
+ & & & & & roadlength25 + majorroadlength300 \\
+\midrule
+PM\textsubscript{2.5} & 2019 & 34 & 0.45 & 1.47 & hdres500 + trafnear + ZBE \\
+\midrule
+BC & 2019 & 30  & 0.83 & 0.18 & hdres50 + linesnear + pop300 + trafload + roads500 \\
+\midrule
+PM\textsubscript{2.5 - Fe} & 2019 & 34 & 0.89 & 0.03 & trafnear + lat + pop + ZBE  \\
+\midrule
+PM\textsubscript{2.5 - Cu} & 2019 & 31 & 0.87 & 0.72 & trafnear + roads1000 + ind1000 + pop25 \\
+\midrule
+PM\textsubscript{2.5 - Zn} & 2019 & 31  & 0.85 & 6.99 & ZBE + roads50 + ind1000 + build25  \\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\small
+\item[1] Predictor variables (roadlength: total road length (m), majorroadlength: total major road length (m), trafload: total traffic intensity (veh/day), trafnear: traffic intensity at the nearest road (veh/day), linesnear: number traffic lines on nearest street, ZBE: Low Emissions Zone (Yes/No, ref value=No), hdres: high-density residential area (m2), roads: roads surface area (m2), ind: industry area (m2), pop: population density (inhabitants), build: building area (m2), lat: latitude (m), sqralt: squared root altitude (m)).
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+The best models were selected based on its better validation performance and on the availability of the sampling campaign that could better represent the study population.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Back-extrapolation procedure --- %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\subsection{Back-extrapolation procedure}
+
+In the BiSC cohort study, we aim to study the effect of air pollution exposure using different windows of exposure (i.e., entire pregnancy, trimesters, and weeks of pregnancy).
+For each subject, back-extrapolated concentrations for different time periods can be calculated using daily concentrations from the background monitoring station.
+The station used was Palau Reial, the spatial location and the daily temporal concentration variability are shown for \textit{NO\textsubscript{2}}, \textit{PM\textsubscript{2.5}}, \textit{BC} in \textbf{\Cref{FigureS1}}.
+
+%%% --- Figure. Monitoring stations and time-series used for LUR Back-Exptrapolation --- %%%
+
+% Redefine the figure numbering to include 'S' for Supplementary
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+%%% Figure: Monitoring stations and time-series used for LUR Back-Extrapolation %%%
+\captionsetup[figure]{skip=6pt} % Adjusting the space between the figure and its caption
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/location_ts_stations.png} % Adjust the path to your image
+    \caption{Spatial location and daily temporal concentrations at Palau Reial station.}
+    \label{FigureS1} % Label for referencing
+\end{figure}
 
@@
 \subsection{Performance metrics}
+The following metrics were used to evaluate the HM for all the air pollutants assessed.
+
+% --- R2 --- %
+\begin{equation}
+R^2 = \left( \frac{\sum (\hat{y}_i - \bar{y})^2}{\sum (y_i - \bar{y})^2} \right)
+\end{equation}
+
+% --- root mean squared error (RMSE) --- % 
+\begin{equation}
+\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)^2}
+\end{equation}
+
+% --- correlation coefficient (r) --- %
+\begin{equation}
+r = \frac{\sum (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum (x_i - \bar{x})^2 \sum (y_i - \bar{y})^2}}
+\end{equation}
+\vspace{0.5 cm}
 
@@
 \begin{table}[ht]
+\caption{Performance metric for hybrid models.}
 \centering
-\caption{Performance metric for hybrid models.}
+\begin{threeparttable}
 \begin{tabular}{llllll}
 \toprule
-\textbf{Dataset campaign} & \textbf{Pollutant} & \textbf{N} & \textbf{R\textsuperscript{2} (10-CV)} & \textbf{RMSE (10-CV)} \\
+\textbf{Pollutant} & \textbf{Year} & \textbf{N} & \textbf{R\textsuperscript{2}\textsubscript{10-CV}} & \textbf{RMSE\textsubscript{10-CV}} & \textbf{Predictors}\tnote{2} \\
 \midrule
-BiSC-home & NO\textsubscript{2} & 1232 & 0.64 & 7.48 \\
-\multirow{5}{*}{BiSCAPE} & PM\textsubscript{2.5} & 161 & 0.66 & 3.45 \\
-& BC & 74 & 0.86 &  0.23 \\
-& Fe & ... & 0.54 & ... \\
-& Cu & ... & 0.70 & ... \\
-& Zn & ... & 0.42 & ... \\
+\multirow{6}{*}{NO\textsubscript{2}} & \multirow{6}{*}{2018-2021} & \multirow{6}{*}{1232} & \multirow{6}{*}{0.64} & \multirow{6}{*}{7.48} & no2\_dispersion +  idw\_no2\_monitoring\_station + \\
+ & & & & & distinvmajor1 + majorroadlength100 + \\
+ & & & & & majorroadlength50 + trafload25 + trafnear + \\ 
+ & & & & & avg\_traffic\_stations + avg\_solar\_radiation +\\ 
+ & & & & & sqralt + ldres1000 +ldres500 + hdres1000 +  \\
+ & & & & & build1000 + roads100 + lat + lon \\
+\midrule
+\multirow{4}{*}{PM\textsubscript{2.5}} & \multirow{4}{*}{2021} & \multirow{4}{*}{161} & \multirow{4}{*}{0.66} & \multirow{4}{*}{3.45} & pm25\_dispersion + idw\_pm25monitoring\_stations + \\
+& & & & & avg\_traffic\_stations + ldres1000 + build25 +  pop1000 + \\
+& & & & & roads25 + sqralt +  majorroadlength25 + \\
+& & & & & trafload300 + build\_height\_25 + lat + lon \\
+\midrule
+BC & 2021 & 74 & 0.86 &  0.23 & bc\_Dispersion + idw\_nox\_monitoring\_stations +
+ avg\_atmospheric\_pressure + avg\_bc\_palau\_reial +  pop100 + hdres300 + roads25 + avg\_wind\_speed + roadlength25 +  hdres50 + distinvmajor1 + avg\_traffic\_stations +  linesnear +  pop300 + 
+ trafload500 + roads500
+ \\
+\midrule
+PM\textsubscript{2.5 - Fe} & 2021 & 161 & 0.54 & 0.08 & pm25\_road\_nonexhaust +idw\_pm25monitoring\_stations + roads300 + roads1000 + ldres500 + sqralt + zbe + lat + lon \\
+\midrule
+PM\textsubscript{2.5 - Cu} & 2021 & 154 & 0.70 & 2.15 & pm25\_road\_nonexhaust + idw\_pm25monitoring\_stations +  linesnear + industry1000 + roads300 + ldres1000 + roads1000 + avg\_traffic\_stations + zbe + lat + lon \\
+\midrule
+PM\textsubscript{2.5 - Zn} & 2021 & 141 & 0.44 & 27.7 & pm25\_road\_nonexhaust + pm25\_background + idw\_pm25monitoring\_stations + linesnear + industry1000 + roads50 + build25 + roads300 + ldres1000 + roads1000 + avg\_traffic\_stations +  trafmajor +trafmajorload1000 +trafload1000 + hdres25 + zbe + lat + lon \\
 \bottomrule
 \end{tabular}
+\begin{tablenotes}
+\small
+\item [2] Predictor variables (no2\_Dispersion: dispersion estimates (ug/m$^3$), pm25\_Dispersion: dispersion estimates PM2.5 (ug/m$^3$), BC\_Dispersion: dispersion estimates BC (ug/m$^3$), PM25\_road\_nonexhaust: dispersion estimates for nonexhaust PM2.5 (ug/m$^3$), PM25\_background: dispersion estimates for background PM2.5 (ug/m$^3$), idw\_no2\_monitoring\_station: weekly NO$_2$ inverse distance weighting interpolation estimates from XVCPA (ug/m$^3$), idw\_pm25\_monitoring\_station: weekly PM2.5 inverse distance weighting interpolation estimates from XVCPA (ug/m$^3$), idw\_nox\_monitoring\_station: weekly NOx inverse distance weighting interpolation estimates from XVCPA (ug/m$^3$), avg\_bc\_palau\_reial: weekly BC average concentration from Palau Reial monitoring station (ug/m$^3$), avg\_traffic\_stations: weekly average count of vehicles in AMB (count), avg\_solar\_radiation: weekly average solar radiation from Raval station (MJ/m$^2$), avg\_wind\_speed: weekly average wind speed from Raval station (km/h), avg\_atmospheric\_pressure: atmospheric pressure from Raval station (hPA), roadlength: total road length (m), majorroadlength: total major road length (m), trafload: total traffic intensity (veh/day), trafnear: traffic intensity at the nearest road (veh/day), linesnear: number traffic lines on nearest street, zbe: Low Emissions Zone (Yes/No, ref value=No), hdres: high-density residential area (m$^2$), roads: roads surface area (m$^2$), ind: industry area (m$^2$), pop: population density (inhabitants), build: building area (m$^2$), ldress: low density residential area (m$^2$), industry: industrial area (m$^2$), trafmajor: total major roads traffic intensity (veh/day), build\_height: averaged building height (meters), lat: latitude (m), sqralt: squared root altitude (m$^{1/2}$)).
+\end{tablenotes}
 \label{tablesX}
+\end{threeparttable}
 \end{table}
 
-\subsection{Variable importance}
+\newpage
+\subsection{Variable importance Random Forest}
 
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/fig_HM_allmodels_importance.png}
-\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
+\includegraphics[width=1.0\textwidth]{figures/fig_HM_allmodels_importance_v2.png}
+\caption{Variable importance for HM for \textit{NO\textsubscript{2}}, \textit{PM\textsubscript{2.5}}, \textit{BC}, \textit{PM\textsubscript{2.5 - Fe}}, \textit{PM\textsubscript{2.5 - Cu}} and \textit{PM\textsubscript{2.5 - Zn}.}}
 \label{fig2}
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/fig_HM_test_all_models.png}
-\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
+\includegraphics[width=1.0\textwidth]{figures/fig_HM_test_all_models_v2.png}
+\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{PM\textsubscript{2.5 - Fe}}, \textit{PM\textsubscript{2.5 - Cu}}, and \textit{PM\textsubscript{2.5 - Zn}}.}
 \label{fig2}

### 2024-01-03 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Material and methods--- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Material and methods--- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \midrule
-BiSC-home, BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\
+BiSC-home + BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\
 \hline
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/boxplot_all_models_estimates.png}
+\includegraphics[width=0.9\textwidth]{figures/boxplot_all_models_estimates.png}
 \caption{Distribution of exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/all_maps.png}
+\includegraphics[width=0.9\textwidth]{figures/combined_figure1.png}
 \caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
+\end{figure}
+
+\captionsetup[figure]{skip=6pt}
+% We add the figure with the study domain 
+\begin{figure}[!htb]
+\includegraphics[width=0.9\textwidth]{figures/combined_figure2.png}
+\caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and hm(\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
 \end{figure}
@@
 \newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We successfully developed spatiotemporal models using three different approaches (LUR, DM, and HM). Models show moderate to good performance 
-
-spatiotemporal models using three different approaches (LUR, DM and HM).
-All models show moderate to good performance evaluated by the defined metrics in the \textbf{methods section} and displayed in more detail in the \textbf{Supplementary materials}.
-
+evaluated by the defined metrics previously and displayed in more detail in the \textbf{supplementary materials}.
 As we expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed.

### 2024-01-03 supplementary_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
     \item Calculate the concentrations of the pollutant at the routine monitoring site for the exact same period (t) as the sample collected at the BiSC and BiSCAPE locations in the different periods: CP\textsubscript{PR}(t).
-    \item Calculate for the routine monitoring site for each sample the difference between the concentration (C\textsubscript{PR}) and the average covering the baseline measurement period: DC\textsubscript{PR}(t) = C\textsubscript{PR}(t) - C\textsubscript{PR}(avg).
-    \item Calculate for each sample at site i (C\textsubscript{i}(t)) the adjusted concentration (C\textsubscript{i, adj}(t)) by computing: C\textsubscript{i,adj}(t) = C\textsubscript{i}/DC\textsubscript{PR}(t).
+    \item Calculate for the routine monitoring site for each sample the difference between the concentration (C\textsubscript{PR}) and the average covering the baseline measurement period: D\_C\textsubscript{PR}(t) = C\textsubscript{PR}(t) - C\textsubscript{PR}(avg).
+    \item Calculate for each sample at site i (C\textsubscript{i}(t)) the adjusted concentration (C\textsubscript{i, adj}(t)) by computing: C\textsubscript{i,adj}(t) = C\textsubscript{i}/D\_C\textsubscript{PR}(t).
 \end{enumerate}
 
-\subsubsection{PM\textsubscript{2.5}, PM\textsubscript{2.5} components (Fe, Cu, Zn), and BC deseasonalization}
+\subsubsection{PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} deseasonalization}
 
-For the PM\textsubscript{2.5} data (BiSCAPE + BiSC) the \textit{difference method} was used for deseasonalisation.
+For the PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} the data available was obtained during the BiSCAPE campaigns.
+The \textit{ratio method} was used for deseasoanlisation.
+The method procedure is very similar to the difference method explained.
+The rationale for using a different method for these pollutants is that we aim to create LUR models for some PM\textsubscript{2.5} components.
+Since we do not have hourly data from the PM\textsubscript{2.5} components for the baseline period (data is only available as a daily average every four days), the deseasonalisation of these components needs to be done with total PM\textsubscript{2.5} conctrations.
+As the \textit{ratio method} is based on subtraction, this method cannot be applied to different pollutants than the reference.
+Samples collected within the BiSCAPE campaigns at Palau Reial showed lower concentrations than those collected online at Palau Reial.
+Using the ratio method in PM\textsubscript{2.5} concentrations may lead to over-subtracting.
+
+\begin{enumerate}
+  \item Define the baseline period, that is, the total period in which the air pollution measurements were done.
+For PM\textsubscript{2.5} and its chemical components and BC the baseline period is from 01/01/2021 to 16/02/2022.
+  \item Collect hourly air pollution data from a routine monitoring site covering the baseline period.
+We used data from Palau Reial monitoring site, as this is the reference site for the BiSC cohort.
+  \begin{enumerate}
+      \item PM\textsubscript{2.5} data at Palau Reial was missing for 26.8\% of the period due instrument maintenance.
+We imputed missing data by doing a linear regression with collocated BC concentrations (r=0.57) as no other station in Barcelona had hourly data for PM\textsubscript{2.5} during this period.
+After recovering those data, the percentage of missing data at Palau Reial decreased to 1.0\%.
+      \item BC data at Palau Reial was missing for only 2.3\% of the period
+  \end{enumerate}
+ \item Calculate the average concentration of the corresponding pollutant (j) for the routine monitoring site covering the baseline measurement period: C\textsubscript{j,PR}(avg).
+For PM\textsubscript{2.5} components, in this step j corresponds to total PM\textsubscript{2.5} mass concentrations.
+ \item Calculate the concentrations of the pollutant j at the routine monitoring site for the exact same period (t) as the sample collected at BiSCAPE locations in the different periods: C\textsubscript{j,PR}(t).
+For PM\textsubscript{2.5} components, in this step j corresponds to total PM\textsubscript{2.5} mass concentrations.
+ \item Calculate the routine monitoring site for each sample the ratio between the concentration of pollutant j (C\textsubscript{j, PR}(t)) and the average covering the baseline measurement period: R\_C\textsubscript{j,PR} = C\textsubscript{PR}(t) / C\textsubscript{PR}(avg).
+For PM\textsubscript{2.5} components, in this step j correspond to total PM\textsubscript{2.5} mass concentrations.
+ \item Calculate for each sample at site i (C\textsubscript{i}(t)) the adjusted concentration of pollutant j (C\textsubscript{i,adj}(t)) by computing: C\textsubscript{j,i,adj}(t) = c\textsubscript{j,i}(t)/R\_C\textsubscript{j,PR}(t)
+\end{enumerate}
+
+\subsubsection{Averages for LUR model development}
+Once the measurements done at different time points (t) and sites (i) for the different pollutants (j) were deseasonalised, we proceeded to calculate the average of the adjusted cocnentrations for each site i.
+Therefore, the average value obtained for each site was considered to be the average for the baseline period at site i.
+These averages were the input values for the development of the LUR model.
 
@@
 \subsection{Model development}
+
+Following the ESCAPE guidelines
 
@@
 
+\subsubsection{Spatial variables}
+
+%%% Figure: Monitoring stations and time-series used for LUR Back-Extrapolation %%%
+\captionsetup[figure]{skip=6pt} % Adjusting the space between the figure and its caption
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.6\textwidth]{figures/lur_pred_corr.png} % Adjust the path to your image
+    \caption{Spatial location and daily temporal concentrations at Palau Reial station.}
+    \label{FigureS1} % Label for referencing
+\end{figure}
+
+\subsubsection{Air pollution monitoring stations (XVPCA stations)}
+%%% Figure: Monitoring stations and time-series used for LUR Back-Extrapolation %%%
+\captionsetup[figure]{skip=6pt} % Adjusting the space between the figure and its caption
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/location_ts_stations_hm.png} % Adjust the path to your image
+    \caption{Spatial location and daily temporal concentrations at Palau Reial station.}
+    \label{FigureS1} % Label for referencing
+\end{figure}
+
+\subsubsection{Traffic count measurement and meteorological data}
+While our dispersion model initiall
+
+\subsubsection{Dispersion model estimates}
+
 \subsection{Performance metrics}
@@
 
+\clearpage
+\begin{samepage}
 \subsection{Model performance}
+The performance metrics of the final HM for NO\textsubscript{2}, PM\textsubscript{2.5}, BC and PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} is described in \textbf{\Cref{Table S4}}
 
 %%% --- Table with performance metrics for Hybrid models --- %%%
-\begin{table}[ht]
+\begin{table}[!htbp]
 \caption{Performance metric for hybrid models.}
+\label{Table S4}
 \centering
 \begin{threeparttable}
+\small % Making the table text smaller
 \begin{tabular}{llllll}
@@
  & & & & & majorroadlength50 + trafload25 + trafnear + \\ 
- & & & & & avg\_traffic\_stations + avg\_solar\_radiation +\\ 
+ & & & & & avg\_traffic\_stations + avg\_solar\_radiation + \\ 
  & & & & & sqralt + ldres1000 +ldres500 + hdres1000 + \\
@@
 \midrule
-\multirow{4}{*}{PM\textsubscript{2.5}} & \multirow{4}{*}{2021} & \multirow{4}{*}{161} & \multirow{4}{*}{0.66} & \multirow{4}{*}{3.45} & pm25\_dispersion + idw\_pm25monitoring\_stations + \\
-& & & & & avg\_traffic\_stations + ldres1000 + build25 +  pop1000 + \\
-& & & & & roads25 + sqralt +  majorroadlength25 + \\
-& & & & & trafload300 + build\_height\_25 + lat + lon \\
+\multirow{5}{*}{PM\textsubscript{2.5}} & \multirow{5}{*}{2021} & \multirow{5}{*}{161} & \multirow{5}{*}{0.66} & \multirow{5}{*}{3.45} & pm25\_dispersion + idw\_pm25monitoring\_stations + \\
+& & & & & avg\_traffic\_stations + ldres1000 + build25 + \\
+& & & & & pop1000 + roads25 + sqralt  + \\
+& & & & &  majorroadlength25 + trafload300 + \\
+& & & & & build\_height\_25 + lat + lon \\
 \midrule
-BC & 2021 & 74 & 0.86 &  0.23 & bc\_Dispersion + idw\_nox\_monitoring\_stations +
- avg\_atmospheric\_pressure + avg\_bc\_palau\_reial +  pop100 + hdres300 + roads25 + avg\_wind\_speed + roadlength25 +  hdres50 + distinvmajor1 + avg\_traffic\_stations +  linesnear +  pop300 + 
- trafload500 + roads500
- \\
+\multirow{6}{*}{BC} & \multirow{6}{*}{2021} &\multirow{6}{*}{74} & \multirow{6}{*}{0.86} & \multirow{6}{*}{0.23} & bc\_dispersion + idw\_nox\_monitoring\_stations +\\ 
+& & & & & avg\_atmospheric\_pressure + avg\_bc\_palau\_reial +   \\ 
+& & & & &  pop100 + hdres300 + roads25 + avg\_wind\_speed +  \\
+& & & & & roadlength25 + hdres50 + distinvmajor1 +  \\
+& & & & & avg\_traffic\_stations + linesnear + \\
+& & & & & pop300 + trafload500 + roads500 \\
 \midrule
-PM\textsubscript{2.5 - Fe} & 2021 & 161 & 0.54 & 0.08 & pm25\_road\_nonexhaust +idw\_pm25monitoring\_stations + roads300 + roads1000 + ldres500 + sqralt + zbe + lat + lon \\
+\multirow{4}{*}{PM\textsubscript{2.5 - Fe}} & \multirow{4}{*}{2021} & \multirow{4}{*}{161}  & \multirow{4}{*}{0.54} & \multirow{4}{*}{0.08} & pm25\_road\_nonexhaust + \\ 
+& & & & & idw\_pm25\_monitoring\_stations + \\
+& & & & & roads300 + roads1000 + ldres500 + \\
+& & & & & sqralt + zbe + lat + lon \\
 \midrule
-PM\textsubscript{2.5 - Cu} & 2021 & 154 & 0.70 & 2.15 & pm25\_road\_nonexhaust + idw\_pm25monitoring\_stations +  linesnear + industry1000 + roads300 + ldres1000 + roads1000 + avg\_traffic\_stations + zbe + lat + lon \\
+\multirow{5}{*}{PM\textsubscript{2.5 - Cu}} & \multirow{5}{*}{2021} & \multirow{5}{*}{154} & \multirow{5}{*}{0.70} & \multirow{5}{*}{2.15} & pm25\_road\_nonexhaust + \\
+& & & & & idw\_pm25\_monitoring\_stations + \\
+& & & & & linesnear + industry1000 + roads300 + \\
+& & & & & ldres1000 + roads1000 + avg\_traffic\_stations + \\
+& & & & &  zbe + lat + lon \\
 \midrule
-PM\textsubscript{2.5 - Zn} & 2021 & 141 & 0.44 & 27.7 & pm25\_road\_nonexhaust + pm25\_background + idw\_pm25monitoring\_stations + linesnear + industry1000 + roads50 + build25 + roads300 + ldres1000 + roads1000 + avg\_traffic\_stations +  trafmajor +trafmajorload1000 +trafload1000 + hdres25 + zbe + lat + lon \\
+\multirow{6}{*}{PM\textsubscript{2.5 - Zn}} &\multirow{6}{*}{2021} & \multirow{6}{*}{141} & \multirow{6}{*}{0.44} & \multirow{6}{*}{27.7} & pm25\_road\_nonexhaust + pm25\_background + \\
+& & & & & idw\_pm25monitoring\_stations + linesnear + \\
+& & & & & industry1000 + roads50 + build25 + roads300 + \\
+& & & & & ldres1000 + roads1000 + avg\_traffic\_stations + \\
+& & & & & trafmajor + trafmajorload1000 +  trafload1000 +\\
+& & & & & hdres25 + zbe + lat + lon \\
 \bottomrule
@@
 \begin{tablenotes}
-\small
+\scriptsize
 \item [2] Predictor variables (no2\_Dispersion: dispersion estimates (ug/m$^3$), pm25\_Dispersion: dispersion estimates PM2.5 (ug/m$^3$), BC\_Dispersion: dispersion estimates BC (ug/m$^3$), PM25\_road\_nonexhaust: dispersion estimates for nonexhaust PM2.5 (ug/m$^3$), PM25\_background: dispersion estimates for background PM2.5 (ug/m$^3$), idw\_no2\_monitoring\_station: weekly NO$_2$ inverse distance weighting interpolation estimates from XVCPA (ug/m$^3$), idw\_pm25\_monitoring\_station: weekly PM2.5 inverse distance weighting interpolation estimates from XVCPA (ug/m$^3$), idw\_nox\_monitoring\_station: weekly NOx inverse distance weighting interpolation estimates from XVCPA (ug/m$^3$), avg\_bc\_palau\_reial: weekly BC average concentration from Palau Reial monitoring station (ug/m$^3$), avg\_traffic\_stations: weekly average count of vehicles in AMB (count), avg\_solar\_radiation: weekly average solar radiation from Raval station (MJ/m$^2$), avg\_wind\_speed: weekly average wind speed from Raval station (km/h), avg\_atmospheric\_pressure: atmospheric pressure from Raval station (hPA), roadlength: total road length (m), majorroadlength: total major road length (m), trafload: total traffic intensity (veh/day), trafnear: traffic intensity at the nearest road (veh/day), linesnear: number traffic lines on nearest street, zbe: Low Emissions Zone (Yes/No, ref value=No), hdres: high-density residential area (m$^2$), roads: roads surface area (m$^2$), ind: industry area (m$^2$), pop: population density (inhabitants), build: building area (m$^2$), ldress: low density residential area (m$^2$), industry: industrial area (m$^2$), trafmajor: total major roads traffic intensity (veh/day), build\_height: averaged building height (meters), lat: latitude (m), sqralt: squared root altitude (m$^{1/2}$)).
@@
 \end{table}
+\end{samepage}
+\clearpage
 

### 2024-01-03 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 
-\textbf{\Cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5} and BC concentrations from the LUR, DM, and HM.
-The data shown are long-term estimated concentrations of pollutants, after averaging weekly estimates for all the study period (2019 - 2021).
+\textbf{\Cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM.
+The data shown are long-term estimated concentrations of pollutants during the entire pregnancy period, after averaging weekly estimates for all the study period (2019 - 2021).
 

### 2024-01-18 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 \usepackage{graphicx}
-\usepackage{authblk} % this is for multiple autho  rs 
+\usepackage{authblk} % this is for multiple authors 
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 \author[1, 2, 3]{Xavier Basagaña}
-\author[1, 3]{Ioar Rivas}
+\author[1, 2, 3]{Ioar Rivas}
 
@@
 
-Air pollution is the leading environmental risk factor for human health worldwide linked to wide range of adverse health outcomes.
-A detailed and comprehensive exposure assessment of outdoor air pollution considering both spatial and temporal variability in urban areas is essential to identify critical windows of susceptibility during pregnancy to traffic-related air pollution.
+Air pollution is the leading environmental risk factor for human health worldwide linked to a wide range of adverse health outcomes.
+A detailed and comprehensive assessment of outdoor air pollution exposure considering both spatial and temporal variability in urban areas is essential to identify critical windows of susceptibility during pregnancy to traffic-related air pollution.
 Within the framework of the Barcelona Life Study Cohort (BiSC), we developed Land use regression (LUR), dispersion (DM), and hybrid (HM) models to estimate outdoor air pollution exposure at residential addresses during pregnancy for several traffic-related air pollutants; NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn).
-The overall aim is to obtain accurate estimates of traffic-related air pollution, intended to be used in the prenatal exposure assessment for a cohort of pregnant women in the city of Barcelona.
+The overall aim is to obtain accurate weekly estimates of traffic-related air pollution, intended to be used in the prenatal exposure assessment for a cohort of pregnant women in the city of Barcelona.
 We create 101 spatial predictor variables
@@
 
-While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains, such models (multi-country and national domain models) often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
+While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains (multi-country and national domain models), such models often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
 This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potentially lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.
@@
 
-Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that make difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}.
+Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that make difficult an accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}.
 In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees-based models \cite{liu2022}.
-This family of models is often used due to its ability to model potential, nonlinear interaction between predictor variables \cite{liu2022treebased}.
+This family of models is often used due to its ability to model potential nonlinear interaction between predictor variables \cite{liu2022treebased}.
 Among the different ML algorithms, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.
 Furthermore, the COVID-19 pandemic has introduced unprecedented challenges for air pollution modelling and exposure assessment.
-These challenges encompass changes in air pollution patterns, constraints on data collection due to health and safety restrictions (lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}.
+These challenges encompass changes in air pollution patterns, constraints on data collection due to health and safety restrictions (e.g.
+lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}.
 \\
 
-In this study, within the framework of the Barcelona Life Study Cohort (BiSC), we developed LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses.
-LUR and DM, were constructed using conventional techniques, while HM was designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
+This study is within the framework of the Barcelona Life Study Cohort (BiSC), a pregnancy cohort that enrolled 1,080 pregnant women at their first trimester with the aim of evaluating the impact of the exposure to air pollution during pregnancy and fetal and early life development of their offspring.
+We developed LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses of the BiSC participants during their pregnancies.
+LUR and DM were constructed using conventional techniques, while HM was designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
 The overall aim is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates, intended to be used in the prenatal exposure assessment within the BiSC cohort.
@@
 The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021.
-The second campaign was carried out in accordance with the ESCAPE guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
+The second campaign was carried out in accordance with the ESCAPE project guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
 During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and HM.
@@
 LUR models for NO\textsubscript{2} were developed using only data from the locations that had measured home-outdoor (BiSC-home campaign) levels for both sampling campaigns (i.e., we did not use data from participants who had data available for only one campaign, either first or third trimester).
-Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites were during the pandemic period and its resulting restrictions.
+Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites were during and after the pandemic period and its resulting restrictions.
 For PM\textsubscript{2.5}, BC  and PM\textsubscript{2.5} constituents models were developed using BiSCAPE data from the 34 sites.
@@
 An emission inventory was compiled for the period 2018-2021, which included data from diverse emission sources such as vehicular traffic, household activities, industrial plants, airports, harbors, factories, and the commercial sector.
-The Atmospheric Dispersion Modelling System software (ADMS-Urban) from Cambridge Environmental Research Consultants (CERC) was employed to calculate the concentration of pollutants like NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC originating from these sources.
+The Atmospheric Dispersion Modelling System software (ADMS-Urban) from Cambridge Environmental Research Consultants (CERC) was employed to calculate the concentration of pollutants like NO\textsubscript{2}, PM\textsubscript{2.5}, and BC originating from these sources.
 ADMS-Urban is a specialized Gaussian model designed for analyzing pollutant emissions in urban and metropolitan areas, particularly at road level \cite{mchugh1997adms}.
@@
 
-In addition, bias-adjustment techniques have been used to fit the model data to the observations.
-Through hourly bias spatial interpolation and generation of hourly maps.
+In addition, bias-adjustment techniques have been used to fit the model data to the observations through hourly bias spatial interpolation and generation of hourly maps.
 After that, data were processed  by cluster analysis techniques to obtain a mean cluster map.
@@
 For a more comprehensive approach, we integrated additional data from the Catalan Air Pollution Monitoring and Forecasting Network (XVPCA) for Barcelona and surrounding areas \cite{xarxa2012}.
-This data was interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
+These data was interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
 The data came from 16 monitoring stations, including Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St.
@@
 While our dispersion model initially utilized meteorological and traffic data, the hybrid model broadened its scope by integrating daily traffic counts from 180 locations throughout the city \cite{trafficbcn}.
-This broader data helps us better understand the spatial and time-specific variations in urban air pollution \cite{pinto2020}.
+These broader data helps us better understand the spatial and time-specific variations in urban air pollution \cite{pinto2020}.
 To further refine our insights into the influence of weather patterns on pollution, we sourced additional meteorological variables, such as wind speed and air temperature, from the Raval monitoring station managed by Servei Meteorològic de Catalunya \cite{xema2013}.
@@
 For each iteration, the model was trained on k-1 folds and then validated using the leftover fold \cite{ziegel2003}.
-For every fold, we computed metrics such as \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)).
+For every fold, we computed metrics such as the coefficient of determination \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)).
 After each iteration, both the error and adjusted \(R\textsuperscript{2}\) were recorded.

### 2024-01-18 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 The overall aim is to obtain accurate weekly estimates of traffic-related air pollution, intended to be used in the prenatal exposure assessment for a cohort of pregnant women in the city of Barcelona.
-We create 101 spatial predictor variables
+Two separate monitoring campaigns were conducted: the "BiSC-home" campaign took place at the home addresses of BiSC participants and involved measuring only NO2 between 2018-2021.
+Additionally, a second campaign was conducted, named "BiSCAPE", and included four distinct measurement periods between 2021-2022.
+LUR and DM were constructed using conventional state-of-the-art techniques, while HM was designed by harnessing Random Forest algorithm, integrating outputs from DM with explanatory variables from the LUR model, monitoring stations and detailed traffic counts.
+Model performance was evaluated using a cross-validation procedure for LUR and HM, and by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations.
+The models demonstrated moderate to good performance for LUR, DM and HM.
+HM outperformed LUR and DM models for NO\textsubscript{2} (R\textsuperscript{2} = 0.64), PM\textsubscript{2.5} (R\textsuperscript{2} = 0.66), and BC (R\textsuperscript{2} = 0.86).
+LUR models show better performance for PM\textsubscript{2.5 - Fe} (R\textsuperscript{2} = 0.89), PM\textsubscript{2.5 - Cu} (R\textsuperscript{2} = 0.87), PM\textsubscript{2.5 - Zn} (R\textsubscript{2} = 0.85).
+Correlations between the air pollutant estimates for the different models were moderate to strong for both long-and-short term exposures.
+Long-term exposure (entire pregnancy) at the residential addresses exhibited correlations ranging from rs = 0.48 to 0.92, while for short-term exposure (weekly exposure) varied from rs = 0.37 to 0.90.
+exposure.
+Our work emphasizes the importance of employing diverse exposure assessment methods to evaluate the exposure to traffic-related air pollutants and to disentangle differences in effect estimates in the BiSC cohort.
+The exposure estimates from these different models can be used in the BiSC cohort to potentially identify critical windows of susceptibility to air pollution during pregnancy.
 
 \end{abstract}
+
+\newpage
 
@@
 In this context, urbanization has been playing a major role in increasing air pollution exposure as cities grow and more population moves to urban areas \cite{nieuwenhuijsen2016}.
-These trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas \cite{boogaard2022}.
-\\
+These trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas during critical developmental stages \cite{boogaard2022}.
 
 Over the past few decades, several modeling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution \cite{hoek2017, di2019no2, di2019pm25, stafoggia2019, stafoggia2020}.
-Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches for estimating air pollution exposure concentrations in epidemiological studies \cite{gulliver2015, dehoogh2014}.
+Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches \cite{gulliver2015, dehoogh2014}.
 LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that work as proxies for traffic-related air pollution \cite{briggs1997, hoek2008}.
@@
 To overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}.
-\\
 
 While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains (multi-country and national domain models), such models often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
-This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potentially lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.
-\\
+This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to capture the differences in exposure for the population living in relatively small areas (i.e.
+city compared to country-level) to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potentially lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.
 
@@
 For a more comprehensive approach, we integrated additional data from the Catalan Air Pollution Monitoring and Forecasting Network (XVPCA) for Barcelona and surrounding areas \cite{xarxa2012}.
-These data was interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
+The data was interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
 The data came from 16 monitoring stations, including Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St.

### 2024-01-20 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 exposure.
-Our work emphasizes the importance of employing diverse exposure assessment methods to evaluate the exposure to traffic-related air pollutants and to disentangle differences in effect estimates in the BiSC cohort.
+Our work emphasizes the importance of employing diverse exposure assessment methods to evaluate the exposure to traffic-related air pollutants and to disentangle differences in effect estimates during pregnancy.
 The exposure estimates from these different models can be used in the BiSC cohort to potentially identify critical windows of susceptibility to air pollution during pregnancy.
@@
 
-While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains (multi-country and national domain models), such models often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
+While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains (multi-country and national domain models), these models often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
 This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to capture the differences in exposure for the population living in relatively small areas (i.e.
@@
 lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}.
-\\
 
-This study is within the framework of the Barcelona Life Study Cohort (BiSC), a pregnancy cohort that enrolled 1,080 pregnant women at their first trimester with the aim of evaluating the impact of the exposure to air pollution during pregnancy and fetal and early life development of their offspring.
+This study was developed within the framework of the Barcelona Life Study Cohort (\href{https://projectebisc.org/en/home/}{BiSC}), a pregnancy cohort that enrolled 1,080 pregnant women in their first trimester to evaluate the impact of exposure to air pollution during pregnancy, fetal and early life development of their offspring.
 We developed LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses of the BiSC participants during their pregnancies.
 LUR and DM were constructed using conventional techniques, while HM was designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
-The overall aim is to obtain accurate NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents (Fe, Zn, Cu, Sb), and BC estimates, intended to be used in the prenatal exposure assessment within the BiSC cohort.
+The overall aim is to obtain accurate weekly estimates for NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn), intended to be used in the prenatal exposure assessment for the BiSC cohort.
 
@@
 Barcelona (\textbf{\cref{fig1}}), is one of the most populated and dense cities in Spain, with 1,656,725 inhabitants ($\approx16,255.4$ inhabitants/km\textsuperscript{2}).
-It is located on the Mediterranean coast of the Iberian peninsula, protected by the Collserola mountain range and delimited by two delta rivers (Besòs and LLobregat).
+The Barcelona Metropolitan Area (AMB) is located on the Mediterranean coast of the Iberian peninsula, protected by the Collserola mountain range and delimited by two delta rivers (Besòs and LLobregat).
 The city has a Mediterranean climate, the wind patterns are dominated by a breeze that blows from the sea during the daytime and from the land during night-time.
@@
 The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021.
-The second campaign was carried out in accordance with the ESCAPE project guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
-During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and HM.
+The second campaign, named "BiSCAPE", was carried out following the European Study of Cohorts for Air Pollution Effects (ESCAPE) project guidelines \cite{ESCAPE2010}, and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022.
+During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, Cornellà de Llobregat, Hospitalet de Llobregat and Sant Just Desvern) (\textbf{\cref{fig1}}).
+These monitoring campaigns were used in the development of the LUR and HM.
+Air pollution levels were measured using specialized  equipment for each pollutant.
+NO\textsubscript{2} was measured using Gradko NO\textsubscript{2} diffusion tubes.
+PM\textsubscript{2.5} and its chemical organic components were measured using a 37mm Polytetrafluroethylene - Teflon® - filters collected using a BGI-400 pump (working at 4 L/min) and a PCIS impactor, and BC measurements were conducted using a MicroAeth®.
 
@@
 Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites were during and after the pandemic period and its resulting restrictions.
-For PM\textsubscript{2.5}, BC  and PM\textsubscript{2.5} constituents models were developed using BiSCAPE data from the 34 sites.
-For the development of the LUR models, spatial variables that better represent the air pollutant concentration distribution across the study area were created.
-Following the ESCAPE guidelines \cite{eeftens2012, beelen2013},  predictor variables were selected from three broad categories: traffic, land use, and urban configuration.
-A wide variety of buffer sizes were created, i.e, 25m, 50m, 100m, 300m, 500m, and 1000m to capture the area-specific predictor variable at every single location.
+For PM\textsubscript{2.5}, BC  and PM\textsubscript{2.5} constituents, models were developed using BiSCAPE data from the 34 sites.
+To develop the LUR models, spatial variables that better represent the air pollutant concentration distribution across the study area were created.
+Following the ESCAPE guidelines \cite{ESCAPE2010, eeftens2012, beelen2013},  predictor variables were selected from three broad categories: traffic, land use, and urban configuration.
+Various buffer sizes were created, i.e, 25m, 50m, 100m, 300m, 500m, and 1000m to capture the area-specific predictor variable at every location.
 Predictor variables were calculated using specific coordinates for each site.
@@
 We then repeated the process using the previous model and all the remaining predictor variables.
-Details of the LUR modelling and variables used can be found in the \textbf{Table S1}.
+Details of the LUR modeling and predictvariables can be found in the \textbf{Table S1}.
 
@@
 For the hybrid models, we sought to improve their predictive power by integrating comprehensive data from multiple information sources and leveraging ML.
-For NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC were largely based on daily concentrations from dispersion models, summarized as weekly averages, and spatial predictor variables from the LUR models.
+NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC were largely based on daily concentrations from dispersion models, summarized as weekly averages, and spatial predictor variables from the LUR models.
 
 For a more comprehensive approach, we integrated additional data from the Catalan Air Pollution Monitoring and Forecasting Network (XVPCA) for Barcelona and surrounding areas \cite{xarxa2012}.
-The data was interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
-The data came from 16 monitoring stations, including Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St.
+Daily air pollution concentration levels were interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
+The data were collected from 16 monitoring stations, including Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St.
 Adria, St.
@@
 In summary, by combining data from multiple sources, our hybrid models aim to provide more accurate and reliable predictions.
-These predictions are essential for studying and mitigating urban air pollution.
 The distribution of the air pollution monitoring stations, meteorological stations, and traffic count measures are shown in the (\textbf{\cref{fig1}}, \textbf{Panel D}).
@@
 
-We evaluated model performance using various statistical approaches and on different temporal bases.
-The LUR models were evaluated on an annual basis, dispersion models were assessed both hourly and daily, and the hybrid models were evaluated on a weekly basis.
+We evaluated model performance using various statistical approaches and on different temporal resolutions.
+The LUR models were evaluated on an annual basis, DM were assessed both hourly and daily, and HM were evaluated on a weekly basis.
 
@@
 This process was repeated \(k-times\), and the average of these measures was reported as the indicator of the model validity.
-Details from the metrics used for model validation can be found in the \textbf{Supplementary material X}.\\
+Details from the metrics used for model validation can be found in the \textbf{supplementary material}.\\
 
@@
 For the external validation of the NO\textsubscript{2} LUR model, we applied data from BiSC-home participants' homes for those participants who had data available only for one measurement campaign, and therefore their data were not used for developing NO\textsubscript{2} LUR model.
-For LUR models BC, PM\textsubscript{2.5}, and its constituents (Cu, Fe, Sb, and Zn), internal validation was conducted using the \textit{leave-one-out-cross-validation (LOOCV)} process.
-For these models, we were not able to conduct external validation due to the lack of external data.
+For PM\textsubscript{2.5}, and its constituents (Fe, Cu, Zn), and BC, internal validation was conducted using the \textit{leave-one-out-cross-validation (LOOCV)} process.
+For these models, we were not able to conduct external validation due to the lack of data.
 
 \subsection{Dispersion model evaluation}
- Briefly, once input data has been introduced to the ADMS Urban model, the calculation of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC concentration levels has been carried out.
-Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
-The closer these values are, the better is the air quality modelling.
-Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material X}.\vspace{0.5cm}
+ Briefly, once the input data has been introduced to the ADMS Urban model, the calculation of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC concentration levels has been carried out.
+Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality monitoring stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
+The closer these values are, the better is the performance of the model.
+Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material}.\vspace{0.5cm}
 
@@
 We assessed the performance of each hybrid model by contrasting predictions with observations.
-Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and the \textit{out-of-bag (OBB) score}.
+Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and an external validation using 20\% of the data to test the performance of the model.
 While both methods vary in the sample size designated for training and testing, their core difference lies in the data-splitting strategy.
-In the 10-fold CV, the model trains on nine out of the ten data partitions and then validates against the remaining one.
-In contrast, the OOB approach trains the model on two-thirds of the data and uses the residual third for external validation.
+In the 10-fold CV, the model trains on nine out of the ten data partitions and then validates against the remaining one, while in the external validation procedure the model uses all the data observations available to check the performance of the model.
 We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distribution of monitoring campaigns}
+\subsection{Distribution of air pollution measurements}
 
 Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants is shown in \Cref{table1}.
-NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 and 97.3 \(\mu \text{g/m}^3\) with a mean of  38.8 \(\pm\) 12.4 \( \mu \text{g/m}^3\).
-PM\textsubscript{2.5} observed levels ranged between 6.14 and 35.8 \( \mu \text{g/m}^3\) with a mean of 13.4 \(\pm\) 6.14 \(\mu \text{g/m}^3\), while BC levels showed less variation during the BiSCAPE campaign, ranging from 0.40 to 2.95 \(  \mu \text{g/m}^3\) with a mean of 1.27 \(\pm\) 0.50 \( \mu \text{g/m}^3\).
-For PM\textsubscript{2.5} constituents  observed values during the BiSCAPE campaign varied substantially between elements with means of 0.25 \(\pm\) 0.13 \(  \mu \text{g/m}^3\)  for  Fe,  7.10 \(\pm\) 3.82 \( \text{ng/m}^3 \) for Cu, and 40.7 \(\pm\) 36.5 \( \text{ng/m}^3 \) for Zn.
-\\
+NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 and 97.3 \(\mu \text{g/m}^3\) with a median of  37.5 \(\pm\) 15.8 \( \mu \text{g/m}^3\).
+PM\textsubscript{2.5} observed levels ranged between 5.01 and 35.8 \( \mu \text{g/m}^3\) with a median of 11.3 \(\pm\) 7.14 \(\mu \text{g/m}^3\), while BC levels showed less variation during the BiSCAPE campaign, ranging from 0.40 to 2.95 \(  \mu \text{g/m}^3\) with a median of 1.22 \(\pm\) 0.64 \( \mu \text{g/m}^3\).
+For PM\textsubscript{2.5} constituents  observed values during the BiSCAPE campaign varied substantially between elements with medians of 0.21 \(\pm\) 0.17 \(  \mu \text{g/m}^3\)  for  Fe,  6.36 \(\pm\) 4.61 \( \text{ng/m}^3 \) for Cu, and 28.2 \(\pm\) 27.8 \( \text{ng/m}^3 \) for Zn.
 
@@
 \label{table1} 
-\begin{tabular}{lllllllllll} % each one indicates a new column in the table 
+\begin{tabular}{llllll} % each one indicates a new column in the table 
 \toprule
-Campaign & Pollutant & Year  & mean & sd & min & p25 & p50 & p75 & max \\ 
+Campaign & Pollutant & Year  & median \(\pm\) IQR & min & max \\ 
 \midrule
-BiSC-home + BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\
+BiSC-home + BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 37.5 \(\pm\) 15.8 & 6.85 & 97.3 \\
 \hline
 \multirow{5}{*}{BiSCAPE} 
-& PM\textsubscript{2.5} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 13.4 & 6.14 & 5.01 & 9.06 & 11.3 & 16.2 & 35.8 \\ 
-& BC (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 1.27 & 0.50 & 0.40 & 0.88 & 1.22 & 1.52 & 2 .95 \\ 
-& PM\textsubscript{2.5 - Fe} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 0.25 & 0.13 & 0.07 & 0.15 & 0.21 & 0.32 & 0.77 \\ 
-& PM\textsubscript{2.5 - Cu} (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
-& PM\textsubscript{2.5 - Zn} (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
+& PM\textsubscript{2.5} (\(\mu\)g/m\textsuperscript{3}) & 
+2021-2022  & 11.3 \(\pm\) 7.14 & 5.01 & 35.8 \\ 
+& BC (\(\mu\)g/m\textsuperscript{3}) & 
+2021-2022 & 1.22 \(\pm\) 0.64 & 0.40 & 2.95  \\ 
+& PM\textsubscript{2.5 - Fe} (\(\mu\)g/m\textsuperscript{3}) & 
+2021-2022 & 0.21 \(\pm\) 0.17 & 0.07 & 0.77 \\ 
+& PM\textsubscript{2.5 - Cu} (ng/m\textsuperscript{3}) & 
+2021-2022 & 6.36 \(\pm\) 4.61 & 1.78 & 22.7 \\ 
+& PM\textsubscript{2.5 - Zn} (ng/m\textsuperscript{3}) & 
+2021-2022 & 28.2 \(\pm\) 27.8 & 6.71 & 273 \\ 
 \bottomrule
@@
 
-The spatial distribution of the outdoor air pollution sampling campaigns is shown in \Cref{fig1}.
-Figure 1A illustrates a high density coverage in the central area of Barcelona, where the BiSC-home campaign was conducted.
-Figure 1B exhibit a similar pattern of spatial distribution of the monitoring campaign but with a reduced density in sampling locations.
+The spatial distribution of the outdoor air pollution campaigns is shown in \Cref{fig1}.
+\textbf{Figure 1A} illustrates a high density coverage in the central area of Barcelona, where the BiSC-home campaign was conducted.
+\textbf{Figure 1B} exhibit a similar pattern of spatial distribution of the monitoring campaign but with a reduced density in sampling locations.
 
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                       
@@
 
-Since different approaches and sampling campaigns were used to built outdoor air pollution models were not directly comparable.
-
-The performance of the models varied substantially by pollutant.
-In table x.
-performance metrics for the HM are shown.
-Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE.
-Performance metrics for LUR, DM and HM are shown in the supplementary materials.
-
-\newpage
-
-\subsection{Comparison between models} % broad title then change this
-
-% --- Exposure concentrations --- %
-Predictions from outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig2}.
-The exposure range was larger for the LUR model for all the pollutants assessed.
-Dispersion model predicted lower exposure concentration values than LUR and Hybrid.
-In addition to that HM show less variability in the exposure concentration.
-
-%%% -------------------------------------------%%%%
-%%% --- Figure 2. Boxplot models comparison --- %%%
-%%% -------------------------------------------%%%%
-\begin{figure}[!htb]
-\includegraphics[width=0.9\textwidth]{figures/boxplot_all_models_estimates.png}
-\caption{Distribution of exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
-\label{fig2}
-\end{figure}
-
-\textbf{\Cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM.
-The data shown are long-term estimated concentrations of pollutants during the entire pregnancy period, after averaging weekly estimates for all the study period (2019 - 2021).
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Spatial distribution of outdoor air pollution estimates  --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The performance of the models varied substantially by pollutant \Cref{fig3}. Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE.
 
@@
 \caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
+\label{fig2a}
 \end{figure}
@@
 \caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and hm(\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
+\label{fig2b}
 \end{figure}
@@
 \newpage
+\subsection{Comparison between models} % broad title then change this
+% --- Exposure concentrations --- %
+Predictions from outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig2}.
+The exposure range was larger for the LUR model for all the pollutants assessed.
+Dispersion model predicted lower exposure concentration values than LUR and Hybrid.
+In addition to that HM show less variability in the exposure concentration.
+
+%%% -------------------------------------------%%%%
+%%% --- Figure 3. Boxplot models comparison --- %%%
+%%% -------------------------------------------%%%%
+\begin{figure}[!htb]
+\includegraphics[width=0.9\textwidth]{figures/boxplot_all_models_estimates.png}
+\caption{Distribution of exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
+\label{fig3}
+\end{figure}
+
+\textbf{\Cref{fig3}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM.
+The data shown are long-term estimated concentrations of pollutants during the entire pregnancy period, after averaging weekly estimates for all the study period (2019 - 2021).
+\newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-01-21 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Tittle of the paper --- %%%
+%%% --- Title of the paper --- %%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\textbf{Keywords:} Air Pollution; Traffic-related air pollutants; Exposure assessment; Spatiotemporal models.
+
 \newpage
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Further details regarding the ADMS-Urban model and the variables used can be found in \textbf{Supplementary materials X}.
+Further details regarding the ADMS-Urban model and the variables used can be found in \textbf{supplementary materials}.
 
@@
 We evaluated model performance using various statistical approaches and on different temporal resolutions.
-The LUR models were evaluated on an annual basis, DM were assessed both hourly and daily, and HM were evaluated on a weekly basis.
+The LUR models were evaluated on an annual basis, DM was assessed both hourly and daily, and HM was evaluated weekly.
 
@@
 Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality monitoring stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
-The closer these values are, the better is the performance of the model.
+The closer these values are, the better the model's performance.
 Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material}.\vspace{0.5cm}
@@
 \subsection{Hybrid model evaluation}
-We assessed the performance of each hybrid model by contrasting predictions with observations.
-Two distinct validation approaches were employed: the \textit{10-fold cross-validation} and an external validation using 20\% of the data to test the performance of the model.
+We assessed the performance of each hybrid model by contrasting predictions with observations. Two distinct validation approaches were employed. Firstly,  
+\textit{10-fold cross-validation} was used to provide an internal validation measure to this purpose the model was trained using 80\% of the data available.
+Secondly, external validation was performed by allocating 20\% of the data as a test set to further evaluate the model's performance.
 While both methods vary in the sample size designated for training and testing, their core difference lies in the data-splitting strategy.

### 2024-01-22 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 We then repeated the process using the previous model and all the remaining predictor variables.
-Details of the LUR modeling and predictvariables can be found in the \textbf{Table S1}.
+Details of the LUR modeling and predictor variables can be found in the \textbf{supplementary materials, Table 1}.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% --- Model Evaluation and Comparison--- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Performance assessment --- %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Performance assessment}
@@
 We evaluated model performance using various statistical approaches and on different temporal resolutions.
-The LUR models were evaluated on an annual basis, DM was assessed both hourly and daily, and HM was evaluated weekly.
+The LUR models were evaluated on an annual basis, DM was assessed both hourly and daily, and HM was evaluated on a weekly basis.
 
@@
 In this approach, the data was split into ten distinct sub-samples.
-For each iteration, the model was trained on k-1 folds and then validated using the leftover fold \cite{ziegel2003}.
+For each iteration, the model was trained on \textit{k-1} folds and then validated using the leftover fold \cite{ziegel2003}.
 For every fold, we computed metrics such as the coefficient of determination \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)).
@@
 \subsection{Hybrid model evaluation}
-We assessed the performance of each hybrid model by contrasting predictions with observations. Two distinct validation approaches were employed. Firstly,  
-\textit{10-fold cross-validation} was used to provide an internal validation measure to this purpose the model was trained using 80\% of the data available.
-Secondly, external validation was performed by allocating 20\% of the data as a test set to further evaluate the model's performance.
-While both methods vary in the sample size designated for training and testing, their core difference lies in the data-splitting strategy.
-In the 10-fold CV, the model trains on nine out of the ten data partitions and then validates against the remaining one, while in the external validation procedure the model uses all the data observations available to check the performance of the model.
-We follow a grid search for hyperparameter tuning  to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}.
+We assessed the performance of each hybrid model by contrasting predictions with observations.
+Two distinct validation approaches were employed.
+Firstly, \textit{10-fold cross-validation (10-fold CV} was used to provide an internal validation measure, where the model was trained on 80\% of the data.
+Secondly, external validation was performed by allocating 20\% of the data as a test set to further evaluate the model's performance against previously unseen data.
+While both methods differ in their approach to partitioning the data, the fundamental distinction lies in how the data is utilized during the training and validation phases.
+The \textit{10-fold CV} method systematically rotates through the data, ensuring that each subset serves as a validation set once, while the external validation approach provides a straightforward assessment against a separate data portion.
+
+Additionally, we employed a grid search for hyperparameter tuning to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}.
 
@@
 
-The spatial distribution of the outdoor air pollution campaigns is shown in \Cref{fig1}.
-\textbf{Figure 1A} illustrates a high density coverage in the central area of Barcelona, where the BiSC-home campaign was conducted.
-\textbf{Figure 1B} exhibit a similar pattern of spatial distribution of the monitoring campaign but with a reduced density in sampling locations.
+The spatial distribution of the outdoor air pollution campaigns is illustrated in \Cref{fig1}.
+\textbf{Figure 1A} shows a high-density coverage in the central area of Barcelona, where the BiSC-home campaign was conducted.
+\textbf{Figure 1B} exhibit a similar pattern in the spatial distribution of the monitoring campaign, albeit with a reduced density of sampling locations compared to the BiSC-home campaign.
 
@@
 
-The performance of the models varied substantially by pollutant \Cref{fig3}. Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE.
+The performance of the models varied substantially across different pollutants as shown in the metrics displayed in the \Cref{fig3}, albeit all models showed moderate to good performance.
+HM outperformed LUR and DM models in predicting  NO\textsubscript{2} (R\textsubscript{2} = 0.64), PM\textsubscript{2.5} (R\textsubscript{2} = 0.66), BC (R\textsubscript{2} = 0.86) with corresponding RMSE values 7.48 \mu\)g/m\textsuperscript{3}, 3.45 \mu\)g/m\textsuperscript{3} and 0.23 \mu\)g/m\textsuperscript{3} respectively.
+LUR models show better performance for traffic-related air pollutants tracers PM\textsubscript{2.5 - Fe} (R\textsuperscript{2} = ), PM\textsubscript{2.5 - Cu} (R\textsuperscript{2} = ), and PM\textsubscript{2.5 - Zn} (R\textsuperscript{2} = ) compared to HM.
 
@@
 \section{Discussion}
-We successfully developed spatiotemporal models using three different approaches (LUR, DM, and HM). Models show moderate to good performance 
+We successfully developed spatiotemporal models using three different approaches (LUR, ADM, and HM). Models show moderate to good performance 
 evaluated by the defined metrics previously and displayed in more detail in the \textbf{supplementary materials}.

### 2024-01-22 supplementary_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 \subsubsection{NO\textsubscript{2} deseasonalization}
-
 For the NO\textsubscript{2} data (BiSCAPE + BiSC) the \textit{difference method} was used for deseasonalisation.
@@
 \subsubsection{Traffic count measurement and meteorological data}
-While our dispersion model initiall
+While our dispersion model initially
 
@@
 
-\bibliography{references}
+%\bibliography{references}
 

### 2024-01-23 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 
-The performance of the models varied substantially across different pollutants as shown in the metrics displayed in the \Cref{fig3}, albeit all models showed moderate to good performance.
-HM outperformed LUR and DM models in predicting  NO\textsubscript{2} (R\textsubscript{2} = 0.64), PM\textsubscript{2.5} (R\textsubscript{2} = 0.66), BC (R\textsubscript{2} = 0.86) with corresponding RMSE values 7.48 \mu\)g/m\textsuperscript{3}, 3.45 \mu\)g/m\textsuperscript{3} and 0.23 \mu\)g/m\textsuperscript{3} respectively.
-LUR models show better performance for traffic-related air pollutants tracers PM\textsubscript{2.5 - Fe} (R\textsuperscript{2} = ), PM\textsubscript{2.5 - Cu} (R\textsuperscript{2} = ), and PM\textsubscript{2.5 - Zn} (R\textsuperscript{2} = ) compared to HM.
+The performance of the models varied substantially across different pollutants as shown in \Cref{fig3}.
+All models exhibit relatively good performance, evaluated by the following metrics; $R^{2}_{\text{10-CV}}$, $RMSE_{\text{10-CV}}$, $R^{2}_{\text{LOOCV}}$, and $RMSE_{\text{LOOCV}}$.
+The HM performance for NO\textsubscript{2} ($R^{2}_{\text{10-CV}} = 0.64$), PM\textsubscript{2.5} ($R^{2}_{\text{10-CV}} = 0.66$), and BC ($R^{2}_{\text{10-CV}} = 0.86$) was moderate to good, with corresponding $RMSE_{\text{10-CV}}$ values of $7.48 \, \mu\text{g/m}^{3}$, $3.45 \, \mu\text{g/m}^{3}$, and $0.23 \, \mu\text{g/m}^{3}$ while for PM\textsubscript{2.5} componentes, the performance was more conservative with PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{10-CV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{10-CV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{10-CV}} = 0.85$).
+Conversely, LUR models showed better performance in predicting PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{LOOCV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{LOOCV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{LOOCV}} = 0.85$) concentration levels in comparison to  NO\textsubscript{2} ($R^{2}_{\text{LOOCV}} = 0.62$), PM\textsubscript{2.5} ($R^{2}_{\text{LOOCV}} = 0.45$), and BC ($R^{2}_{\text{LOOCV}} = 0.83$).
 
@@
 % --- Exposure concentrations --- %
-Predictions from outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig2}.
+Predictions from outdoor air pollution concentrations exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig2}.
 The exposure range was larger for the LUR model for all the pollutants assessed.

### 2024-01-24 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

--- 2024-01-22 manuscript_BiSC_outdoor_AP.tex
@@
 
+ADM results were evaluated considering the locations of XVPCA stations, to evaluate the model performance on those locations.
+The model was calibrated to fit the model results to the data measured in each XVPCA station.
+
 In addition, bias-adjustment techniques have been used to fit the model data to the observations through hourly bias spatial interpolation and generation of hourly maps.
@@
 
-We evaluated model performance using various statistical approaches and on different temporal resolutions.
+We evaluated model performance using various statistical approaches, using different external validation datasets, and on different temporal resolutions.
 The LUR models were evaluated on an annual basis, DM was assessed both hourly and daily, and HM was evaluated on a weekly basis.
@@
 For each iteration, the model was trained on \textit{k-1} folds and then validated using the leftover fold \cite{ziegel2003}.
-For every fold, we computed metrics such as the coefficient of determination \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)).
+For every fold, we computed metrics such as the coefficient of determination \(R^2\) and the root-mean-square-error (RMSE).
 After each iteration, both the error and adjusted \(R\textsuperscript{2}\) were recorded.
@@
 For the external validation of the NO\textsubscript{2} LUR model, we applied data from BiSC-home participants' homes for those participants who had data available only for one measurement campaign, and therefore their data were not used for developing NO\textsubscript{2} LUR model.
-For PM\textsubscript{2.5}, and its constituents (Fe, Cu, Zn), and BC, internal validation was conducted using the \textit{leave-one-out-cross-validation (LOOCV)} process.
-For these models, we were not able to conduct external validation due to the lack of data.
+For PM\textsubscript{2.5}, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn}, and BC, internal validation was conducted using the \textit{leave-one-out-cross-validation (LOOCV)} process.
+For these models, we were not able to conduct external validation due to the lack of data availability.
 
@@
 Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality monitoring stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
+Additionally, we compared the results of the model with data from the 4 campaigns carried out at BiSCAPE sites.
+In total, 98 periods were used to evaluate the performance of NO\textsubscript{2} model, 66 periods for PM\textsubscript{2.5}, and 93 periods for BC.
 The closer these values are, the better the model's performance.
+
 Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material}.\vspace{0.5cm}
@@
 
-\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                       
@@
 
-The performance of the models varied substantially across different pollutants as shown in the metrics displayed in the \Cref{fig3}, albeit all models showed moderate to good performance.
-HM outperformed LUR and DM models in predicting  NO\textsubscript{2} (R\textsubscript{2} = 0.64), PM\textsubscript{2.5} (R\textsubscript{2} = 0.66), BC (R\textsubscript{2} = 0.86) with corresponding RMSE values 7.48 \mu\)g/m\textsuperscript{3}, 3.45 \mu\)g/m\textsuperscript{3} and 0.23 \mu\)g/m\textsuperscript{3} respectively.
-LUR models show better performance for traffic-related air pollutants tracers PM\textsubscript{2.5 - Fe} (R\textsuperscript{2} = ), PM\textsubscript{2.5 - Cu} (R\textsuperscript{2} = ), and PM\textsubscript{2.5 - Zn} (R\textsuperscript{2} = ) compared to HM.
+\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using land use (LUR) regression models, dispersion models, and hybrid models.
+The performance of the models varied substantially across different pollutants as shown in \Cref{fig2} and \Cref{fig3}.
+All models exhibit relatively good performance, evaluated by the $R^{2}_{\text{10-CV}}$, $RMSE_{\text{10-CV}}$, $R^{2}_{\text{LOOCV}}$, and $RMSE_{\text{LOOCV}}$.
+
+The HM demonstrated moderate to good performance for NO\textsubscript{2} ($R^{2}_{\text{10-CV}} = 0.64$), PM\textsubscript{2.5} ($R^{2}_{\text{10-CV}} = 0.66$), and BC ($R^{2}_{\text{10-CV}} = 0.86$), with corresponding $RMSE_{\text{10-CV}}$ values of $7.48 \, \mu\text{g/m}^{3}$, $3.45 \, \mu\text{g/m}^{3}$, and $0.23 \, \mu\text{g/m}^{3}$, respectively.
+However, for PM\textsubscript{2.5} components, the performance was more conservative, with PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{10-CV}} = 0.54$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{10-CV}} = 0.70$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{10-CV}} = 0.44$).
+
+In contrast, LUR models showed better performance in predicting PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{LOOCV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{LOOCV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{LOOCV}} = 0.85$) concentration levels in comparison to  NO\textsubscript{2} ($R^{2}_{\text{LOOCV}} = 0.62$), PM\textsubscript{2.5} ($R^{2}_{\text{LOOCV}} = 0.45$), and BC ($R^{2}_{\text{LOOCV}} = 0.83$) exposure concentrations.
+
+Dispersion models were evaluated against observations from the XVPC monitoring stations, we average the metrics obtained during this process to gauge the overall performance of the dispersion model.
+A comparison between predicted and observed values for all the monitoring stations used in the DM is provided in the \textbf{supplementary materials}.
+
+The maps reveals distinct patterns of variability in estimated outdoor air pollution exposure concentrations.
+LUR models indicate high variability in exposure concentrations during pregnancy, whereas DM exhibit less variability.
+HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.
 
@@
 \caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
-\label{fig2a}
+\label{fig2}
 \end{figure}
@@
 \caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and hm(\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
-\label{fig2b}
+\label{fig3}
 \end{figure}
@@
 % --- Exposure concentrations --- %
-Predictions from outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig2}.
+Predictions from outdoor air pollution concentrations exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig4}.
 The exposure range was larger for the LUR model for all the pollutants assessed.
@@
 \caption{Distribution of exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
-\label{fig3}
+\label{fig4}
 \end{figure}
 
-\textbf{\Cref{fig3}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM.
+\textbf{\Cref{fig5}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM.
 The data shown are long-term estimated concentrations of pollutants during the entire pregnancy period, after averaging weekly estimates for all the study period (2019 - 2021).
@@
 \caption{Correlation between traffic-related air pollutants for LUR, DM, HM in different windows of exposure \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
+\label{fig5}
 \end{figure}
@@
 \caption{Comparison between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, Zn).}}
+\label{fig6}
 \end{figure}

### 2024-01-24 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 
-\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using land use (LUR) regression models, dispersion models, and hybrid models.
+\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using land use (LUR) regression models, dispersion models, and hybrid models.
 The performance of the models varied substantially across different pollutants as shown in \Cref{fig2} and \Cref{fig3}.
-All models exhibit relatively good performance, evaluated by the $R^{2}_{\text{10-CV}}$, $RMSE_{\text{10-CV}}$, $R^{2}_{\text{LOOCV}}$, and $RMSE_{\text{LOOCV}}$.
+In general, LUR and HM models exhibit relatively good performance, while DM shown a more conservative performance evaluated by the $R^{2}$ and $RMSE$.
 
-The HM demonstrated moderate to good performance for NO\textsubscript{2} ($R^{2}_{\text{10-CV}} = 0.64$), PM\textsubscript{2.5} ($R^{2}_{\text{10-CV}} = 0.66$), and BC ($R^{2}_{\text{10-CV}} = 0.86$), with corresponding $RMSE_{\text{10-CV}}$ values of $7.48 \, \mu\text{g/m}^{3}$, $3.45 \, \mu\text{g/m}^{3}$, and $0.23 \, \mu\text{g/m}^{3}$, respectively.
-However, for PM\textsubscript{2.5} components, the performance was more conservative, with PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{10-CV}} = 0.54$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{10-CV}} = 0.70$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{10-CV}} = 0.44$).
+The HM demonstrated moderate to good performance for NO\textsubscript{2} ($R^{2}_{\text{ 10-CV}} = 0.64$), PM\textsubscript{2.5} ($R^{2}_{\text{ 10-CV}} = 0.66$), and BC ($R^{2}_{\text{ 10-CV}} = 0.86$), with corresponding $RMSE_{\text{ 10-CV}}$ values of $7.48 \, \mu\text{g/m}^{3}$, $3.45 \, \mu\text{g/m}^{3}$, and $0.23 \, \mu\text{g/m}^{3}$, respectively.
+However, for PM\textsubscript{2.5} components, the performance was more conservative, with PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ 10-CV}} = 0.54$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ 10-CV}} = 0.70$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ 10-CV}} = 0.44$).
 
-In contrast, LUR models showed better performance in predicting PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{LOOCV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{LOOCV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{LOOCV}} = 0.85$) concentration levels in comparison to  NO\textsubscript{2} ($R^{2}_{\text{LOOCV}} = 0.62$), PM\textsubscript{2.5} ($R^{2}_{\text{LOOCV}} = 0.45$), and BC ($R^{2}_{\text{LOOCV}} = 0.83$) exposure concentrations.
+In contrast, LUR models showed better performance in predicting PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ LOOCV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ LOOCV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ LOOCV}} = 0.85$) concentration levels in comparison to  NO\textsubscript{2} ($R^{2}_{\text{ LOOCV}} = 0.62$), PM\textsubscript{2.5} ($R^{2}_{\text{ LOOCV}} = 0.45$), and BC ($R^{2}_{\text{ LOOCV}} = 0.83$) exposure concentrations.
 
-Dispersion models were evaluated against observations from the XVPC monitoring stations, we average the metrics obtained during this process to gauge the overall performance of the dispersion model.
+Dispersion models, when compared with observations from the XVPCA monitoring stations were evaluated against observations from the XVPCA monitoring stations, yielded an $R^{2} = 0.39$ for NO\textsubscript{2}, $R^{2} = 0.26$ for PM\textsubscript{2.5}, and $R^{2} = 0.65$ for BC.
 A comparison between predicted and observed values for all the monitoring stations used in the DM is provided in the \textbf{supplementary materials}.
@@
 The maps reveals distinct patterns of variability in estimated outdoor air pollution exposure concentrations.
-LUR models indicate high variability in exposure concentrations during pregnancy, whereas DM exhibit less variability.
+LUR models shown a high variability in exposure concentrations during pregnancy, whereas DM exhibit less variability.
 HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.
 
-%%% --- Figure. Average exposure during the entire pregnancy period --- %%%
+%%% --- Figure 3. Average exposure during the entire pregnancy period --- %%%
 % Put the figure text closer to the Figure 3
@@
 
+%%% --- Figure 4. Average exposure during the entire pregnancy period --- %%%
 \captionsetup[figure]{skip=6pt}
@@
 
+Results from the external validation of the LUR, DM, and HM models, conducted where information was available is depicted in \Cref{table2}. For NO\textsubscript{2}
+
+%%% --- Table 2. Performance from external validation for LUR, DM, HM --- %%%
+
+\begin{table}
+    \centering
+    \caption{Performance from external validation for LUR, DM, and HM.}
+    \label{table2}
+    \begin{tabular}{llllllll}
+    \toprule
+         Model & Validation dataset & \multicolumn{2}{c}{NO\textsubscript{2}} &  \multicolumn{2}{c}{PM\textsubscript{2.5}} &  \multicolumn{2}{c}{BC} \\
+         \cmidrule(lr){3-8}
+         & & R\textsuperscript{2} & RMSE & R\textsuperscript{2} & RMSE & R\textsuperscript{2} & RMSE \\
+    \midrule
+         LUR & External Validation & - & - & - & - & - & - \\
+         DM & External Validation & - & - & - & - & - & - \\
+         HM & External Validation & - & - & - & - & - & - \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Comparison between models --- %%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newpage
 \subsection{Comparison between models} % broad title then change this
+
 % --- Exposure concentrations --- %
-Predictions from outdoor air pollution concentrations exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig4}.
+Estimated outdoor air pollution concentrations exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig4}.
 The exposure range was larger for the LUR model for all the pollutants assessed.
@@
 Correlation between the air pollutant estimates for the different models assessed was moderate to strongly correlated for both long-and-short term exposure.
-Nitrogen dioxide
+
+NO\textsubscript{2}
 

### 2024-01-25 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 \begin{figure}[!htb]
-\includegraphics[width=0.9\textwidth]{figures/combined_figure2.png}
+\includegraphics[width=0.9a\textwidth]{figures/combined_figure2.png}
 \caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and hm(\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
@@
 
-Results from the external validation of the LUR, DM, and HM models, conducted where information was available is depicted in \Cref{table2}. For NO\textsubscript{2}
+Results from validation of the LUR, DM, and HM models, conducted where information for validated the model was available are shown in \Cref{table2}.
+The validation of each model has been conducted using different validation datasets.
+NO\textsubscript{2} LUR model was validated using BiSC - home campaign data.
+NO\textsubscript{2}, PM\textsubscript{2.5} and BC from DM model were validated using data from BiSCAPE monitoring campaign.
+While HM models were validated using 20\% of data available initially before model training
 
 %%% --- Table 2. Performance from external validation for LUR, DM, HM --- %%%
-
-\begin{table}
+\begin{table}[h!]
 \centering
-    \caption{Performance from external validation for LUR, DM, and HM.}
+    \caption{Performance from validation for LUR, DM, and HM models.}
     \label{table2}
-    \begin{tabular}{llllllll}
+    \begin{tabular}{llcccccc}
     \toprule
-         Model & Validation dataset & \multicolumn{2}{c}{NO\textsubscript{2}} &  \multicolumn{2}{c}{PM\textsubscript{2.5}} &  \multicolumn{2}{c}{BC} \\
+         Model & Validation & \multicolumn{6}{c}{R\textsuperscript{2 \textsubscript{ Validation}}} \\
          \cmidrule(lr){3-8}
-         & & R\textsuperscript{2} & RMSE & R\textsuperscript{2} & RMSE & R\textsuperscript{2} & RMSE \\
+         & & NO\textsubscript{2} & PM\textsubscript{2.5} & BC & PM\textsubscript{2.5 - Fe} & PM\textsubscript{2.5 - Cu} & PM\textsubscript{2.5 - Zn} \\
     \midrule
-         LUR & External Validation & - & - & - & - & - & - \\
-         DM & External Validation & - & - & - & - & - & - \\
-         HM & External Validation & - & - & - & - & - & - \\
+         LUR & External validation BiSC-Home & 0.56 & - & - & - & - & - \\
+         DM & External validation BiSCAPE & 0.65 & 0.22 & 0.56 & - & - & - \\
+         HM & External validation split (test data) & 0.64 & 0.65 & 0.69 & 0.65 & 0.56 & 0.36  \\
     \bottomrule
@@
 \end{table}
+
+Due to lack of data availability we only evaluate the performance of NO\textsubscript{2} for LUR models.
+The performance of the LUR model decreased in the external validation by 16\%.
+When validated DM against BiSCAPE measurements model performance increase for NO\textsubscript{2} but decrease for PM\textsubscript{2.5} and BC.
+HM showed consistent results in performance using both approaches to validate the models for NO\textsubscript{2} and PM\textsubscript{2.5} models showed a similar performance that the one obtained using the \textit{10-CV}.
+While for BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} performance varied between validation approaches by -19.7\%, +20.37\%, -20.0\% , and -18.8\% respectively.
 
@@
 % --- Exposure concentrations --- %
-Estimated outdoor air pollution concentrations exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig4}.
+\textbf{\Cref{fig5}} presents the distribution of estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM at the address level.
+The data shown are long-term estimated concentrations of pollutants during the entire pregnancy period, after averaging weekly estimates for all the study period (2019 - 2021).
 The exposure range was larger for the LUR model for all the pollutants assessed.
-Dispersion model predicted lower exposure concentration values than LUR and Hybrid.
+Dispersion model predicted lower exposure concentration values than LUR and HM.
 In addition to that HM show less variability in the exposure concentration.
@@
 \begin{figure}[!htb]
-\includegraphics[width=0.9\textwidth]{figures/boxplot_all_models_estimates.png}
-\caption{Distribution of exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
+\includegraphics[width=0.7\textwidth]{figures/boxplot_all_models_estimates.png}
+\caption{Distribution of estimated exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
 \label{fig4}
 \end{figure}
-
-\textbf{\Cref{fig5}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM.
-The data shown are long-term estimated concentrations of pollutants during the entire pregnancy period, after averaging weekly estimates for all the study period (2019 - 2021).
-\newpage
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 \textbf{(figure 4.A)}.
-Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 \textbf{(figure 4.B)}.
+Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 \textbf{(figure 5.A)}.
+Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 \textbf{(figure 5.B)}.
 Correlation between the air pollutant estimates for the different models assessed was moderate to strongly correlated for both long-and-short term exposure.
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/correlation_fig.png}
+\includegraphics[width=0.7\textwidth]{figures/correlation_fig.png}
 \caption{Correlation between traffic-related air pollutants for LUR, DM, HM in different windows of exposure \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
@@
 \begin{figure}[!htb]
-\includegraphics[width=1.0\textwidth]{figures/final_HEXBIN.png}
+\includegraphics[width=0.7\textwidth]{figures/final_HEXBIN.png}
 \caption{Comparison between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, Zn).}}

### 2024-01-26 02.Code/01.-descriptive_AP.R
changes figure

@@ -1,7 +1,7 @@
 # Install packages
 pacman::p_load(tidyverse, purrr, skimr, caret, tictoc, lubridate, 
                mice, ggmice, plyr, data.table, mapview, ggmap, omsdata, sf, sp, patchwork,
-               ggpattern, GGally)
+               ggpattern, GGally, viridis)
 
 ##############################
 ### --- Load the data --- ###
@@ -469,26 +469,35 @@ ggplot() +
 
 
 lur_estimates <- read.csv("01.Data/lur_estimates.csv")
-
+colnames(lur_estimates)
 
 # summarise all the pollutants 
 lur_estimates_plot <- lur_estimates %>% 
-  dplyr::ungroup()
-
-lur_estimates_plot <- lur_estimates_plot %>% dplyr::group_by(gid) %>%  
-  dplyr::summarise_all(funs(mean))
+                      dplyr::ungroup()
 
 dplyr::glimpse(lur_estimates_plot)
 
+lur_estimates_plot <- lur_estimates_plot %>%
+                      dplyr::select(subject_id, 
+                                    no2_lur:zn_lur, lon, lat) %>% 
+                      dplyr::group_by(subject_id) %>%  
+                      dplyr::summarise_all(funs(mean))
+
+dplyr::glimpse(lur_estimates_plot)
+view(lur_estimates_plot)
+
 lur_estimates_plot <- lur_estimates_plot %>% 
-  dplyr::select(gid, subject_id, 
-                no2_lur, pm25_lur, bc_lur, fe_lur, cu_lur, zn_lur,
-                lon, lat) %>% 
-  tidyr::drop_na()
+                      dplyr::select(subject_id, 
+                                    no2_lur, pm25_lur, bc_lur, fe_lur, cu_lur, zn_lur,
+                                    lon, lat) %>% 
+                      tidyr::drop_na()
 
 
 lur_estimates_sf <- st_as_sf(lur_estimates_plot, coords = c("lon", "lat"), crs = 25831)
 lur_estimates_wgs84 <- st_transform(lur_estimates_sf, 4326)
+
+
+dplyr::glimpse(lur_estimates_sf)
 
 # Plotting with ggplot2
 dplyr::glimpse(amb_shp)
@@ -608,19 +617,25 @@ ggsave(plot = lur_map_plot, "03.Outputs/figures/lur_map_plot.png",
 ##################################################
 
 dm_estimates <- read.csv("01.Data/dm_estimates.csv")
-
[…]

### 2024-01-28 manuscript_BiSC_outdoor_AP.tex

@@
 
+\subsection{Comparison between model estimates}
+Exposure estimates obtained from LUR, DM, and HM were compared at the BiSC participant's address level for long- and short-term exposure.
+We compare the distribution of estimated concentrations between models and pollutants and calculate Spearman (Rho) correlation coefficients.
+We used scatterplots and Bland-Altmant plots to investigate the level of agreement between the different models.
+
 All statistical analyses were performed in R version 4.3.1 \cite{Rstudio}. GIS predictor variables were computed using X software.
@@
 
-Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants is shown in \Cref{table1}.
+Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants are shown in \Cref{table1}.
 NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 and 97.3 \(\mu \text{g/m}^3\) with a median of  37.5 \(\pm\) 15.8 \( \mu \text{g/m}^3\).
@@
 
-\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using land use (LUR) regression models, dispersion models, and hybrid models.
+\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using LUR, HM, and HM.
 The performance of the models varied substantially across different pollutants as shown in \Cref{fig2} and \Cref{fig3}.
-In general, LUR and HM models exhibit relatively good performance, while DM shown a more conservative performance evaluated by the $R^{2}$ and $RMSE$.
+In general, LUR and HM models exhibit relatively good performance, while DM shows a more conservative performance evaluated by the $R^{2}$ and $RMSE$.
 
@@
 \begin{figure}[!htb]
-\includegraphics[width=0.9\textwidth]{figures/combined_figure1.png}
+\includegraphics[width=0.9\textwidth]{figures/combined_figure1_v2.png}
 \caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
@@
 \begin{figure}[!htb]
-\includegraphics[width=0.9a\textwidth]{figures/combined_figure2.png}
-\caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and hm(\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
+\includegraphics[width=0.9\textwidth]{figures/combined_figure2_v2.png}
+\caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and HM (\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
 \label{fig3}
@@
 
-%%% --- Table 2. Performance from external validation for LUR, DM, HM --- %%%
+%%% --- Table 2. Performance from validation for LUR, DM, HM --- %%%
 \begin{table}[h!]
@@
 The performance of the LUR model decreased in the external validation by 16\%.
-When validated DM against BiSCAPE measurements model performance increase for NO\textsubscript{2} but decrease for PM\textsubscript{2.5} and BC.
+When validated DM against BiSCAPE measurements model performance increase for NO\textsubscript{2}, but decrease for PM\textsubscript{2.5} and BC.
 HM showed consistent results in performance using both approaches to validate the models for NO\textsubscript{2} and PM\textsubscript{2.5} models showed a similar performance that the one obtained using the \textit{10-CV}.
@@
 \section{Discussion}
-We successfully developed spatiotemporal models using three different approaches (LUR, ADM, and HM). Models show moderate to good performance 
-evaluated by the defined metrics previously and displayed in more detail in the \textbf{supplementary materials}.
-As we expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed.
-We believe that by combining multiple data sources, and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.
+We successfully developed spatiotemporal models using three approaches (LUR, ADM, and HM). Models show moderate to good performance 
+evaluated by previously defined metrics and displayed in more detail in the \textbf{supplementary materials}.
+As expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed.
+We believe that by combining multiple data sources and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.
 

### 2024-01-31 02.Code/01.-descriptive_AP.R
bland altmant plots

@@ -996,8 +996,18 @@ models_estimates$model <- factor(models_estimates$model,
                                 levels = c("LUR", "DM", "HM"),
                                 ordered = TRUE)
 
-dplyr::glimpse(model_estimates)
+dplyr::glimpse(models_estimates)
 view(model_estimates_elements)
+
+dplyr::glimpse(models_estimates_elements)
+
+
+models_estimates %>% dplyr::distinct(subject_id) %>% dplyr::count()
+models_estimates_elem,enmt
+
+models_estimates %>% skimr::skim()
+
+
 
 library(ggplot2)
 library(ggpattern)
@@ -2796,14 +2806,14 @@ skimr::skim(pm25_constituents_measures)
 # this is ok in the ggsave object
 NO2_LUR_model_plot_v2 <- NO2_LUR_model_plot +
                          annotate("text", x = Inf, y = Inf, label = expression(R^2~(LOOCV) == 0.62), 
-                                  hjust = 2.45, vjust = 1.96, size = 2.5, colour = "black") +
+                                  hjust = 2.10, vjust = 1.96, size = 2.5, colour = "black") +
                          annotate("text", x = Inf, y = Inf, label = "RMSE (LOOCV) = 4.05", 
-                                  hjust = 2.00, vjust = 4.7, size = 2.5, colour = "black")
+                                  hjust = 1.76, vjust = 4.7, size = 2.5, colour = "black")
 
 # this is alligned :)
 NO2_DM_model_plot_v2 <- NO2_DM_model_plot + 
                         annotate("text", x = Inf, y = Inf, label = expression(R^2 == 0.39), 
-                                 hjust = 4.60, vjust = 1.96, size = 2.5, colour = "black") +
+                                 hjust = 4.42, vjust = 1.96, size = 2.5, colour = "black") +
                         annotate("text", x = Inf, y = Inf, label = "RMSE = 16.3", 
                                  hjust = 3.10, vjust = 4.7, size = 2.5, colour = "black")
 
@@ -2826,18 +2836,18 @@ ggsave(plot = plot_prueba, "03.Outputs/figures/plot_prueba.png",
 
 # this is ok in the ggsave object
 PM25_LUR_model_plot_v2 <- PM25_LUR_model_plot +
-  annotate("text", x = Inf, y = Inf, label = expression(R^2~(LOOCV) == 0.9), 
[…]

### 2024-01-31 02.Code/04.-bland_altmant.R

@@ -1,0 +1,258 @@
+## 01.- bland_altmant
+
+library(quantr)
+
+
+#####################
+### --- Data --- ###
+####################
+estimates_correlation_data <- read.csv("01.Data/estimates_correlation_data.csv")
+dplyr::glimpse(estimates_correlation_data)
+
+########################################
+### --- NO2 Bland-Altmant plots --- ###
+######################################
+BA_no2_hm_lur <- quantr::plot_bland_altman(estimates_correlation_data, 
+                                   lcs_column= "no2_hm", 
+                                   reference_column = "no2_lur") +
+                 scale_x_continuous(breaks = c(0, 30, 60, 90, 120),
+                                    limits = c(0, 120)) +
+                 ylim(c(-60, 60)) + 
+                 xlab(expression(NO[2]~mean~HM ~and~ LUR~(mu*g/m^3))) + 
+                 ylab(expression(NO[2]~diff.~HM - LUR~(mu*g/m^3)))
+
+
+BA_no2_dm_lur <- quantr::plot_bland_altman(estimates_correlation_data,
+                                   lcs_column= "no2_dm", 
+                                   reference_column = "no2_lur") +
+                 scale_x_continuous(breaks = c(0, 25, 50, 75, 100, 125),
+                                    limits = c(0, 125)) +
+                 ylim(c(-60, 60)) + 
+                 xlab(expression(NO[2]~mean~DM ~and~ LUR~(mu*g/m^3))) + 
+                 ylab(expression(NO[2]~diff.~DM - LUR~(mu*g/m^3))) 
+
+
+BA_no2_hm_dm <- quantr::plot_bland_altman(estimates_correlation_data,
+                                           lcs_column= "no2_hm", 
+                                           reference_column = "no2_dm") +
+                scale_x_continuous(breaks = c(0, 30, 60, 90, 120),
+                                   limits = c(0, 120)) +
+                ylim(c(-60, 60)) + 
+                xlab(expression(NO[2]~mean~HM ~and~ DM~(mu*g/m^3))) + 
+                ylab(expression(NO[2]~diff.~HM - DM~(mu*g/m^3)))
+
+
+
+
+# putting the plots together
[…]

### 2024-01-31 manuscript_BiSC_outdoor_AP.tex

@@
 The maps reveals distinct patterns of variability in estimated outdoor air pollution exposure concentrations.
-LUR models shown a high variability in exposure concentrations during pregnancy, whereas DM exhibit less variability.
+LUR models shown higher variability in exposure concentrations during pregnancy, whereas DM exhibit less variability.
 HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.
 
-%%% --- Figure 3. Average exposure during the entire pregnancy period --- %%%
+%%% --- Figure 2. Average exposure during the entire pregnancy period --- %%%
 % Put the figure text closer to the Figure 3
@@
 % We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=0.9\textwidth]{figures/combined_figure1_v2.png}
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{figures/combined_figure1_v2.png}
 \caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
@@
 
-%%% --- Figure 4. Average exposure during the entire pregnancy period --- %%%
+%%% --- Figure 3. Average exposure during the entire pregnancy period --- %%%
 \captionsetup[figure]{skip=6pt}
 % We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=0.9\textwidth]{figures/combined_figure2_v2.png}
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{figures/combined_figure2_v3.png}
 \caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and HM (\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
@@
 \newpage
-\subsection{Comparison between models} % broad title then change this
+\subsection{Comparison of LUR, DM and HM at address level} % broad title then change this
 
 % --- Exposure concentrations --- %
-\textbf{\Cref{fig5}} presents the distribution of estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM at the address level.
+\textbf{\Cref{fig4}} presents the distribution of estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM at the address level.
 The data shown are long-term estimated concentrations of pollutants during the entire pregnancy period, after averaging weekly estimates for all the study period (2019 - 2021).
@@
 
+The overall estimated concentration for NO\textsubscript{2} was higher in the HM (36.7 \(\mu\)g/m\textsuperscript{3}) in comparison to LUR (36.4 \(\mu\)g/m\textsuperscript{3}) and DM (27.5 \(\mu\)g/m\textsuperscript{3}).
+PM\textsubscript{2.5} median concentration was higher for LUR (16.7 \mu\)g/m\textsuperscript{3}) compared to hM (12.2\mu\)g/m\textsuperscript{3}) and DM (11.8\mu\)g/m\textsuperscript{3}).
+
 %%% -------------------------------------------%%%%
-%%% --- Figure 3. Boxplot models comparison --- %%%
-%%% -------------------------------------------%%%%
-\begin{figure}[!htb]
-\includegraphics[width=0.7\textwidth]{figures/boxplot_all_models_estimates.png}
-\caption{Distribution of estimated exposure to outdoor air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
+%%% --- Figure 4. Boxplot models comparison --- %%%
+%%% ------------------------------------------%%%%
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{figures/boxplot_all_models_estimates.png}
+\caption{Distribution of estimated exposure to outdoor air pollution for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
 \label{fig4}
@@
 Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 \textbf{(figure 5.B)}.
-Correlation between the air pollutant estimates for the different models assessed was moderate to strongly correlated for both long-and-short term exposure.
-
-NO\textsubscript{2}
+Across the different models assessed, the correlation between the air pollutant estimates was moderate to strongly correlated for both long-and short-term exposure periods.
+Estimated concentrations of NO\textsubscript{2} ranged from 0.85 to 0.92.
+In terms of PM\textsubscript{2.5}, the correlation for estimated concentrations from LUR, DM, and HM varied from r = 0.54 to 0.66, while correlations for BC were stronger ranging from r = 0.64 to 0.88.
+Overall, NO\textsubscript{2} and BC from LUR, DM, and HM were strongly correlated, while for PM\textsubscript{2.5} the correlation was weaker.
+Highest correlation was found for pollutants estimated using DM than LUR and HM based pollutants.
 
@@
 % We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=0.7\textwidth]{figures/correlation_fig.png}
+\begin{figure}[]
+\includegraphics[width=0.8\textwidth]{figures/correlation_fig.png}
 \caption{Correlation between traffic-related air pollutants for LUR, DM, HM in different windows of exposure \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}

### 2024-01-31 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 Long-term exposure (entire pregnancy) at the residential addresses exhibited correlations ranging from rs = 0.48 to 0.92, while for short-term exposure (weekly exposure) varied from rs = 0.37 to 0.90.
-exposure.
 Our work emphasizes the importance of employing diverse exposure assessment methods to evaluate the exposure to traffic-related air pollutants and to disentangle differences in effect estimates during pregnancy.
@@
 ADMS-Urban is a specialized Gaussian model designed for analyzing pollutant emissions in urban and metropolitan areas, particularly at road level \cite{mchugh1997adms}.
-
 ADMS-Urban employs various algorithms to assess the chemical transportation and dispersion of pollutants, factoring in ground-level turbulence and turbulence induced by surrounding buildings \cite{stocker2012adms}.
-The model also features a meteorological pre-processing system that calculates fluxes and turbulence, enabling it to model airflow over complex terrains.
-
-For enhanced accuracy, the dispersion model integrates data from a three-dimensional building and street width model from the Institut Cartogràfic i Geològic de Catalunya \cite{stoter2016state, 3Dcity}.
+The model also features a meteorological pre-processing system that calculates fluxes and turbulence, enabling it to model airflow over complex terrains.For enhanced accuracy, the dispersion model integrates data from a three-dimensional building and street width model from the Institut Cartogràfic i Geològic de Catalunya \cite{stoter2016state, 3Dcity}.
 This data aids in the calculation of the street canyon effect, a phenomenon characterized by tall buildings situated adjacent to heavily trafficked roads, which hinders atmospheric dispersion, leading to higher and relatively consistent pollutant concentrations compared to areas unaffected by the canyon effect.
-
 The model also utilized meteorological data including wind speed, direction, air temperature, cloud cover, solar radiation, relative humidity, precipitation rate, and Planetary Boundary Layer Height (PBLH).
 This meteorological data was collected from the Zona Universitària station in Barcelona, managed by the Servei Meteorològic de Catalunya \cite{xema2013}.
-
 To incorporate background pollution levels at a regional scale, data from Copernicus Atmosphere Monitoring Service Information (CAMS) was used.
 Background air pollution data has been extracted from the CAMS reanalysis ensemble model for 2018, and from CAMS ensemble forecast model for 2019-2020 \cite{cams2020, franceinstitut}.
-
 ADM results were evaluated considering the locations of XVPCA stations, to evaluate the model performance on those locations.
 The model was calibrated to fit the model results to the data measured in each XVPCA station.
-
 In addition, bias-adjustment techniques have been used to fit the model data to the observations through hourly bias spatial interpolation and generation of hourly maps.
@@
 Each bias map by pollutant was associated to a cluster map class and an hourly-varying profile.
-
 Further details regarding the ADMS-Urban model and the variables used can be found in \textbf{supplementary materials}.
@@
 \subsection{Hybrid models}
-
 For the hybrid models, we sought to improve their predictive power by integrating comprehensive data from multiple information sources and leveraging ML.
 NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC were largely based on daily concentrations from dispersion models, summarized as weekly averages, and spatial predictor variables from the LUR models.
-
 For a more comprehensive approach, we integrated additional data from the Catalan Air Pollution Monitoring and Forecasting Network (XVPCA) for Barcelona and surrounding areas \cite{xarxa2012}.
@@
 Coloma, Vall Hebron, Zona Universitária \cite{xarxa2012}.
-
 While our dispersion model initially utilized meteorological and traffic data, the hybrid model broadened its scope by integrating daily traffic counts from 180 locations throughout the city \cite{trafficbcn}.
@@
 To further refine our insights into the influence of weather patterns on pollution, we sourced additional meteorological variables, such as wind speed and air temperature, from the Raval monitoring station managed by Servei Meteorològic de Catalunya \cite{xema2013}.
-
 We employed the RF algorithm to capture non-linear relationships and potential interactions between predictor variables and the response variable, utilizing \textit{the ranger package} in R implemented in the \textit{caret} framework \cite{wright2019}, a set of functions that attempt to streamline the process for creating predictive models \cite{caret2008}.
@@
 Predictions are then derived by averaging the outputs from all trees.
-
 In summary, by combining data from multiple sources, our hybrid models aim to provide more accurate and reliable predictions.
@@
 \subsection{Performance assessment}
-
-We evaluated model performance using various statistical approaches, using different external validation datasets, and on different temporal resolutions.
+We evaluated model performance using various statistical approaches, validation datasets, and on different temporal resolutions.
 The LUR models were evaluated on an annual basis, DM was assessed both hourly and daily, and HM was evaluated on a weekly basis.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- LUR model evaluation --- %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{LUR model evaluation}
@@
 This process was repeated \(k-times\), and the average of these measures was reported as the indicator of the model validity.
-Details from the metrics used for model validation can be found in the \textbf{supplementary material}.\\
+Details from the metrics used for model validation can be found in the \textbf{supplementary material}.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Dispersion model evaluation --- %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 \subsection{Dispersion model evaluation}
@@
 The closer these values are, the better the model's performance.
+Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material}.
 
-Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material}.\vspace{0.5cm}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Hybrid model evaluation --- %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Hybrid model evaluation}
@@
 The \textit{10-fold CV} method systematically rotates through the data, ensuring that each subset serves as a validation set once, while the external validation approach provides a straightforward assessment against a separate data portion.
-
 Additionally, we employed a grid search for hyperparameter tuning to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% --- Comparison between model estimates --- %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Comparison between model estimates}
@@
 We compare the distribution of estimated concentrations between models and pollutants and calculate Spearman (Rho) correlation coefficients.
-We used scatterplots and Bland-Altmant plots to investigate the level of agreement between the different models.
+We Bland-Altmant plots to investigate the level of agreement between the different types of models.
 
@@
 In general, LUR and HM models exhibit relatively good performance, while DM shows a more conservative performance evaluated by the $R^{2}$ and $RMSE$.
-
 The HM demonstrated moderate to good performance for NO\textsubscript{2} ($R^{2}_{\text{ 10-CV}} = 0.64$), PM\textsubscript{2.5} ($R^{2}_{\text{ 10-CV}} = 0.66$), and BC ($R^{2}_{\text{ 10-CV}} = 0.86$), with corresponding $RMSE_{\text{ 10-CV}}$ values of $7.48 \, \mu\text{g/m}^{3}$, $3.45 \, \mu\text{g/m}^{3}$, and $0.23 \, \mu\text{g/m}^{3}$, respectively.
-However, for PM\textsubscript{2.5} components, the performance was more conservative, with PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ 10-CV}} = 0.54$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ 10-CV}} = 0.70$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ 10-CV}} = 0.44$).
-
-In contrast, LUR models showed better performance in predicting PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ LOOCV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ LOOCV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ LOOCV}} = 0.85$) concentration levels in comparison to  NO\textsubscript{2} ($R^{2}_{\text{ LOOCV}} = 0.62$), PM\textsubscript{2.5} ($R^{2}_{\text{ LOOCV}} = 0.45$), and BC ($R^{2}_{\text{ LOOCV}} = 0.83$) exposure concentrations.
-
+However, for PM\textsubscript{2.5} components, the performance was more conservative, with PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ 10-CV}} = 0.54$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ 10-CV}} = 0.70$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ 10-CV}} = 0.44$).In contrast, LUR models showed better performance in predicting PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ LOOCV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ LOOCV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ LOOCV}} = 0.85$) concentration levels in comparison to  NO\textsubscript{2} ($R^{2}_{\text{ LOOCV}} = 0.62$), PM\textsubscript{2.5} ($R^{2}_{\text{ LOOCV}} = 0.45$), and BC ($R^{2}_{\text{ LOOCV}} = 0.83$) exposure concentrations.
 Dispersion models, when compared with observations from the XVPCA monitoring stations were evaluated against observations from the XVPCA monitoring stations, yielded an $R^{2} = 0.39$ for NO\textsubscript{2}, $R^{2} = 0.26$ for PM\textsubscript{2.5}, and $R^{2} = 0.65$ for BC.
@@
 
-The maps reveals distinct patterns of variability in estimated outdoor air pollution exposure concentrations.
-LUR models shown higher variability in exposure concentrations during pregnancy, whereas DM exhibit less variability.
-HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.
-
 %%% --- Figure 2. Average exposure during the entire pregnancy period --- %%%
-% Put the figure text closer to the Figure 3
-\captionsetup[figure]{skip=6pt}
-% We add the figure with the study domain 
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{figures/combined_figure1_v2.png}
+\captionsetup[figure]{skip=4pt} % Reduced space after caption
+\begin{figure}[!h]
+\includegraphics[width=0.75\textwidth]{figures/combined_figure1_v2.png}
 \caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
@@
 %%% --- Figure 3. Average exposure during the entire pregnancy period --- %%%
-\captionsetup[figure]{skip=6pt}
-% We add the figure with the study domain 
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{figures/combined_figure2_v3.png}
+\begin{figure}[!h]
+\includegraphics[width=0.75\textwidth]{figures/combined_figure2_v3.png}
 \caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and HM (\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
@@
 
-Results from validation of the LUR, DM, and HM models, conducted where information for validated the model was available are shown in \Cref{table2}.
-The validation of each model has been conducted using different validation datasets.
-NO\textsubscript{2} LUR model was validated using BiSC - home campaign data.
-NO\textsubscript{2}, PM\textsubscript{2.5} and BC from DM model were validated using data from BiSCAPE monitoring campaign.
-While HM models were validated using 20\% of data available initially before model training
+The maps reveals distinct patterns of variability in estimated outdoor air pollution exposure concentrations.
+LUR models shown higher variability in exposure concentrations during pregnancy, whereas DM exhibit less variability.
+HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.
 
@@
 
-Due to lack of data availability we only evaluate the performance of NO\textsubscript{2} for LUR models.
+Results from validation of the LUR, DM, and HM models, conducted where information for validated the model was available are shown in \Cref{table2}.
+The validation of each model has been conducted using different validation datasets.
+NO\textsubscript{2} LUR model was validated using BiSC - home campaign data.
+NO\textsubscript{2}, PM\textsubscript{2.5} and BC from DM model were validated using data from BiSCAPE monitoring campaign.
+While HM models were validated using 20\% of data available initially before model training.
+Due to lack of data availability we were able to evaluate the performance of NO\textsubscript{2} for LUR models.
 The performance of the LUR model decreased in the external validation by 16\%.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\subsection{Comparison of LUR, DM and HM at address level} % broad title then change this
+\subsection{Comparison of LUR, DM and HM estimates at address level} % broad title then change this
 
@@
 \textbf{\Cref{fig4}} presents the distribution of estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM at the address level.
-The data shown are long-term estimated concentrations of pollutants during the entire pregnancy period, after averaging weekly estimates for all the study period (2019 - 2021).
-The exposure range was larger for the LUR model for all the pollutants assessed.
-Dispersion model predicted lower exposure concentration values than LUR and HM.
-In addition to that HM show less variability in the exposure concentration.
-
-The overall estimated concentration for NO\textsubscript{2} was higher in the HM (36.7 \(\mu\)g/m\textsuperscript{3}) in comparison to LUR (36.4 \(\mu\)g/m\textsuperscript{3}) and DM (27.5 \(\mu\)g/m\textsuperscript{3}).
-PM\textsubscript{2.5} median concentration was higher for LUR (16.7 \mu\)g/m\textsuperscript{3}) compared to hM (12.2\mu\)g/m\textsuperscript{3}) and DM (11.8\mu\)g/m\textsuperscript{3}).
+The data shown are estimated concentrations of pollutants during the entire pregnancy period (weekly estimates) for all the study period (2019 - 2021).
+In our analysis, estimated median concentration for NO\textsubscript{2} was higher in the HM ($36.7 \, \mu\text{g/m}^{3}$) in comparison to LUR ($36.4 \, \mu\text{g/m}^{3}$) and DM ($27.5 \, \mu\text{g/m}^{3}$).
+However, for PM\textsubscript{2.5} median concentration was higher for LUR ($16.7 \, \mu\text{g/m}^{3}$) compared to HM ($12.2 \mu\text{g/m}^{3}$) and DM ($11.8 \, \mu\text{g/m}^{3}$).
+This trend was also observed in BC estimates, where LUR's median concentration was $1.34  \mu\text{g/m}^{3}$, higher than HM's ($1.21 , \mu\text{g/m}^{3}$) and significantly greater than DM's ($0.76 , \mu\text{g/m}^{3}$).
+For PM\textsubscript{2.5} constituents like Fe, Cu, and Zn, a different pattern was observed, with HM presenting higher median concentrations than those observed in LUR models.
+Overall, while HM generally indicated higher median concentrations estimates, LUR models demonstrated greater variability with larger ranges across all pollutants assessed.
+Dispersion models predicted lower exposure concentration values in all pollutants than LUR and HM.
+\\\\
 
@@
 %%% ------------------------------------------%%%%
-\begin{figure}
+\begin{figure}[!h]
 \includegraphics[width=0.8\textwidth]{figures/boxplot_all_models_estimates.png}
@@
 In terms of PM\textsubscript{2.5}, the correlation for estimated concentrations from LUR, DM, and HM varied from r = 0.54 to 0.66, while correlations for BC were stronger ranging from r = 0.64 to 0.88.
-Overall, NO\textsubscript{2} and BC from LUR, DM, and HM were strongly correlated, while for PM\textsubscript{2.5} the correlation was weaker.
+For PM\textsubscript{2.5} constituents we observed higher correlation coefficients for PM\textsubscript{2.5 - Fe} in LUR and HM for both long and short term exposure.
+Estimates for PM\textsubscript{2.5 - Cu} and PM\textsubscript{2.5 - Zn} showed lower correlations for both methods compared to Fe constituent.
+Overall, NO\textsubscript{2} and BC from LUR, DM, and HM were strongly correlated, while for PM\textsubscript{2.5} and constituents the correlation was weaker.
 Highest correlation was found for pollutants estimated using DM than LUR and HM based pollutants.
+Lowest correlation between outdoor air pollutants was found for HM estimates.
 
@@
 % We add the figure with the study domain 
-\begin{figure}[]
-\includegraphics[width=0.8\textwidth]{figures/correlation_fig.png}
+\begin{figure}[!h]
+\includegraphics[width=1.0\textwidth]{figures/correlation_fig_v2.png}
 \caption{Correlation between traffic-related air pollutants for LUR, DM, HM in different windows of exposure \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
@@
 \newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\Cref{fig6} shown Bland-Altmant plots to assess the level of agreement over the concentrations ranges between the models.
 
@@
 % We add the figure with the study domain 
-\begin{figure}[!htb]
-\includegraphics[width=0.7\textwidth]{figures/final_HEXBIN.png}
-\caption{Comparison between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, Zn).}}
+\begin{figure}[!h]
+\includegraphics[width=0.7\textwidth]{figures/final_bland_altmant.png}
+\caption{Bland-Altmant plots for comparing outdoor air pollution estimates for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, Zn).
+using LUR, DM, and HM}}
 \label{fig6}
@@
 \newpage
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Discussion}
-We successfully developed spatiotemporal models using three approaches (LUR, ADM, and HM). Models show moderate to good performance 
+We successfully developed outdoor air pollutants models using three approaches (LUR, DM, and HM). Models show moderate to good performance 
 evaluated by previously defined metrics and displayed in more detail in the \textbf{supplementary materials}.

### 2024-02-19 manuscript_BiSC_outdoor_AP.tex
finished paper for co-authors

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{\textbf{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}
+\title{\textbf{\Large Development and comparison of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}
 
@@
 %%% --- Authors --- %%%
-\author[1, 2, 3]{Alan Domínguez}
-\author[1, 3, 4]{Payam Dadvand}
-\author[1]{Marta Cirach}
+\author[1, 2, 3]{\small Alan Domínguez}
+\author[1, 2, 3]{Payam Dadvand}
+\author[1, 2, 3]{Marta Cirach}
+\author[4]{Gustavo Arevalo}
+\author[4]{Lluís Barril}
+\author[1, 2, 3, 5]{María Foraster}
+\author[1, 2, 3]{Mireia Gascon}
 \author[1]{Bruno Raimbault}
 \author[1]{Toni Galmes}
-\author[1]{Karl Samuelsson}
+\author[1, 2, 3]{Laura Goméz-Herrara}
+\author[1, 2, 3]{Cecilia Persavento}
+\author[1, 6]{Karl Samuelsson}
+\author[4]{Jose Lao}
+\author[7]{Teresa Moreno}
+\author[7]{Xavier Querol}
+\author[1, 2, 3]{Cathryn Tonne}
 \author[1, 2, 3]{Mark Nieuwenhuijsen}
 \author[1, 2, 3]{Jordi Sunyer}
-\author[1, 2, 3]{Xavier Basagaña}
-\author[1, 2, 3]{Ioar Rivas}
+\author[1, 2, 3, *]{Xavier Basagaña}
+\author[1, 2, 3, *]{Ioar Rivas}
 
 %%% --- Affiliations --- %%%
-\affil[1]{Barcelona Institute for Global Health (ISGlobal), Barcelona, Spain.}
+\affil[1]{\small Barcelona Institute for Global Health (ISGlobal), Barcelona, Spain.}
 \affil[2]{Universitat Pompeu Fabra (UPF), Barcelona, Spain.}
 \affil[3]{CIBER Epidemiología y Salud Pública (CIBERESP), Madrid, Spain.}
-\affil[4]{London School of Hygiene and Tropical Medicine (LSHTM), London, UK.}
+\affil[4]{Barcelona Regional}
+\affil[5]{Blanquerna School of Health Science, Universitat Ramon Llull (URL), Barcelona, Spain.}
+\affil[6]{Department of Building Engineering, Energy Systems and Sustainability Science, Faculty of Engineering and Sustainability Science, University of Gävle, Gävle, Sweden}
+\affil[7]{Institute of Environmental Assessment and Water Research (IDAEA-CSIC), Barcelona, Spain.}
+\affil[*]{Join last authors}
+
+% --- removing date from manuscript --- % 
+\date{}
 
@@
 Air pollution is the leading environmental risk factor for human health worldwide linked to a wide range of adverse health outcomes.
-A detailed and comprehensive assessment of outdoor air pollution exposure considering both spatial and temporal variability in urban areas is essential to identify critical windows of susceptibility during pregnancy to traffic-related air pollution.
-Within the framework of the Barcelona Life Study Cohort (BiSC), we developed Land use regression (LUR), dispersion (DM), and hybrid (HM) models to estimate outdoor air pollution exposure at residential addresses during pregnancy for several traffic-related air pollutants; NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn).
-The overall aim is to obtain accurate weekly estimates of traffic-related air pollution, intended to be used in the prenatal exposure assessment for a cohort of pregnant women in the city of Barcelona.
-Two separate monitoring campaigns were conducted: the "BiSC-home" campaign took place at the home addresses of BiSC participants and involved measuring only NO2 between 2018-2021.
+A detailed and comprehensive assessment of outdoor air pollution exposure considering both spatial and temporal variability in urban areas is essential to identify critical windows of susceptibility
+during developmental periods over the lifespan to traffic-related air pollution (TRAP).
+Within the framework of the Barcelona Life Study Cohort (BiSC), we developed Land use regression (LUR), dispersion (DM), and hybrid (HM) models to estimate outdoor air pollution exposure at residential addresses for several TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn).
+The overall aim is to obtain accurate weekly estimates of traffic-related air pollution, intended to be used in the exposure assessment for a cohort of pregnant women in the city of Barcelona.
+Two separate monitoring campaigns were conducted: the "BiSC-home" campaign took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} between 2018-2021.
 Additionally, a second campaign was conducted, named "BiSCAPE", and included four distinct measurement periods between 2021-2022.
@@
 The exposure estimates from these different models can be used in the BiSC cohort to potentially identify critical windows of susceptibility to air pollution during pregnancy.
+\\
+
+\textbf{Keywords:} Air Pollution; Traffic-related air pollutants; Exposure assessment; Spatiotemporal models.
 
 \end{abstract}
-
-\textbf{Keywords:} Air Pollution; Traffic-related air pollutants; Exposure assessment; Spatiotemporal models.
 
@@
 
-While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains (multi-country and national domain models), these models often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
+While significant advancements have been made in spatiotemporal modeling of air pollution over large-scale domains (multi-country and national domain models), these models often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}.
 This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to capture the differences in exposure for the population living in relatively small areas (i.e.
@@
 \subsection{Dispersion model evaluation}
-Briefly, once the input data has been introduced to the ADMS Urban model, the calculation of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC concentration levels has been carried out.
+Briefly, once the input data has been introduced to the ADMS-Urban model, the calculation of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC concentration levels has been carried out.
 Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality monitoring stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
@@
 %%% --- Hybrid model evaluation --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Hybrid model evaluation}
@@
 %%% --- Comparison between model estimates --- %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Comparison between model estimates}
 Exposure estimates obtained from LUR, DM, and HM were compared at the BiSC participant's address level for long- and short-term exposure.
-We compare the distribution of estimated concentrations between models and pollutants and calculate Spearman (Rho) correlation coefficients.
-We Bland-Altmant plots to investigate the level of agreement between the different types of models.
+We compare the distribution of estimated concentrations between models and pollutants and calculate Spearman correlation coefficients.
+Bland-Altmant plots were used to investigate the level of agreement between the different types of models.
+\\
 
@@
 
-\Cref{fig6} shown Bland-Altmant plots to assess the level of agreement over the concentrations ranges between the models.
+\textbf{\Cref{fig6}} shown Bland-Altmant plots to assess the level of agreement across the concentrations ranges between the models.
+The limits of agreement, represented by the red dashed lines  contains the majority of the points with some outliers present in the plots indicating good agreement in the estimated concentration, but with potential disagreement at higher estimated concentrations, showing an upward trend when comparing the model`s estimates.
+Overall, when comparing outdoor air pollution estimates at address level the models show good agreement at lower and medium ranges, with low bias that is observed as the blue ticked line is close to 0 for almost all comparison between models.
 
 %%% --- Figure. Comparison/agreement between moodel predictions estimates --- %%%
-% Put the figure text closer to the Figure 3
-
 % Put the figure text closer to the Figure 3
@@
 \includegraphics[width=0.7\textwidth]{figures/final_bland_altmant.png}
-\caption{Bland-Altmant plots for comparing outdoor air pollution estimates for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, Zn).
+\caption{Bland-Altmant plots for comparing outdoor air pollution estimates for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, and Zn).
 using LUR, DM, and HM}}
@@
 \end{figure}
+\vspace*{-\baselineskip}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Discussion}
-We successfully developed outdoor air pollutants models using three approaches (LUR, DM, and HM). Models show moderate to good performance 
-evaluated by previously defined metrics and displayed in more detail in the \textbf{supplementary materials}.
-As expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed.
-We believe that by combining multiple data sources and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.
 
-\subsection{Comparison between LUR, DM, and Hybrid model}
+% --- General --- % 
+We successfully developed outdoor air pollution models (NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents) to assess 
+exposure within the city of Barcelona for the period 2018-2021.
+Utilizing three different approaches: land use regression, dispersion modeling, and hybrid models were tailored to provide exposure estimates across various time windows releveant to our study population in the BiSC cohort, including entire pregnancy duration, specific trimesters, and weeks of pregnancy for each participant.
+The performance of these models was moderate to good, as determined by different evaluation metrics.
+The LUR models accounted for approximately 45\% to 89\% of the variance in pollutant levels, indicating their strength in capturing spatial variability in air pollution concentrations.
+HM models, which integrate a broader range of data inputs and modeling techniques, explained 44\% to 86\% of the variance.
+In contrast, DM models, focusing on physical and chemical processes affecting pollutants, explained a lower range of variability, between 26\% and 65\%.
+These findings align with the outcomes of similar air pollution modeling studies conducted across Europe and specifically within Barcelona \cite{fung2024, mila2023, criado2023, rovira2022, chen2020, dehoogh2018, dehoogh2013}.
+For both LUR and HM approaches, the selection of predictor variables demonstrated considerable variability across different pollutants, reflecting the influence of distinct emission sources.
+Notably, BC achieved the highest model performance across all methods, followed by NO\textsubscript{2}, PM\textsubscript{2.5}, and their constituents.
+This pattern highlights the critical role of source-specific factors in influencing the accuracy and reliability of pollution exposure estimates.\\
 
-\subsection{Comparison with Previous studies}
+% --- NO2 models --- %
+The performance of our LUR NO\textsubscript{2} model was higher than previous European models \cite{rivera2012} as well as earlier models specific to Barcelona \cite{criado2023}, indicating an enhanced capacity to capture the intra-urban variability of NO\textsubscript{2} concentrations levels.
+Notably, our LUR model outperformed a recently developed microscale LUR model, which was part of a data fusion method designed to minimize prediction uncertainty by integrating a LUR base map with hourly dispersion estimates from the CALLIOPE-Urban model within Barcelona \cite{criado2023}.
+The predictor variables of our model were predominantly traffic-related, and were selected due to their strong association with NO\textsubscript{2} emissions \cite{casallas2018}.
+Conversely, our DM model exhibits a lower performance compared to the CALLIOPE-Urban system, a regional-to-street-scale modelling framework that offers high-spatial-and-high-temporal-resolution (10m x 10m, hourly) by integrating the mesoscale airquality forecasting system CALIOPE (WRF-HERMES-CMAQ-BSS-DREAM8b) with an urban roadwide dispersion model (R-LINE) \cite{benavides2019}.
+Regarding our HM, where we integrated detailed information on traffic counts across the city of Barcelona, air pollution and meteorological data from monitoring stations, and estimates from the DM model through a RF algorithm, we achieved better performance compared to both our LUR and DM models.
+Most relevant predictor variables were estimates from the DM, monitoring station, traffic counts, solar radiation traffic-related spatial variables.
+We attribute the enhanced variance explanation by the HM to two main factors.
+Firstly, the incorporation of outputs from the DM model allows to improve the spatiotemporal variability in our model \cite{michanowicz2016, korek2017}, in addition traffic information allows to capture other emission sources not previously captured \cite{pinto2020} in our other two approaches.
+Secondly, the integration of the RF algorithm enables the model to account for non-linear relationships and potential interactions between predictor variables that the LUR and DM models did not fully exploit \cite{araki2018, chen2019}.
+Despite these advancements, a recent hybrid model developed for the entire region of Catalonia, which utilized information on air pollution measurements from monitoring stations, land use variables, and satellite remote sensing products, outperformed the performance of our HM for NO\textsubscript{2} \cite{mila2023}.
+This improvement is likely attributable to the use of data from the TROPOspheric Monitoring Instrument (TROPOMI), providing deeper insights into the vertical atmospheric distribution of NO\textsubscript{2} \cite{petetin2023}.\\
 
-\subsection{Strengths and Limitations}
-One of the main strengths of this study is the extensive exposure assessment
+% --- PM25 particles --- % add predictor variables 
+PM\textsubscript{2.5} have a moderate performance in LUR models similar to  previous models that aim to predict within-city variability in spanish cities \cite{rivera2012, wang2014} but lower than other multi-cities LUR models developed in Europe \cite{eeftens2012, gulliver2018}.
+Predictor variables used in the LUR model were mainly traffic, road density, and residential variables.
+We believe that the performance obtained from our LUR model can only explain a part of the variation of the  PM\textsubscript{2.5} concentrations across Barcelona.
+One potential cause that could partially explain the moderate performance of our LUR model is that PM\textsubscript{2.5} in Barcelona came from nine to ten major contribution sources that have been previously identified in two chemical species and source contribution studies for PM\textsubscript{2.5} \cite{in2021, amato2024}, we believe that our model was not able to take into account the variability of this sources since LUR models mainly use traffic-related variables to predict PM\textsubscript{2.5} concentrations levels, and previously have been described that LUR models are less effective to predict sources other than traffic \cite{dehoogh2014}.
+The performance of our DM model was less effective than the one obtained from the LUR model.
+One potential explanation could be the limited coarser spatial resolution from DM compared to LUR.
+Our lUR model uses observed data and land use characteristics to estimate air pollution levels, potentially offering a finer spatial resolution that better captures local variations, especially within urban settings \cite{hoek2017}.
+Since there were no previously available published DM models PM\textsubscript{2.5} developed in the city of Barcelona or at a Catalunya scale we were not able to compare the performance.
+The enhanced performance of our PM\textsubscript{2.5} HM model compared to our previous approaches can largely be attributed to the incorporation of comprehensive information during the model-building process and  the employment of the RF algorithm significantly increased our model's accuracy in predicting air pollution levels.
+This approach enabled us to account for additional sources of variability in PM\textsubscript{2.5} concentration levels, identifying emissions sources not captured by the other models.
+Additionally, our HM model demonstrated improved performance metrics compared to a previously developed RF model \cite{mila2023}.
+Regardless of these advancements in modeling PM\textsubscript{2.5} concentrations by our HM, it was found that recent European multi-stage HM models, utilizing RF algorithm for predicting daily PM\textsubscript{2.5} levels at a country-level with a high spatiotemporal resolution, exhibit superior performance metrics.
+These models integrated satellite Aerosol Optical Depth (AOD) measurements along with PM and meteorological data monitored at various locations and spatial predictors at the grid cell level, enhancing predictive accuracy beyond our HM model \cite{dehoogh2018, stafoggia2019, stafoggia2020, schneider2020}.
+This difference in model performance can be attributed to two main factors.
+First, the comprehensive exposure assessment period employed by these models, which utilized routine several years of data from routine monitoring stations to more effectively capture temporal trends in air pollution exposure levels.
+Second, the inclusion of AOD measurements provided additional insights into the background and regional variability of air pollution levels with higher correlations in Western Europe compared to the rest of the Region \cite{chen2024}.\\
+
+% --- Black Carbon (BC) --- %
+BC models have the highest performance across all pollutants, this was systematically observed in all our exposure assessment methods explored. 
+The LUR-developed model has greater performance than previous BC models developed in Europe \cite{hoek2015, van2020}.
+However, there are only a few studies that aim to predict BC concentrations levels across Europe, this limited our comparison with other models.
+Our HM outperformed a previous hybrid European-wide model developed using large-scale satellite data, chemical transport models, and small-scale and land use predictors to predict BC \cite{dehoogh2018}, and a hybrid LUR model that
+to estimate BC and metal components across the metropolitan area of Pittsburgh, USA \cite{tripathy2019}.
+Predictor variables used for the development of LUR and HM models were mainly traffic, residential, and population density-related.
+We believe that the high performance in both LUR and HM is that BC levels proportionally reflect traffic-related process  \\
+
+% --- PM2.5 constituents particles --- % 
+LUR has a very good performance in predicting PM\textsubscript{2.5} constituents pollutants, while HM has a moderate performance.
+In both LUR and HM models, traffic-related variables emerged as significant predictors for PM\textsubscript{2.5} constituents, this was consistent with results obtained from stepwise and RF analysis.
+This aligns with previous evidence indicating the influence of traffic-related variables that explain a large proportion of the variability in the measured concentrations \cite{dehoogh2013, chen2020}.
+Additionally, industry-related variables appears as relevant predictors in predicting PM\textsubscript{2.5 - Cu} and \textsubscript{2.5 - Zn} in both LUR and HM, possibly reflecting emissions from industry sector \cite{belis2013, chen2020}.
+Our PM\textsubscript{2.5} constituents models outperformed in performance to those published in previous studies \cite{}.
+Differences in performance between LUR and HM models in predicting PM\textsubscript{2.5} constituents could be explained due to stability for long-term concentrations in PM\textsubscript{2.5} composition.\\
+
+% --- Predicted concentrations ranges --- %
+
+% --- Correlations and agreement between models --- %
+Comparisons of air pollution estimates at the address level indicated that the correlations between the different models' estimates were moderate to strong for both long-term and short-term exposure periods.
+For NO\textsubscript{2}, PM\textsubscript{2.5} and BC we observed a positive correlation for both long-and-short exposures.
+Long-term correlations were strong for NO\textsubscript{2}, PM\textsubscript{2.5} and BC in comparison to short-term correlations.
+This is mainly due to a smoothing effect over the pregnancy period that could lead to more stable correlations between pollutants and also a seasonality effect between trimesters \cite{wang2018, morgan2023, hao2023}.
+This was not observed for PM\textsubscript{2.5} constituents, on the contrary, we observed the reverse pattern in correlation showing stronger correlations for short-term exposure (weekly exposures).
+Correlation between pollutants estimated for the same exposure assessment method was higher for DM, this pattern was also observed in  a previous study \cite{klompmaker2021}.
+One potential explanation could be that DM models often rely on the same emission source inventory to predict different air pollutants that could lead to higher correlations between different pollutants in comparison to other approaches such as LUR and HM models.
+PM\textsubscript{2.5} constituents were positive correlated between them for both long- and-short term windows of exposure for LUR and HM models.
+PM\textsubscript{2.5 - Zn} estimated using LUR and HM models were poorly correlated against estimated pollutants (NO\textsubscript{2}, PM\textsubscript{2.5}, and BC) from other exposure assessment methods, this pattern was observed for long-term and short-term exposure but was more pronounced for HM estimated pollutants.
+We believe that this results could be partially explained since both PM\textsubscript{2.5 - Fe} and PM\textsubscript{2.5 - Cu} come mainly from traffic, particularly from venhicle non-exhaust factors, while PM\textsubscript{2.5 - Zn} could came from industry emission sources as well this could explain the magnitude of some correlations in \textbf{Fig 5.A} and \textbf{Fig 5.B} \cite{amato2024}.
+Overall, there was good agreement among the LUR, DM, and HM model estimates for all pollutants.
+However, the lowest agreement was observed for higher air pollution estimates, suggesting that at elevated pollution levels, discrepancies between model estimates become more pronounced.
+Our reported findings were consistent with similar studies, which report similar performance metrics for the outdoor air pollutants assessed and relatively good agreement across different \cite{mila2023} and models exposure assessment methods \\
+
+% --- Strengths --- %
+Our study has several strengths.
+First, we use three different approaches to develop outdoor air pollution models for different TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn), allowing for a comprehensive assessment of the effects of air pollution in further epidemiological studies.
+Secondly, we take into account the potential measurement error introduced by the COVID pandemic in the estimations of our models by incorporating detailed information from pre-and- post pandemic traffic across the city, XVPCA monitoring station network, and traffic scenarios, allowing us to capture changes in air pollution trends across Barcelona.
+Thirdly, we report different metrics to assess the performance of outdoor air pollution models that are commonly employed to assess performance in different approaches to estimate air pollution levels.
+Additionally, we report the results for DM models which are frequently overlooked in the exposure assessment components of epidemiological studies.
+Fourthly, our LUR and HM models had the advantage of a large training dataset in measured concentrations by combining measurements from the BiSC-home and BiSCAPE monitoring campaign with different contrasts in exposure levels across Barcelona, enhancing the robustness and depth of our analysis.
+Finally, to our knowledge, this is one of the first studies to comprehensively assess different TRAP's using a multifaceted modeling strategy that spans both pre- and post-pandemic periods.
+It assesses their relative performance at an urban scale, enabling the disentanglement of differences in effect estimates derived from distinct exposure assessment methodologies.
+This approach allows for a detailed examination of within-city variability in exposure, contributing valuable insights into further epidemiological studies.\\
+
+% --- Limitations --- % 
+The results obtained in this study, which stands as one of the first comprehensive assessments of TRAPs, must be interpreted within the context of its limitations. 
+First, the methodology employed in the construction of our different models difficult direct comparisons between models, thus making it challenging to draw strong conclusions about their relative performance and accuracy.
+This inherent complexity in model comparison underscores the need for cautious interpretation of our results.
+Second, the DM was designed to predict air pollution levels at specific locations and within particular temporal frames.
+This specificity limits the reproducibility and broader applicability of the HM, which rely on dispersion estimates as a fundamental component of their predictive variables.
+However, this does not limit the potential utility of our LUR models that could be extended and used in  following epidemiological studies, out of the initial period for what the model was built.
+These models, despite the study's limitations, offer valuable insights and could be adapted for use in future epidemiological research beyond the initial scope for which they were developed.
+Third, the absence of external validation data restricts our capacity to evaluate  the accuracy of our preferred modelling approach.
+Despite this, we assess the performance of the NO\textsubscript{2} model given some lights regarding the most suitable approach for our context.
+However, we can no extending these findings to other pollutants due to differences in their sources, behavior, and environmental interactions.
+Despite these challenges, our study lays the groundwork for future research in exposure assessment in urban areas, providing a basis for methodological refinement and encouraging the development of more generalized models that can be  externally validated.
 
@@
 
-Air pollution exposure concentration estimates from LUR, dispersion and hybrid models were moderately to strongly correlated. We observed
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In conclusion, our study innovatively developed and compared three different exposure assessment methods to estimate outdoor air pollution exposures to several TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn) within the urban context of Barcelona.
+Our work provides a comprehensive evaluation of outdoor air pollution exposure, highlighting the critical role of diverse exposure assessment methods in understanding the effects and discrepancies of TRAPs exposure.
+Finally, our exposure estimates derived from the models developed will be used in the BiSC cohort and potentially in further epidemiological studies in Barcelona and its metropolitan area.
 

### 2024-02-19 supplementary_BiSC_outdoor_AP.tex
Updates from Overleaf
finished paper for co-authors

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{\textbf{Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}
+\title{\textbf{\large Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}
 
@@
 %%% --- Authors --- %%%
-\author{Alan Domínguez, Payam Dadvand, Marta Cirach, Bruno Raimbault, Toni Galmes, Karl Samuelsson, Mark Nieuwenhuijsen, Jordi Sunyer, Xavier Basagaña, Ioar Rivas}
+\author{\small Alan Domínguez, Payam Dadvand, Marta Cirach, Gustavo Arevalo, 
+Lluís Barril, María Foraster, Mireia Gascon, Bruno Raimbault, Toni Galmes, Laura Goméz-Herrera, Cecilia Persavento, Karl Samuelsson, Jose Lao, Teresa Moreno, Xavier Querol,
+Cathryn Tonne, Mark Nieuwenhuijsen, Jordi Sunyer, Xavier Basagaña, Ioar Rivas}
+
+% --- removing date from manuscript --- % 
+\date{}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Model development}
-
 Following the ESCAPE guidelines
@@
 
+\subsection{Validation LUR}
 %%% --- Table with performance metrics for Hybrid models --- %%%
@@
 The best models were selected based on its better validation performance and on the availability of the sampling campaign that could better represent the study population.
+
+\subsection{Validaiton LUR: Comparison between measurements and predictions}
 
@@
 
-\subsection{Comparison between observed measurements and predictions}
+\subsection{Validation HM: Comparison between measurements and predictions}
 

### 2024-02-21 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 %%%%%%%%%%%%%%%%%%%%%%%
+\vspace{-3em}
 \begin{abstract}
@@
 Within the framework of the Barcelona Life Study Cohort (BiSC), we developed Land use regression (LUR), dispersion (DM), and hybrid (HM) models to estimate outdoor air pollution exposure at residential addresses for several TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn).
-The overall aim is to obtain accurate weekly estimates of traffic-related air pollution, intended to be used in the exposure assessment for a cohort of pregnant women in the city of Barcelona.
-Two separate monitoring campaigns were conducted: the "BiSC-home" campaign took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} between 2018-2021.
-Additionally, a second campaign was conducted, named "BiSCAPE", and included four distinct measurement periods between 2021-2022.
-LUR and DM were constructed using conventional state-of-the-art techniques, while HM was designed by harnessing Random Forest algorithm, integrating outputs from DM with explanatory variables from the LUR model, monitoring stations and detailed traffic counts.
-Model performance was evaluated using a cross-validation procedure for LUR and HM, and by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations.
+The overall aim is developed and compared LUR, DM, amd HM for predicting outdoor air pollutiuon levels of NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} consituentes (Fe, Cu, Zn) to obtain accurate weekly estimates of TRAPs, intended to be used in the exposure assessment of BiSC cohort.
+Two separate monitoring campaigns were conducted.
+First, the "BiSC-home" measuring home-outdoor NO\textsubscript{2} levels at the home addressess of 984 BiSC participants (2018-2021).
+Second, the "BiSCAPE" campaign encompassing four distinct measurement of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC in 34 representative points across Barcelona (2021-2022).
+LUR and DM were constructed using conventional state-of-the-art techniques, while HM was designed by harnessing Random Forest (RF) algorithm, integrating outputs from DM with explanatory variables from the LUR model, monitoring stations and detailed traffic counts.
+Model performance was evaluated using a cross-validation (CV) and Leave-one-out cross-validation (LOOCV) procedure for LUR and HM, and by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations.
+Additionally, NO\textsubscript{2} models from LUR, DM, and HM, were externally validated using BiSC-home monitoring campaign.
 The models demonstrated moderate to good performance for LUR, DM and HM.
 HM outperformed LUR and DM models for NO\textsubscript{2} (R\textsuperscript{2} = 0.64), PM\textsubscript{2.5} (R\textsuperscript{2} = 0.66), and BC (R\textsuperscript{2} = 0.86).
-LUR models show better performance for PM\textsubscript{2.5 - Fe} (R\textsuperscript{2} = 0.89), PM\textsubscript{2.5 - Cu} (R\textsuperscript{2} = 0.87), PM\textsubscript{2.5 - Zn} (R\textsubscript{2} = 0.85).
-Correlations between the air pollutant estimates for the different models were moderate to strong for both long-and-short term exposures.
+LUR models show better performance for PM\textsubscript{2.5 - Fe} (R\textsuperscript{2} = 0.89), PM\textsubscript{2.5 - Cu} (R\textsuperscript{2} = 0.87), PM\textsubscript{2.5 - Zn} (R\textsuperscript{2} = 0.85).
+Correlations between outdoor air pollution estimates for the different models were moderate to strong for both long-and-short term exposures.
 Long-term exposure (entire pregnancy) at the residential addresses exhibited correlations ranging from rs = 0.48 to 0.92, while for short-term exposure (weekly exposure) varied from rs = 0.37 to 0.90.
 Our work emphasizes the importance of employing diverse exposure assessment methods to evaluate the exposure to traffic-related air pollutants and to disentangle differences in effect estimates during pregnancy.
-The exposure estimates from these different models can be used in the BiSC cohort to potentially identify critical windows of susceptibility to air pollution during pregnancy.
-\\
+The exposure estimates from these different models can be used in the BiSC cohort to potentially identify critical windows of susceptibility to air pollution.\\
 
@@
 This family of models is often used due to its ability to model potential nonlinear interaction between predictor variables \cite{liu2022treebased}.
-Among the different ML algorithms, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.
+Among the different ML algorithms, Random Forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}.
 Furthermore, the COVID-19 pandemic has introduced unprecedented challenges for air pollution modelling and exposure assessment.
@@
 
-This study was developed within the framework of the Barcelona Life Study Cohort (\href{https://projectebisc.org/en/home/}{BiSC}), a pregnancy cohort that enrolled 1,080 pregnant women in their first trimester to evaluate the impact of exposure to air pollution during pregnancy, fetal and early life development of their offspring.
-We developed LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses of the BiSC participants during their pregnancies.
+This study was developed within the framework of the  \href{https://projectebisc.org/en/home/}{Barcelona Life Study Cohort (BiSC)}, a pregnancy cohort that enrolled 1,080 pregnant women in their first trimester to evaluate the impact of exposure TRAPs during pregnancy, fetal and early life development of their offspring.
+We developed LUR, DM, and HM models to comprehensively estimate outdoor air pollution exposure at residential addresses of the BiSC participants during their pregnancies.
 LUR and DM were constructed using conventional techniques, while HM was designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model.
-The overall aim is to obtain accurate weekly estimates for NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn), intended to be used in the prenatal exposure assessment for the BiSC cohort.
+The overall aim was developed and compared LUR, DM, and HM for predicting outdoor air pollution levels of NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn) to obtain accurate weekly estimates of TRAPs,intended to be used in the exposure assessment of the BiSC cohort.
 
@@
 For a more comprehensive approach, we integrated additional data from the Catalan Air Pollution Monitoring and Forecasting Network (XVPCA) for Barcelona and surrounding areas \cite{xarxa2012}.
-Daily air pollution concentration levels were interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}.
+Daily air pollution concentration levels were interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017}.
 The data were collected from 16 monitoring stations, including Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St.
@@
 Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality monitoring stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others.
-Additionally, we compared the results of the model with data from the 4 campaigns carried out at BiSCAPE sites.
-In total, 98 periods were used to evaluate the performance of NO\textsubscript{2} model, 66 periods for PM\textsubscript{2.5}, and 93 periods for BC.
+Additionally, we compared the results of the model with data from the BiSC-home and BiSCAPE campaign to externally validated the models.
+We use both campaign to validate the results of the NO\textsubscript{2} model and only the BiSCAPE campaign to validate PM\textsubscript{2.5} and BC models.
+In total, 98 periods were used to evaluate the performance of NO\textsubscript{2} model, 66 periods for PM\textsubscript{2.5}, and 93 periods for BC from the BiSCAPE campaign.
 The closer these values are, the better the model's performance.
@@
 Secondly, external validation was performed by allocating 20\% of the data as a test set to further evaluate the model's performance against previously unseen data.
-While both methods differ in their approach to partitioning the data, the fundamental distinction lies in how the data is utilized during the training and validation phases.
+While both methods differ in their approach to partitioning the data, the fundamental distinction lies in how the data is utilized during the training and validation process.
 The \textit{10-fold CV} method systematically rotates through the data, ensuring that each subset serves as a validation set once, while the external validation approach provides a straightforward assessment against a separate data portion.
@@
 \begin{figure}[!h]
-\includegraphics[width=0.75\textwidth]{figures/combined_figure1_v2.png}
+\includegraphics[width=0.8\textwidth]{figures/combined_figure1_v3.png}
 \caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
@@
 \begin{figure}[!h]
-\includegraphics[width=0.75\textwidth]{figures/combined_figure2_v3.png}
+\includegraphics[width=0.8\textwidth]{figures/combined_figure2_v3.png}
 \caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and HM (\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
@@
 LUR models shown higher variability in exposure concentrations during pregnancy, whereas DM exhibit less variability.
-HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.
+HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.\\
 
@@
     \toprule
-         Model & Validation & \multicolumn{6}{c}{R\textsuperscript{2 \textsubscript{ Validation}}} \\
+         Model & Validation & \multicolumn{6}{c}{R\textsuperscript{2 \textsubscript{Validation}}} \\
          \cmidrule(lr){3-8}
@@
          LUR & External validation BiSC-Home & 0.56 & - & - & - & - & - \\
-         DM & External validation BiSCAPE & 0.65 & 0.22 & 0.56 & - & - & - \\
-         HM & External validation split (test data) & 0.64 & 0.65 & 0.69 & 0.65 & 0.56 & 0.36  \\
+         \midrule % Separating LUR from DM
+         \multirow{2}{*}{DM} & External validation BiSCAPE & 0.65 & 0.22 & 0.56 & - & - & - \\
+         & External validation BiSC-Home & 0.44 & - & - & - & - & - \\ 
+         \midrule % Separating DM from HM
+         \multirow{2}{*}{HM} & External validation split (test data) & 0.64 & 0.65 & 0.69 & 0.65 & 0.56 & 0.36 \\
+         & External validation BiSC-Home & 0.64 & - & - & - & - & - \\ 
     \bottomrule
@@
 NO\textsubscript{2} LUR model was validated using BiSC - home campaign data.
-NO\textsubscript{2}, PM\textsubscript{2.5} and BC from DM model were validated using data from BiSCAPE monitoring campaign.
-While HM models were validated using 20\% of data available initially before model training.
-Due to lack of data availability we were able to evaluate the performance of NO\textsubscript{2} for LUR models.
+NO\textsubscript{2}, PM\textsubscript{2.5} and BC from DM model were validated using data from BiSCAPE and BiSC-home monitoring campaigns.
+While HM models were validated using 20\% of data available initially before model training and BiSC-home campaign.
 The performance of the LUR model decreased in the external validation by 16\%.
@@
 While for BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} performance varied between validation approaches by -19.7\%, +20.37\%, -20.0\% , and -18.8\% respectively.
+Additionally, we use the BiSC - home campaign to validate and compare NO\textsubscript{2} models.
+To do that we develop and additional HM using only observations from the locations that had measured home-outdoor levels for both sampling campaigns (first and third trimesters) to make possible compare the performance of the three different approaches.
+We observed that HM model explain more variance than LUR and DM models.
 
@@
 %%% ------------------------------------------%%%%
+\vspace{-2em}
 \begin{figure}[!h]
-\includegraphics[width=0.8\textwidth]{figures/boxplot_all_models_estimates.png}
+\includegraphics[width=1\textwidth]{figures/boxplot_all_models_estimates.png}
 \caption{Distribution of estimated exposure to outdoor air pollution for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
@@
 For PM\textsubscript{2.5} constituents we observed higher correlation coefficients for PM\textsubscript{2.5 - Fe} in LUR and HM for both long and short term exposure.
-Estimates for PM\textsubscript{2.5 - Cu} and PM\textsubscript{2.5 - Zn} showed lower correlations for both methods compared to Fe constituent.
+Estimates for PM\textsubscript{2.5 - Cu} and PM\textsubscript{2.5 - Zn} showed lower correlations for both methods compared to PM\textsubscript{2.5 - Fe} constituent.
 Overall, NO\textsubscript{2} and BC from LUR, DM, and HM were strongly correlated, while for PM\textsubscript{2.5} and constituents the correlation was weaker.
@@
 \includegraphics[width=1.0\textwidth]{figures/correlation_fig_v2.png}
-\caption{Correlation between traffic-related air pollutants for LUR, DM, HM in different windows of exposure \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
+\caption{Correlation between traffic-related air pollutants for LUR, DM, HM in different windows of exposure \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{PM$_{2.5 - Fe}$}, \textit{PM$_{2.5 - Cu}$}, and \textit{PM$_{2.5 - Zn}$}.}
 \label{fig5}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{-1em}
 
@@
 \begin{figure}[!h]
-\includegraphics[width=0.7\textwidth]{figures/final_bland_altmant.png}
-\caption{Bland-Altmant plots for comparing outdoor air pollution estimates for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, and Zn).
-using LUR, DM, and HM}}
+\includegraphics[width=0.58\textwidth]{figures/final_bland_altmant.png}
+\caption{Bland-Altmant plots for comparing outdoor air pollution estimates for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{PM$_{2.5 - Fe}$}, \textit{PM$_{2.5 - Cu}$}, and \textit{PM$_{2.5 - Zn}$}, using LUR, DM and HM.}
 \label{fig6}
@@
 % --- General --- % 
-We successfully developed outdoor air pollution models (NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents) to assess 
-exposure within the city of Barcelona for the period 2018-2021.
+We successfully developed outdoor air pollution models to TRAPs (NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents) to assess exposure within the city of Barcelona for the period 2018-2021.
 Utilizing three different approaches: land use regression, dispersion modeling, and hybrid models were tailored to provide exposure estimates across various time windows releveant to our study population in the BiSC cohort, including entire pregnancy duration, specific trimesters, and weeks of pregnancy for each participant.
 The performance of these models was moderate to good, as determined by different evaluation metrics.
-The LUR models accounted for approximately 45\% to 89\% of the variance in pollutant levels, indicating their strength in capturing spatial variability in air pollution concentrations.
-HM models, which integrate a broader range of data inputs and modeling techniques, explained 44\% to 86\% of the variance.
+The LUR models accounted for approximately 45\% to 83\% of explanied variance in NO\textsubscript{2}, PM\textsubscript{2.5} and BC levels, and between 85\% and 85\% in PM\textsubscript{2.5} constituents, indicating their strength in capturing spatial variability in air pollution concentrations.
+HM models, which integrate a broader range of data inputs and modeling techniques, explained between 64\% and 86\% of the variance in NO\textsubscript{2}, PM\textsubscript{2.5} and BC levels, and between 44\% and 77\% in PM\textsubscript{2.5} constituents levels.
 In contrast, DM models, focusing on physical and chemical processes affecting pollutants, explained a lower range of variability, between 26\% and 65\%.
@@
 % --- NO2 models --- %
-The performance of our LUR NO\textsubscript{2} model was higher than previous European models \cite{rivera2012} as well as earlier models specific to Barcelona \cite{criado2023}, indicating an enhanced capacity to capture the intra-urban variability of NO\textsubscript{2} concentrations levels.
+The performance of our LUR NO\textsubscript{2} model was higher than previous European models \cite{rivera2012} as well as earlier models specific to Barcelona \cite{criado2023}, indicating an enhanced capacity of our model to capture the intra-urban variability in NO\textsubscript{2} concentrations levels.
 Notably, our LUR model outperformed a recently developed microscale LUR model, which was part of a data fusion method designed to minimize prediction uncertainty by integrating a LUR base map with hourly dispersion estimates from the CALLIOPE-Urban model within Barcelona \cite{criado2023}.
@@
 Conversely, our DM model exhibits a lower performance compared to the CALLIOPE-Urban system, a regional-to-street-scale modelling framework that offers high-spatial-and-high-temporal-resolution (10m x 10m, hourly) by integrating the mesoscale airquality forecasting system CALIOPE (WRF-HERMES-CMAQ-BSS-DREAM8b) with an urban roadwide dispersion model (R-LINE) \cite{benavides2019}.
-Regarding our HM, where we integrated detailed information on traffic counts across the city of Barcelona, air pollution and meteorological data from monitoring stations, and estimates from the DM model through a RF algorithm, we achieved better performance compared to both our LUR and DM models.
+Regarding our HM, where we integrated detailed information on traffic counts across the city of Barcelona, air pollution and meteorological data from monitoring stations, and estimates from the DM model through a RF algorithm, we achieved a better performance compared to both our LUR and DM models.
 Most relevant predictor variables were estimates from the DM, monitoring station, traffic counts, solar radiation traffic-related spatial variables.
@@
 Our lUR model uses observed data and land use characteristics to estimate air pollution levels, potentially offering a finer spatial resolution that better captures local variations, especially within urban settings \cite{hoek2017}.
-Since there were no previously available published DM models PM\textsubscript{2.5} developed in the city of Barcelona or at a Catalunya scale we were not able to compare the performance.
+Since there were no previously available published DM models PM\textsubscript{2.5} developed in the city of Barcelona or at a Catalunya scale we were not able to compare their performance.
 The enhanced performance of our PM\textsubscript{2.5} HM model compared to our previous approaches can largely be attributed to the incorporation of comprehensive information during the model-building process and  the employment of the RF algorithm significantly increased our model's accuracy in predicting air pollution levels.
 This approach enabled us to account for additional sources of variability in PM\textsubscript{2.5} concentration levels, identifying emissions sources not captured by the other models.
-Additionally, our HM model demonstrated improved performance metrics compared to a previously developed RF model \cite{mila2023}.
+Additionally, our HM model demonstrated improved performance metrics compared to a previously developed RF model \cite{mila2023}.However, drawing definitive conclusions from this comparison is challenging, as their model was specifically evaluated for its ability to predict air pollution levels using data from the monitoring network across Catalonia.
 Regardless of these advancements in modeling PM\textsubscript{2.5} concentrations by our HM, it was found that recent European multi-stage HM models, utilizing RF algorithm for predicting daily PM\textsubscript{2.5} levels at a country-level with a high spatiotemporal resolution, exhibit superior performance metrics.
@@
 This difference in model performance can be attributed to two main factors.
-First, the comprehensive exposure assessment period employed by these models, which utilized routine several years of data from routine monitoring stations to more effectively capture temporal trends in air pollution exposure levels.
+First, the comprehensive exposure assessment period employed by these models, which utilized several years of data from routine monitoring stations to more effectively capture temporal trends in air pollution exposure levels.
 Second, the inclusion of AOD measurements provided additional insights into the background and regional variability of air pollution levels with higher correlations in Western Europe compared to the rest of the Region \cite{chen2024}.\\
@@
 Predictor variables used for the development of LUR and HM models were mainly traffic, residential, and population density-related.
-We believe that the high performance in both LUR and HM is that BC levels proportionally reflect traffic-related process  \\
+We attribute the performance of both the LUR and HM models to their ability to capture the traffic-related processes affecting BC levels throughout the city of Barcelona.
+This alignment suggest that BC concentrations proportionally reflect traffic-related emissions, underscoring the critical role of traffic sources in urban air pollution in the city of Barcelona.\\
 
@@
 Despite this, we assess the performance of the NO\textsubscript{2} model given some lights regarding the most suitable approach for our context.
-However, we can no extending these findings to other pollutants due to differences in their sources, behavior, and environmental interactions.
-Despite these challenges, our study lays the groundwork for future research in exposure assessment in urban areas, providing a basis for methodological refinement and encouraging the development of more generalized models that can be  externally validated.
+However, we can no extend these findings to other pollutants due to differences in their sources, behavior, and environmental interactions.
+Despite these challenges, our study provides a basis for methodological refinement and encouraging the development of more generalized models that can be  externally validated.
 
@@
 
-In conclusion, our study innovatively developed and compared three different exposure assessment methods to estimate outdoor air pollution exposures to several TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn) within the urban context of Barcelona.
+In conclusion, our study developed and compared three different exposure assessment methods to estimate outdoor air pollution exposures to several TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn) within the urban context of Barcelona.
 Our work provides a comprehensive evaluation of outdoor air pollution exposure, highlighting the critical role of diverse exposure assessment methods in understanding the effects and discrepancies of TRAPs exposure.

### 2024-02-21 supplementary_BiSC_outdoor_AP.tex
Updates from Overleaf

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{\textbf{\large Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}
+\title{\textbf{\Large Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}
 
@@
 \maketitle
+
+\vspace*{-1cm}
 
